---
title: "Hierarchical analysis of hourly values"
format: html
author: Johannes Zauner
editor_options: 
  chunk_output_type: console
---

## Preface

This script uses hierarchical generalized additive model to explore the following hypothesis:

> Personal light exposure patterns depend on wearing position

*Type of model:* HGAM (Hierarchical GAM)

*Base formula:* $\text{melanopic EDI} \sim  s(\text{Time}) + s(\text{Time}, \text{position}) + s(\text{Time}, \text{Site}) + s(\text{Time}, \text{Participant})$

*Additional notes:* 
- position and site effects are modelled with the `sz` basis, that allows for the most identifiable modelling of parametric differences
- Participant effects are modelled with the `fs` basis, which models the participant-effects as random smooths
- melanopic EDI is scaled with `log_zero_inflated()` before model fitting.

## Setup

```{r}
library(tidyverse)
library(mgcv)
library(LightLogR)
library(here)
library(gratia)
library(itsadug)
```

### Preparation

```{r}
#| filename: Load prepared data
load(here("data/prepared_time_data.RData"))
time_data <- 
  time_data |> 
  mutate(position = fct(position))
```

```{r}
#| filename: Overview plot of the data (median with IQR)
time_data|>
  group_by(position) |> 
  aggregate_Date(
    date.handler = \(x) ymd("2025-01-01"),
    numeric.handler = \(x) median(x, na.rm = TRUE),
    p75 = quantile(MEDI, p=0.75, na.rm = TRUE),
    p25 = quantile(MEDI, p=0.25, na.rm = TRUE)
  ) |> 
  gg_doubleplot(geom = "blank", jco_color = FALSE, facetting = FALSE,
                x.axis.breaks.repeat = ~Datetime_breaks(.x, by = "12 hours", shift =
    lubridate::duration(0, "hours"))) +
  geom_ribbon(aes(ymin = p25, ymax = p75, fill = position), alpha = 0.5) +
  geom_line(aes(color = position)) +
  geom_hline(yintercept = 250, linetype = "dashed") +
  scale_color_viridis_d()+
  scale_fill_viridis_d() +
  guides(fill = "none") +
  theme(panel.spacing = unit(2, "lines"))
```

```{r}
#| filename: H11 formulas
formula_1 <- lzMEDI ~ s(Time, k = 12) + s(Time, position, bs = "sz", k = 12) + s(Time, site, bs = "sz", k = 12) + s(Time, Id, bs = "fs") + s(Id_date, bs = "re")
formula_0 <- lzMEDI ~ s(Time, k = 12) + s(Time, site, bs = "sz", k = 12) + s(Time, Id, bs = "fs") + s(Id_date, bs = "re")
```

### Analysis

#### Base model

```{r}
#| filename: Main model
model_1 <- 
  bam(formula_1, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10
      )

r1 <- start_value_rho(model_1, 
                      plot=TRUE)

model_1 <- 
  bam(formula_1, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10,
      rho = r1,
      AR.start = AR.start
      )

acf_resid(model_1, split_pred = "AR.start")

summary_1 <- summary(model_1)
summary_1
model_1 |> glance()

model_1 |> appraise()
model_1 |> draw(select = "s(Time,position)")
```

```{r}
#| filename: Variance explained
Xp <- predict(model_1, type = "terms")
variance <- apply(Xp, 2, var)
cor(Xp)
R2 <- (variance/sum(variance)) |> enframe(value = "R2", name = "parameter")
R2
```

#### Null model (H11_0)

```{r}
#| filename: Null model
model_0 <- 
  bam(formula_0, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 8
      )

r1 <- start_value_rho(model_0, 
                      plot=TRUE)

model_0 <- 
  bam(formula_0, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10,
      rho = r1,
      AR.start = AR.start
      )

acf_resid(model_0, split_pred = "AR.start")

summary_0 <- summary(model_0)
summary_0
model_0 |> glance()

model_0 |> appraise()
model_0 |> draw()
```

#### Model comparison

```{r}
AICs <- AIC(model_1, model_0)
AICs |> as_tibble(rownames = "model") |> gt()
```

> While position only explains very little of the overall structure (<1% of variance explained), is has a ∆AIC >2 and will thus be included

#### Model visualization

```{r}
#| filename: prepare photoperiod model H2 plot
H11_sex_data <- 
H11_model_1 |> 
  conditional_values(
    condition = list(Time = seq(0.25, 23.75, by = 0.5), "sex"),
    exclude = c("s(Time,site)", "s(Time,Id)", "s(Id_date)")
    ) |> 
  mutate(across(c(.fitted, .lower_ci, .upper_ci),
                exp_zero_inflated)
  )

```

```{r}
#| filename: Visualize sex model H11
#| fig-height: 6
#| fig-width: 10
source("scripts/Brown_bracket.R")

H11_pred <-
H11_sex_data |> 
  ggplot(aes()) +
  map(c(1,10,250, 10^4), 
          \(x) geom_hline(aes(yintercept = x), 
                          col = "grey", linetype = "dashed")
      ) +
  geom_ribbon(aes(x = Time, ymin = .lower_ci, ymax = .upper_ci, fill = sex), alpha = 0.5) +
  geom_line(aes(x = Time, y = .fitted, col = sex), linewidth = 0.1) +
  geom_line(aes(x = Time, y = .fitted), linewidth = 1) +
  theme_cowplot() +
  scale_y_continuous(trans = "symlog", breaks = c(0,1,10,100,250, 1000)) +
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  coord_cartesian(xlim = c(0,24), ylim = c(0,10^4), expand = FALSE) +
  labs(y = "Melanopic EDI (lx)", x = "Local time (hrs)", fill = "Sex", col = "Sex") +
  facet_wrap(~sex) +
  # guides( y = guide_axis_stack(Brown_bracket, "axis")) +
  gghighlight(
    label_key = sex,
    label_params = list(x = 12, y = 4, force = 0, hjust = 0.25)
  ) +
  theme_sub_strip(background = element_blank(),
                  text = element_blank()) +
  theme(panel.spacing = unit(1, "lines"),
        plot.caption = element_textbox_simple())
  # labs(caption = Brown_caption)

```

```{r}
#| filename: Visualize model terms
H11_P_terms <-
H11_model_1 |> 
  smooth_estimates(select = "s(Time,sex)") |> 
  mutate(.lower_ci = .estimate -1.96*.se,
         .upper_ci = .estimate +1.96*.se,
         .fitted = .estimate,
         signif = .lower_ci > 0 | .upper_ci < 0,
         signif_id = consecutive_id(signif),
         .by = sex) |> 
  ggplot(aes()) +
  map(c(log10(c(0.1, 0.2, 0.5, 2, 5, 10))),
      \(x) geom_hline(yintercept = x, linewidth = 0.5, col = "grey", linetype = 2)) +
  geom_ribbon(aes(x = Time, ymin = .lower_ci, ymax = .upper_ci, fill = sex), alpha = 0.5) +
  geom_hline(yintercept = 0, linewidth = 1, linetype = 2) +
  geom_line(aes(x = Time, y = .fitted), 
            linewidth = 1) +
  geom_line(data = \(x) filter(x, signif),
            aes(x = Time, y = .fitted, group = signif_id), 
            col = "red", linewidth = 1) +
  # scale_color_manual(values = c(melidos_colors)) +
  theme_cowplot() +
  scale_y_continuous(labels = expression(10^-1, 5^-1, 2^-1, "1  ", "2  ", "5  ", "10  "), 
                     breaks = log10(c(0.1, 0.2, 0.5, 1, 2, 5, 10)))+
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  coord_cartesian(xlim = c(0,24), expand = FALSE) +
  # guides(fill = "none") +
  labs(y = "Factor", x = "Local time (hrs)", fill = "Sex") +
  facet_wrap(~sex) +
  theme_sub_strip(background = element_blank(),
                  text = element_blank()) +
  theme(panel.spacing = unit(1, "lines"),
        plot.caption = element_textbox_simple(),
        legend.position = "bottom")

```

```{r}
H11_P_terms
ggsave("figures/Fig8.pdf")
```

### Interpretation

- Sex is a minor differentiator in terms of melanopic stimulus, that explanes less than 1% of the variance. The only significant difference between the sexes is a lower value during the morning for females compared to males. This phase lasts about 2 hours and has a peak difference of about factor 2

## Session Info

```{r}
sessionInfo()
```

