---
title: "Hierarchical analysis of hourly values"
format: html
author: Johannes Zauner
editor_options: 
  chunk_output_type: console
---

## Preface

This script uses hierarchical generalized additive model to explore the following hypothesis:

> Personal light exposure patterns depend on wearing position

*Type of model:* HGAM (Hierarchical GAM)

*Base formula:* $\text{melanopic EDI} \sim  s(\text{Time}) + s(\text{Time}, \text{position}) + s(\text{Time}, \text{Site}) + s(\text{Time}, \text{Participant})$

*Additional notes:* 
- position and site effects are modelled with the `sz` basis, that allows for the most identifiable modelling of parametric differences
- Participant effects are modelled with the `fs` basis, which models the participant-effects as random smooths
- melanopic EDI is scaled with `log_zero_inflated()` before model fitting.

## Setup

```{r}
library(tidyverse)
library(mgcv)
library(LightLogR)
library(here)
library(gratia)
library(broom)
library(itsadug)
library(cowplot)
library(gghighlight)
library(ggsci)
library(gt)
```

### Preparation

```{r}
#| filename: Load prepared data
colors <- ggsci::pal_jco()(3) |> set_names(c("glasses", "chest", "wrist"))
load(here("data/prepared_time_data.RData"))
time_data <- 
  time_data |> 
  mutate(position = fct(position, levels = c("glasses", "chest", "wrist")))
```

```{r}
#| filename: Overview plot of the data (median with IQR)
time_data|>
  group_by(position) |> 
  mutate(position = fct_relevel(position, "glasses")) |> 
  aggregate_Date(
    date.handler = \(x) ymd("2025-01-01"),
    numeric.handler = \(x) median(x, na.rm = TRUE),
    p75 = quantile(MEDI, p=0.75, na.rm = TRUE),
    p25 = quantile(MEDI, p=0.25, na.rm = TRUE)
  ) |> 
  gg_doubleplot(geom = "blank", jco_color = FALSE, facetting = FALSE,
                x.axis.breaks.repeat = ~Datetime_breaks(.x, by = "12 hours", shift =
    lubridate::duration(0, "hours"))) +
  geom_ribbon(aes(ymin = p25, ymax = p75, fill = position), alpha = 0.5) +
  geom_line(aes(color = position)) +
  geom_hline(yintercept = 250, linetype = "dashed") +
  scale_color_jco()+
  scale_fill_jco() +
  guides(fill = "none") +
  theme(panel.spacing = unit(2, "lines"))
```

```{r}
#| filename: H11 formulas
formula_1 <- lzMEDI ~ s(Time, k = 12) + s(Time, position, bs = "sz", k = 12) + s(Time, site, bs = "sz", k = 12) + s(Time, Id, bs = "fs") + s(Id_date, bs = "re")
formula_0 <- lzMEDI ~ s(Time, k = 12) + s(Time, site, bs = "sz", k = 12) + s(Time, Id, bs = "fs") + s(Id_date, bs = "re")
```

### Analysis

#### Base model

```{r}
#| filename: Main model
model_1 <- 
  bam(formula_1, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10
      )

r1 <- start_value_rho(model_1, 
                      plot=TRUE)

model_1 <- 
  bam(formula_1, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10,
      rho = r1,
      AR.start = AR.start
      )

acf_resid(model_1, split_pred = "AR.start")

summary_1 <- summary(model_1)
summary_1
model_1 |> glance()

model_1 |> appraise()
model_1 |> draw()
```

```{r}
#| filename: Variance explained
Xp <- predict(model_1, type = "terms")
variance <- apply(Xp, 2, var)
cor(Xp)
R2 <- (variance/sum(variance)) |> enframe(value = "R2", name = "parameter")
R2
```

#### Null model (H11_0)

```{r}
#| filename: Null model
model_0 <- 
  bam(formula_0, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 8
      )

r1 <- start_value_rho(model_0, 
                      plot=TRUE)

model_0 <- 
  bam(formula_0, 
      data = time_data, 
      discrete = TRUE,
      method = "fREML",
      nthreads = 10,
      rho = r1,
      AR.start = AR.start
      )

acf_resid(model_0, split_pred = "AR.start")

summary_0 <- summary(model_0)
summary_0
model_0 |> glance()

model_0 |> appraise()
model_0 |> draw()
```

#### Model comparison

```{r}
AICs <- AIC(model_1, model_0)
AICs |> as_tibble(rownames = "model") |> gt()
```

> While position only explains very little of the overall structure (<1% of variance explained), is has a ∆AIC >2 and will thus be included

#### Model visualization

```{r}
#| filename: prepare model plot
model_data <- 
model_1 |> 
  conditional_values(
    condition = list(Time = seq(0.25, 23.75, by = 0.5), "position"),
    exclude = c("s(Time,site)", "s(Time,Id)", "s(Id_date)")
    ) |> 
  mutate(across(c(.fitted, .lower_ci, .upper_ci),
                exp_zero_inflated)
  )

```

```{r}
#| filename: Visualize position model
#| fig-height: 6
#| fig-width: 10
source(here("scripts/Brown_bracket.R"))

# model_pred <-
model_data |> 
  mutate(position = fct_relevel(position, "glasses")) |> 
  ggplot(aes()) +
  map(c(1,10,250, 10^4), 
          \(x) geom_hline(aes(yintercept = x), 
                          col = "grey", linetype = "dashed")
      ) +
  geom_ribbon(aes(x = Time, ymin = .lower_ci, ymax = .upper_ci, fill = position), alpha = 0.5) +
  geom_line(aes(x = Time, y = .fitted, col = position), linewidth = 0.1) +
  geom_line(aes(x = Time, y = .fitted), linewidth = 1) +
  theme_cowplot() +
  scale_y_continuous(trans = "symlog", breaks = c(0,1,10,100,250, 1000)) +
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  coord_cartesian(xlim = c(0,24), ylim = c(0,10^4), expand = FALSE) +
  labs(y = "Melanopic EDI (lx)", x = "Local time (hrs)", fill = "position", col = "Position") +
  facet_wrap(~position) +
  scale_color_jco()+
  scale_fill_jco() +
  # guides( y = guide_axis_stack(Brown_bracket, "axis")) +
  gghighlight(
    label_key = position,
    label_params = list(x = 12, y = 4, force = 0, hjust = 0.25)
  ) +
  theme_sub_strip(background = element_blank(),
                  text = element_blank()) +
  theme(panel.spacing = unit(1, "lines"),
        plot.caption = element_textbox_simple())
  # labs(caption = Brown_caption)

```

```{r}
#| filename: helper function
difference_sz <- function(model, data, x, fac, lev1, lev2,
                          sz_term,
                          exclude = NULL,
                          unconditional = FALSE,
                          level = 0.95) {

  data1 <- data
  data2 <- data

  fac_levels <- levels(model.frame(model)[[fac]])
  
  data1[[fac]] <- factor(rep(lev1, nrow(data1)), levels = fac_levels)
  
  data2[[fac]] <- factor(rep(lev2, nrow(data2)), levels = fac_levels)
  
  X1 <- predict(
    
    model, 
    newdata = data1,
    type = "lpmatrix",
    exclude = exclude,
    newdata.guaranteed = TRUE
  )

  X2 <- predict(
    model, 
    newdata = data2,
    type = "lpmatrix",
    exclude = exclude,
    newdata.guaranteed = TRUE
  )

  Xd <- X1 - X2

  ## keep only the sz smooth columns
  sm_labels <- vapply(model$smooth, `[[`, character(1), "label")
  i <- match(sz_term, sm_labels)

  if (is.na(i)) {
    stop(
      "Smooth not found. Available smooth labels are:\n",
      paste(sm_labels, collapse = "\n")
    )
  }

  keep <- model$smooth[[i]]$first.para:model$smooth[[i]]$last.para
  Xd[, setdiff(seq_len(ncol(Xd)), keep)] <- 0

  beta <- coef(model)
  V <- vcov(model, unconditional = unconditional)

  fit <- drop(Xd %*% beta)
  se <- sqrt(rowSums((Xd %*% V) * Xd))

  crit <- qnorm(1 - (1 - level) / 2)

  tibble(
    {{ x }} := data[[x]],
    comparison = paste0(lev1, "_", lev2),
    diff = fit,
    se = se,
    lower = fit - crit * se,
    upper = fit + crit * se
  )
}
```


```{r}
#| filename: calculate the model comparison
df <- model_1$model |> distinct(Time, .keep_all = TRUE)

model_comparison <- 
rbind(
difference_sz(model_1, df, "Time", "position", "chest", "glasses",
              "s(Time,position)",
              exclude = c("s(Time,site)", "s(Time,Id)", "s(Id_date)"),
              unconditional = FALSE,
              level = 0.95
),
difference_sz(model_1, df, "Time", "position", "wrist", "glasses", 
              "s(Time,position)",
              exclude = c("s(Time,site)", "s(Time,Id)", "s(Id_date)"),
              unconditional = FALSE,
              level = 0.95
)
) |> 
  mutate(position = str_remove(comparison, "_glasses"))

```


```{r}
#| filename: Visualize model terms
model_terms <-
model_comparison |> 
  mutate(
         signif = lower > 0 | upper < 0,
         signif_id = consecutive_id(signif),
         .by = comparison) |> 
  ggplot(aes()) +
  map(c(log10(seq(0.8, 1.2, by = 0.1))),
      \(x) geom_hline(yintercept = x, linewidth = 0.5, col = "grey", linetype = 2)) +
  geom_ribbon(aes(x = Time, ymin = lower, ymax = upper, fill = position), alpha = 0.5) +
  geom_hline(yintercept = 0, linewidth = 1, linetype = 2) +
  geom_line(aes(x = Time, y = diff), 
            linewidth = 1) +
  geom_line(data = \(x) filter(x, signif),
            aes(x = Time, y = diff, group = signif_id), 
            col = "red", linewidth = 1) +
  # scale_color_manual(values = c(melidos_colors)) +
  theme_cowplot() +
  scale_fill_manual(values = colors[-1]) +
  scale_y_continuous(labels = c("-20%", "-10%", "±0%", "+10%", "+20%"),
                     breaks = log10(seq(0.80, 1.20, by = 0.1)))+
  scale_x_continuous(breaks = seq(0, 24, by = 6)) +
  coord_cartesian(xlim = c(0,24), expand = FALSE) +
  guides(fill = "none") +
  labs(y = "Factor", x = "Local time (hrs)", fill = "Position") +
  facet_wrap(~position) +
  theme(panel.spacing = unit(1, "lines"),
        plot.caption = element_textbox_simple(),
        legend.position = "bottom")

```

```{r}
model_terms
ggsave(here("output/figures/Fig_time_diff.pdf"))
```

### Interpretation

- There are significant differences between glasses and chest in the afternoon, growing larger towards the evening in the 8 to 10% range. There are significant differences between glasses and wrist during most waking hours and in the 8 to 15% range. Both chest and wrist tend to underestimate the glasses position.

## Session Info

```{r}
sessionInfo()
```

