---
title: "Exploratory analysis light exposure data"
authors: "Sietse de Vries"
---

## Preface

This is a (work in progress) exploratory analysis of the light exposure dataset gathered in Izmir, Türkiye by Didikoglu et al. 2025.

```{r}
#| include: false
library(dplyr)
library(lubridate)
library(ggplot2)
library(tidyr)
library(cowplot)
library(rio)
library(LightLogR)
library(here)

tz = "Europe/Istanbul"
```

## Data import

The data is imported directly from the Github page [Didikogluetal2025](https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/tree/v1.0.1).

**Q: For the final notebook, do we want to upload all the used data on our "own" Github page? Or do we directly import the datasets from the Github pages of each location?**

As this is a test, only the first 3 participants are considered.

```{r}
# Head
filepaths_h <- sapply(1:3, function(i) {
  id <- sprintf("%03d", i)
  paste0(
    "https://raw.githubusercontent.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/",
    "refs/tags/v1.0.1/data/raw/individual/IZTECH_S", id,
    "/continuous/actlumus_head/IZTECH_S", id, "_h.txt"
  )
})

pattern <- "(IZTECH_S\\d+)(?=_h)"
data_h <- import$ActLumus(filepaths_h, tz = tz, auto.id = pattern)


# Chest
filepaths_c <- sapply(1:3, function(i) {
  id <- sprintf("%03d", i)
  paste0(
    "https://raw.githubusercontent.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/",
    "refs/tags/v1.0.1/data/raw/individual/IZTECH_S", id,
    "/continuous/actlumus_chest/IZTECH_S", id, "_c.txt"
  )
})

pattern <- "(IZTECH_S\\d+)(?=_c)"
data_c <- import$ActLumus(filepaths_c, tz = tz, auto.id = pattern)


# Wrist
filepaths_w <- sapply(1:3, function(i) {
  id <- sprintf("%03d", i)
  paste0(
    "https://raw.githubusercontent.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/",
    "refs/tags/v1.0.1/data/raw/individual/IZTECH_S", id,
    "/continuous/actlumus_wrist/IZTECH_S", id, "_w.txt"
  )
})

pattern <- "(IZTECH_S\\d+)(?=_w)"
data_w <- import$ActLumus(filepaths_w, tz = tz, auto.id = pattern)

# Simplify all 3 dataframes
data_h <- data_h[, c("Id", "Datetime", "MEDI")]|> mutate(position = "h")
data_c <- data_c[, c("Id", "Datetime", "MEDI")]|> mutate(position = "c")
data_w <- data_w[, c("Id", "Datetime", "MEDI")]|> mutate(position = "w")
```

## Data cleaning

Observation outside of the trial period (of each participant) are removed.

```{r}
study_dates <- import("https://raw.githubusercontent.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/v1.0.1/data/Study_dates_MeLiDos_IZTECH.xlsx")

study_dates <-
  study_dates |>
  select(subjectID_device, datetime_trial_start, datetime_trial_end) |>
  separate(subjectID_device, into = c("Id", "position"), sep = "_(?=[^_]+$)") |>
  filter(position %in% c("h","c","w")) |>
  mutate(
    datetime_trial_start = force_tz(datetime_trial_start, "Europe/Istanbul"),
    datetime_trial_end   = force_tz(datetime_trial_end, "Europe/Istanbul")
  )


trial_filter <- function(data, study_dates) {
  data |>
    inner_join(study_dates, by = c("Id", "position")) |>
    filter(Datetime >= datetime_trial_start & Datetime <= datetime_trial_end) |>
    select(-datetime_trial_start, -datetime_trial_end)
}

data_h <- trial_filter(data_h, study_dates)
data_c <- trial_filter(data_c, study_dates)
data_w <- trial_filter(data_w, study_dates)
```

Next, days with more than 30 minutes of missing data are excluded, ensuring that only complete measurement days are included in the analysis.

```{r}
# Head
data_h <- 
  data_h |> 
  remove_partial_data(
    handle.gaps = TRUE, MEDI, by.date = TRUE, threshold.missing = "30 minutes"
  )

data_h |> gg_overview()
data_h |> has_gaps()
data_h |> has_irregulars()


# Chest
data_c <- 
  data_c |> 
  remove_partial_data(
    handle.gaps = TRUE, MEDI, by.date = TRUE, threshold.missing = "30 minutes"
  )

data_c |> gg_overview()
data_c |> has_gaps()
data_c |> has_irregulars()


# Wrist
data_w <- 
  data_w |> 
  remove_partial_data(
    handle.gaps = TRUE, MEDI, by.date = TRUE, threshold.missing = "30 minutes"
  )

data_w |> gg_overview()
data_w |> has_gaps()
data_w |> has_irregulars()
```

## Datetime synchronization

The first and last full day of each dataset (h, c, and w) are identified. If any days are missing in one of the datasets for a given participant, the corresponding days are removed from the head dataset (as this dataset will serve as the "ground truth").

```{r}
# Head
date_range_h <- data_h |> 
  group_by(Id) |> 
  summarise(
    start_day_h = as.Date(min(Datetime, na.rm = TRUE), tz = tz),
    end_day_h = as.Date(max(Datetime, na.rm = TRUE), tz = tz))

# Chest
date_range_c <- data_c |> 
  group_by(Id)|> 
  summarise(
    start_day_c = as.Date(min(Datetime, na.rm = TRUE), tz = tz),
    end_day_c = as.Date(max(Datetime, na.rm = TRUE), tz = tz))

# Wrist
date_range_w <- data_w |> 
  group_by(Id) |> 
  summarise(
    start_day_w = as.Date(min(Datetime, na.rm = TRUE), tz = tz),
    end_day_w = as.Date(max(Datetime, na.rm = TRUE), tz = tz))


common_dates <- date_range_h |> 
  full_join(date_range_c, by = "Id") |> 
  full_join(date_range_w, by = "Id")|> 
  mutate(
    # Find the latest available start and earliest available end
    common_start_day = pmax(start_day_h, start_day_c, start_day_w),
    common_end_day   = pmin(end_day_h, end_day_c, end_day_w)
  )

# Remove any rows from the data_h dataframe that are outside of the common start/end
data_h <- data_h |> 
  mutate(temp_date = as.Date(Datetime, tz=tz)) |> 
  left_join(
    common_dates |>  select(Id, common_start_day, common_end_day),
    by = "Id") |> 
  filter(temp_date >= common_start_day & temp_date <= common_end_day) |>
  select(-temp_date, -common_start_day, -common_end_day) 

```

Next, a single dataframe is created with all measurements synchronized to a common datetime column. Because the dosimeters have slightly different recording times, the LightLogR data2reference function is used to align the datetime of the chest- and wrist-worn dosimeters to the head-worn measurements.

```{r}
data_hcw <- data2reference(data_h, Reference.data = data_c, Reference.column = MEDI_c) |>
  select(-position) |>
  left_join(
    data2reference(data_h, Reference.data = data_w, Reference.column = MEDI_w) |>
      select(Id, Datetime, MEDI_w),
    by = c("Id", "Datetime")
  )

names(data_hcw)[names(data_hcw) == "MEDI"] <- "MEDI_h"

print(data_hcw)
```

## Non-wear filter

For each measurement instance, it is indicated whether the dosimeters were worn (y) or not (n). If any device was not worn, the measurements are labeled N. This feature has not yet been implemented. *Q: Is the wear log data (e.g. available [here](https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/blob/v1.0.1/data/raw/individual/IZTECH_S001/continuous/wearlog/IZTECH_S001_wearlog_20250616.csv?) already ready for use? Or are you planning to do additional cleaning?*

## Calculate the absolute and relative differences

For each chest- and wrist-worn measurement, the absolute and relative difference to the head-worn measurement is calculated. The following exceptions are applied in the calculation: *If MEDI_h = 0 and MEDI_c ≠ 0 → reldif_c = Inf* *If MEDI_h = 0 and MEDI_w ≠ 0 → reldif_w = Inf* *If MEDI_h = 0 and MEDI_c = 0 → reldif_c = 0* *If MEDI_h = 0 and MEDI_w = 0 → reldif_w = 0*

```{r}
# Absolute differences
data_hcw <- data_hcw |>
  mutate(
    absdif_c = MEDI_c - MEDI_h,
    absdif_w = MEDI_w - MEDI_h
  )

# Relative differences
data_hcw <- data_hcw |>
  mutate(
    reldif_c = case_when(
      MEDI_h == 0 & MEDI_c == 0 ~ 0,
      MEDI_h == 0 & MEDI_c != 0 ~ Inf,
      TRUE ~ (MEDI_c - MEDI_h) / MEDI_h * 100
    ),
    reldif_w = case_when(
      MEDI_h == 0 & MEDI_w == 0 ~ 0,
      MEDI_h == 0 & MEDI_w != 0 ~ Inf,
      TRUE ~ (MEDI_w - MEDI_h) / MEDI_h * 100
    )
  )
```

## Exploratory plots

As a starting point for the data exploration, one day from a single participant is used.

```{r}
# Format the data as a long table
long_hcw <- data_hcw |>
  pivot_longer(
    cols = c(MEDI_h, MEDI_c, MEDI_w, absdif_c, absdif_w, reldif_c, reldif_w),
    names_to = c(".value", "position"),
    names_pattern = "(MEDI|absdif|reldif)_(.)"
  )

## Filter for participant 1's first day
long_hcw_id1_day <- long_hcw |>
  filter(Id == "IZTECH_S001") |>
  filter_Date(start = "2024-12-25", length = days(1), tz=tz)

long_hcw_id1_day <- long_hcw_id1_day |>
  mutate(
    position = factor(
      position,
      levels = c("h", "c", "w"),
      labels = c("Head", "Chest", "Wrist")
    )
  )
```

Plot of the measurements and differences over time.

```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| label: fig-meas_over_time
#| fig-cap: "Measurements, relative differences, and absolute differences between simultaneously recorded light levels at the head, chest, and wrist of participant IZTECH_S001."
p1 <- long_hcw_id1_day |>
  gg_day(
    y.axis = MEDI,
    aes_col = position, 
    size = 0.25,
    geom = "point",
    y.axis.label = "Measurements (lx, MEDI)",
    y.scale = "symlog"
  )

p2 <- long_hcw_id1_day |>
  gg_day(
    y.axis = MEDI,
    aes_col = position, 
    size = 0.25,
    geom = "point",
    y.axis.label = "Measurements (lx, MEDI)",
    y.scale = "identity"
  ) +
  scale_y_continuous(
    breaks = c(0, 200, 400, 600, 800, 1000, 1200)
  )


p3 <- long_hcw_id1_day |>
  filter(position %in% c("Chest", "Wrist")) |>
  gg_day(
    y.axis = absdif,
    aes_col = position,
    size = 0.25,
    geom = "point",
    y.scale = "identity",
    y.axis.label = "Absolute difference (lx, MEDI)"
  )+
  ylim(-200, 200) +
  scale_color_manual(
    values = ggsci::pal_jco()(3)[2:3] 
  )

p4 <- long_hcw_id1_day |>
  filter(position %in% c("Chest", "Wrist")) |>
  gg_day(
    y.axis = reldif,
    aes_col = position,
    size = 0.25,
    geom = "point",
    y.scale = "identity",
    y.axis.label = "Relative difference (%)"
  )+
  ylim(-110, 510) +
  scale_color_manual(
    values = ggsci::pal_jco()(3)[2:3]
  )

combined <- plot_grid(p1, p2, p3, p4, ncol = 1, align = "v")

combined
```

Plot of the distribution of measurements and differences. As the non-wear filter has not yet been implemented, a manual correction is applied.

```{r}
# Manual non-wear filter (temporary)
df_plot <- long_hcw_id1_day |>
  filter(
    position %in% c("Chest", "Wrist"),
    !is.na(absdif),
    is.finite(absdif),
    hms::as_hms(Datetime) >= hms("08:35:00"),
    hms::as_hms(Datetime) <= hms("23:55:00")
  )

# Count the number of observations with absdif higher than 300
over300_counts <- df_plot |>
  filter(absdif > 300) |>
  count(position)

# Count values below -300
under300_counts <- df_plot|>
  filter(absdif < -300) |>
  count(position)

label_df <- df_plot |>
  distinct(position) |>
  left_join(under300_counts |> rename(count = n), by = "position") |>
  mutate(count = ifelse(is.na(count), 0, count)) |>
  rename(under_count = count) |>
  left_join(over300_counts |> rename(count = n), by = "position") |>
  mutate(count = ifelse(is.na(count), 0, count)) |>
  rename(over_count = count) |>
  # Create two rows per position: negative and positive
  tidyr::pivot_longer(cols = c(under_count, over_count), names_to = "extreme", values_to = "count") |>
  mutate(
    label = ifelse(extreme == "under_count",
                   paste0("< -300: ", count),
                   paste0("> 300: ", count)),
    x = ifelse(extreme == "under_count", -300, 300),
    y = Inf
  )


ggplot(df_plot, aes(x = absdif, fill = position)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  labs(
    x = "Absolute difference (lx, MEDI)",
    y = "Count (1 observ / 10 sec)"
  )   + 
  scale_x_continuous(
    breaks = seq(-300, 300, by = 50),
    limits = c(-300, 300)
  ) +
  facet_wrap(~position, nrow = 1) +
  geom_text(
    data = label_df,
    aes(x = x, y = y, label = label),
    hjust = ifelse(label_df$extreme == "under_count", 0, 1),
    vjust = 1.5,
    inherit.aes = FALSE
  )


# Count the number of inf
inf_counts <- df_plot |>
  filter(is.infinite(reldif)) |>
  count(position)

# Count the number of observations with reldif higher than 500%
over500_counts <- df_plot |>
  filter(reldif > 500) |>
  count(position)


label_df <- df_plot |>
  distinct(position) |>
  left_join(
    inf_counts |> rename(inf_count = n), by = "position"
  ) |>
  left_join(
    over500_counts |> rename(over500_count = n), by = "position"
  ) |>
  mutate(
    label = paste0("Inf: ", inf_count, "\n>500%: ", over500_count),
    x = 500,
    y = Inf
  )
```


```{r}
#| warning: false
#| message: false
#| fig-width: 12
#| label: fig-rel_dif
#| fig-cap: "Distribution of relative differences between simultaneously recorded light levels at the head, chest, and wrist of participant IZTECH_S001."
ggplot(df_plot, aes(reldif, fill = position)) +
  geom_histogram(binwidth = 5, boundary = 0) +
  labs(
    x = "relative difference (%)",
    y = "Count (1 observ / 10 sec)"
  ) +
  scale_x_continuous(
    breaks = seq(-100, 500, by = 50),
    limits = c(-100, 500)
  ) +
  facet_wrap(~position, nrow = 1) +
  geom_text(
    data = label_df,
    aes(x = x, y = y, label = label),
    hjust = 1, vjust = 1.5,
    inherit.aes = FALSE
  )

ggsave("../output/figures/reldif.png", width = 17, height = 10, scale = 2, units = "cm")

```

## References {.unnumbered}
