---
title: "Stability of wearing-position bias with sample size"
format: html
author: Johannes Zauner
editor_options:
  chunk_output_type: inline
---

## Preface

The metric comparison established that wearing position (chest and wrist,
relative to the glasses reference) introduces an outcome-level *bias* in light
exposure metrics. This document asks a complementary, design-oriented question:

> **If and when does the estimated bias stabilise as data accumulate?**

We probe this along two independent axes with bootstrapping:

1. **Participants** — how the bias and its uncertainty behave as the number of
   participants grows (1–200, resampled with replacement).
2. **Participant-days** — how they behave as the number of days *per participant*
   grows (1–7), holding the full participant set.

At each sample size we bootstrap the *mean bias* and summarise its sampling
variability by its **standard deviation**. Bias is computed on each metric's
**model scale** (the scale on which the comparison notebook analysed it) and
expressed **in percent of the glasses reference level**, so all six metrics share
one interpretable axis. The figure tracks that SD — the uncertainty in the
estimated bias — as sample size grows. A companion table then applies a practical
**5% tolerance**, reporting the sample size at which each metric's bias SD first
falls to or below 5%, or noting that it never does.

The analysis uses one metric per metric category:

| Category (metric family/type) | Metric |
|---|---|
| Duration above threshold | Duration above 250 lx mel EDI |
| Dose | Melanopic dose |
| Mean level | Mean mel EDI |
| Spectral | MDER |
| Pulses | N pulses above 250 lx mel EDI |
| Timing | Brightest 10 h midpoint |

## Setup

Packages for data wrangling (`tidyverse`), reading the metric metadata
(`readxl`), the model-scale transform `log_zero_inflated()` (`LightLogR`),
figures and tables (`cowplot`, `gt`), the wearing-position color palette
(`ggsci`), and project paths (`here`).

```{r}
#| label: setup
#| warning: false
#| message: false
#| filename: Setup
library(tidyverse)
library(readxl)
library(LightLogR)
library(cowplot)
library(gt)
library(here)
```

The pre-processed metrics are loaded from the `.RData` file produced by the
preparation notebook. The object `metrics` is long-format with one row per
participant (`Id`), date (`Date`), wearing `position`, and `metric`.

```{r}
#| label: import-metrics
#| filename: Import metrics
load(here("data/prepared_metrics.RData"))
```

A single seed is set for the whole notebook so every bootstrap is reproducible.

```{r}
#| label: set-seed
#| filename: Set seed
set.seed(7391)
```

## Metric selection

We select one metric per category and confirm that each string is actually
present in the data.

```{r}
#| label: select-metrics
#| filename: Define and verify target metrics
target_metrics <- c(
  "duration_above_250",      # Duration above 250 lx mel EDI
  "dose",                    # Melanopic dose
  "mean_MEDI",               # Mean mel EDI
  "MDER",                    # MDER
  "n_pulses_above_250",      # N pulses above 250 lx mel EDI
  "brightest_10h_midpoint"   # Brightest 10 h midpoint
)

missing <- setdiff(target_metrics, unique(metrics$metric))
stopifnot("Some target metrics are absent from `metrics`" = length(missing) == 0)

metric_labels <- c(
  duration_above_250     = "Duration > 250 lx",
  dose                   = "Melanopic dose",
  mean_MEDI              = "Mean mel EDI",
  MDER                   = "MDER",
  n_pulses_above_250     = "N pulses > 250 lx",
  brightest_10h_midpoint = "Brightest 10h midpoint"
)
```

A shared, named palette assigns one color per wearing position, matching the
comparison notebook (`ggsci::pal_jco()`).

```{r}
#| label: position-colors
#| filename: Wearing-position palette
position_colors <- ggsci::pal_jco()(3) |> set_names(c("glasses", "chest", "wrist"))
```

## Model scale of each metric

Bias must be computed on the same scale each metric was modelled on in the
comparison notebook. That scale is derived from the metric metadata in
`data/metric_types.xlsx` using the same logic as the comparison notebook: a
log-transformed response or a `glmmTMB`/Poisson engine implies a (base-10 or
natural) log scale, a `qlogis` response implies a logit scale, and everything
else is linear. `derive_metric_scales()` (sourced below) returns, per metric, the
scale and the `transform` that brings the *stored* value onto it — `identity`
where the value is already on scale (e.g. `mean_MEDI`, stored as the mean of
`log_zero_inflated(MEDI)`).

```{r}
#| label: load-metric-info
#| filename: Load metric metadata
metric_info <- read_xlsx(here("data/metric_types.xlsx"))
```

## Defining the bias

For each metric we transform values to the model scale, align the three
positions on the same participant-day, and compute the bias of each comparison
position (chest, wrist) against the glasses reference:

$$\text{bias} = \text{value}^{\text{model}}_{\text{position}} - \text{value}^{\text{model}}_{\text{glasses}}$$

To make metrics comparable, bias is expressed **in percent of the glasses
reference level** — the model-scale bias divided by the absolute mean model-scale
glasses value, matching the comparison notebook's "outcome-level bias (%)". Note
that on a log scale this is the ratio of a log-difference to a log-level, so the
logarithm base cancels and `log_zero_inflated()` (zero-safe, base-10) can be used
throughout. The illustrative table below shows, for one metric, the per-position
bias on the model scale alongside the reference level.

```{r}
#| label: bias-table-manual
#| filename: Build the bias table (illustrative, single metric)
metrics |>
  filter(metric == "dose", position %in% c("glasses", "chest", "wrist")) |>
  mutate(value = log_zero_inflated(value)) |>
  select(metric, site, Id, Date, position, value) |>
  pivot_wider(names_from = position, values_from = value) |>
  mutate(bias_chest = chest - glasses,
         bias_wrist = wrist - glasses) |>
  head()
```

## Testing the approach

Before generalising, we validate the full pipeline on a **single metric
(`dose`)** with a **small number of bootstraps (B = 20)**. This keeps the logic
transparent and confirms the end-to-end path (model-scale bias → bootstrap →
median and 90% interval in percent) behaves sensibly before the expensive full
run.

First, the model-scale bias table for the test metric, in long form (one row per
participant-day and comparison position). `dose` is on a log10 scale, so values
are passed through `log_zero_inflated()`; the reference level `ref` is the
absolute mean model-scale glasses value, used to convert bias to percent.

```{r}
#| label: test-bias
#| filename: Test bias table (single metric)
test_ms <-
  metrics |>
  filter(metric == "dose", position %in% c("glasses", "chest", "wrist")) |>
  mutate(value = log_zero_inflated(value))

test_ref <- test_ms |> filter(position == "glasses") |> summarise(abs(mean(value))) |> pull()

test_bias <-
  test_ms |>
  select(metric, site, Id, Date, position, value) |>
  pivot_wider(names_from = position, values_from = value) |>
  pivot_longer(c(chest, wrist), names_to = "position", values_to = "value_pos") |>
  mutate(bias = value_pos - glasses) |>
  filter(!is.na(bias)) |>
  select(metric, position, site, Id, Date, bias)
```

**Participant axis (test).** For a handful of sample sizes, we collapse each
participant to their mean bias, resample participants with replacement, and take
the standard deviation (in percent of `ref`) of the resampled mean across B = 20
replicates.

```{r}
#| label: test-participants
#| filename: Test participant bootstrap
test_A <-
  test_bias |>
  group_by(metric, position) |>
  group_modify(\(d, key) {
    per_id <- d |> group_by(Id) |> summarise(mb = mean(bias), .groups = "drop") |> pull(mb)
    map_dfr(c(1, 5, 25, 100, 200), \(n) {
      est <- replicate(20, mean(sample(per_id, n, replace = TRUE)))
      tibble(n = n, sd = sd(est) * 100 / test_ref)
    })
  }) |>
  ungroup()

test_A
```

**Participant-day axis (test).** For each participant we resample `n_days` of
their days with replacement, average, then average across participants; the
standard deviation across B = 20 replicates is recorded for 1–7 days.

```{r}
#| label: test-days
#| filename: Test participant-day bootstrap
test_B <-
  test_bias |>
  group_by(metric, position) |>
  group_modify(\(d, key) {
    by_id <- split(d$bias, d$Id)
    map_dfr(1:7, \(nd) {
      est <- replicate(20, mean(map_dbl(by_id, \(v) mean(sample(v, nd, replace = TRUE)))))
      tibble(n = nd, sd = sd(est) * 100 / test_ref)
    })
  }) |>
  ungroup()
```

A quick sanity plot of the participant axis: the standard deviation of the bias
(the uncertainty in the estimate) shrinks steeply with more participants, toward
the 5% tolerance.

```{r}
#| label: fig-test
#| filename: Sanity plot of the test run
#| fig-width: 8
#| fig-height: 3
test_A |>
  ggplot(aes(n, sd)) +
  geom_hline(yintercept = 5, linetype = "dashed", colour = "grey40") +
  geom_line() +
  facet_wrap(~position) +
  labs(x = "Number of participants", y = "SD of bias (% of glasses level)") +
  theme_cowplot()
```

The approach is sound: the estimate's uncertainty shrinks with data. We now
generalise.

## Generalising into reusable functions

The test logic is repeated identically across all six metrics and both
positions, so it is factored into `scripts/bootstrap_stability.R`. Sourcing it
exposes the functions used below:

- **`derive_metric_scales()`** — maps the metric metadata to a model scale and a
  value `transform` (`log_zero_inflated` / `qlogis` / `identity`).
- **`compute_bias()`** — transforms values to the model scale, then builds the
  long bias table (comparison − glasses) per participant-day, attaching the
  per-metric percent reference `ref`.
- **`boot_participants()`** — the participant axis: collapses each participant to
  their mean bias and resamples participants with replacement, returning the SD in
  percent.
- **`boot_days()`** — the participant-day axis: draws a fixed number of
  participants (default 50) with replacement, resamples `n_days` per participant,
  and averages (same summary).
- **`bootstrap_stability()`** — runs both axes and binds them into one tidy table.
- **`crossing_table()`** — the smallest sample size at which the bias SD falls to
  or below 5%.
- **`plot_stability()` / `color_strips_grid()`** — the four-panel figure and its
  position-colored strips.

```{r}
#| label: source-functions
#| filename: Source bootstrap-stability helpers
source(here("scripts/bootstrap_stability.R"))
```

The model scale and transform for each target metric:

```{r}
#| label: metric-scales
#| filename: Derive model scales
scales <- derive_metric_scales(metric_info)
scales |> filter(metric %in% target_metrics)
```

The model-scale bias table for all six metrics and both comparison positions:

```{r}
#| label: bias-all
#| filename: Bias table for all target metrics
bias_df <- compute_bias(metrics |> filter(metric %in% target_metrics), scales)
bias_df |> count(metric, position)
```

## Full bootstrap run

We now run both axes for every metric and position with **B = 1000** replicates,
across participants 1–200 and participant-days 1–7. The participant-days axis is
based on a fixed study size, set here; unlike the participant axis (which varies
the number of people), it asks how days-per-person alone affect precision for a
study of this size.

```{r}
#| label: days-n
#| filename: Fixed study size for the participant-days axis
days_n_participants <- 50
```

The chunk is cached because the full run is the expensive step; re-renders reuse
the stored result.

```{r}
#| label: full-run
#| filename: Full bootstrap run
#| cache: true
stability <- bootstrap_stability(
  bias_df,
  n_range             = 1:200,
  day_range           = 1:7,
  B                   = 1000,
  days_n_participants = days_n_participants
)
```

## Stabilisation figure

The four panels cross the two comparison positions (rows, strip color = position)
with the two bootstrap axes (columns). Each metric is a single line showing the
**standard deviation of its bias** (in percent of the glasses level) — a direct
measure of how uncertain the estimated bias is at a given sample size. The dashed
reference marks the 5% SD tolerance. The x-axes are free: the participant column
spans 1–200, while the participant-day column shows 1–7 days. The
participant-days panels are based on `r days_n_participants` participants (noted
in the figure caption), so they isolate the effect of days-per-person at a fixed
study size. The y-axis is clipped so the very large SDs at the smallest samples
are cut off.

```{r}
#| label: fig-stability
#| filename: Four-panel stability figure
#| fig-width: 10
#| fig-height: 7
#| warning: false
fig_stability <- plot_stability(stability, position_colors, metric_labels,
                                tol = 5, ylim = c(0, 20),
                                days_n_participants = days_n_participants)
fig_stability
```

```{r}
#| label: save-figure
#| filename: Save the stability figure
ggsave(here("output/figures/Fig_bias_stability.pdf"), fig_stability,
       width = 10, height = 7, device = cairo_pdf)
ggsave(here("output/figures/Fig_bias_stability.jpeg"), fig_stability,
       width = 10, height = 7)
```

## When does the bias SD fall to 5%?

The table below reports, for each metric, wearing position, and bootstrap axis,
the smallest sample size at which the bias SD first falls to or below 5% of the
glasses level. `never` means the SD stays above 5% throughout the tested range.

```{r}
#| label: tbl-crossing
#| filename: Sample size to reach 5% SD
crossing_table(stability, tol = 5) |>
  mutate(metric = recode(metric, !!!metric_labels),
         n_within = as.character(n_within)) |>
  replace_na(list(n_within = "never")) |>
  pivot_wider(names_from = c(boot_type, position), values_from = n_within,
              names_sep = " / ") |>
  gt(rowname_col = "metric") |>
  tab_header("Sample size for bias SD ≤ 5%") |>
  tab_spanner_delim(delim = " / ") |>
  sub_missing(missing_text = "never")
```

## Interpretation

The figure isolates *precision*: the point bias of each metric is a property of
the population and barely moves with sample size, so the bias SD shows purely how
uncertain that estimate is. It falls steeply with participants and more gently
with participant-days, approaching a between-participant floor that extra days
cannot cross.

Because the SD measures spread rather than magnitude, it reaches the 5% tolerance
for every metric once enough data accumulate — even metrics with a large point
bias such as mean mel EDI, whose estimate is simply *precise* at a non-zero value.
The participant axis buys precision far faster than the participant-day axis: a
few dozen participants bring most metrics' SD below 5%, whereas extending days per
participant helps only up to the between-participant floor. The crossing table
quantifies these thresholds; note that a small SD confirms the bias is well
*estimated*, not that it is small — the magnitude must be read from the comparison
notebook.

## Session info

```{r}
#| label: session-info
sessionInfo()
```
