---
title: "Wrist metric-error summary workbook"
format: html
author: Johannes Zauner
editor_options:
  chunk_output_type: console
execute:
  warning: false
  message: false
---

## Purpose

This workbook is a stripped-down, reproducible version of the metric-based analysis used for the wrist-versus-glasses comparison. It starts from the prepared participant-day metric file (`data/prepared_metrics.RData`), fits the metric models, and builds only the summary table object `metrics_summary`. Chest comparisons, exploratory figures, the full metric table, and ancillary diagnostics are intentionally omitted.

## Methods

### Dataset and reference position

The analysis uses the MeLiDos field-study wearable light-exposure data. The study collected synchronous light measurements from multiple wearing positions. For the metric-based analysis, glasses-mounted measurements are treated as the reference proxy for near-eye exposure, and wrist-worn measurements are compared against that reference.

### Prepared metric data

The prepared data used here are created in `notebooks/JZ_metric_preparation.qmd`. In brief, the preparation workflow:

1. imports glasses, chest, and wrist light logger data from `melidosData`, retaining melanopic equivalent daylight illuminance (`MEDI`) and photopic illuminance (`LIGHT`);
2. removes the MPI site from this analysis because comparable ActLumus chest/wrist data were not collected there;
3. aligns the three wearing positions into one time series and keeps only timestamps where all positions have valid melanopic measurements;
4. merges sleep-diary windows and non-wear logs;
5. removes non-wear intervals unless they are sleep intervals;
6. removes implausible high values (`>= 100000` lx) for melanopic and photopic measurements;
7. removes hours with less than 50% data availability and days with less than 80% availability;
8. calculates daily light-exposure metrics per participant, date, and wearing position, plus participant-level rhythm metrics; and
9. stores the resulting long metric table as `data/prepared_metrics.RData`.

Although the preparation stage uses all three positions to enforce matched valid timestamps, this workbook filters the prepared metric data to the `glasses` and `wrist` positions before modelling. This keeps the analysis focused on the wrist error relative to the glasses reference.

### Metric modelling

Each metric is modelled according to `data/metric_types.xlsx`, which defines the metric family, model engine, response transformation, and random-effects structure. The full model tests whether metric values differ by wearing position while accounting for study site and repeated observations. In formula form, the model is constructed as:

```{=markdown}
`response ~ site * position + [metric-specific random effects]`
```

where `position` has `glasses` as the reference level and `wrist` as the comparison level. Gaussian, Tweedie, or Poisson families are used according to the metric specification. If a mixed model does not converge or is singular, the random-effects structure is simplified stepwise, following the original metric-comparison notebook.

### Summary statistic

For each metric, the wrist error is the estimated wrist-minus-glasses contrast from the fitted model, divided by the absolute estimated glasses reference value. The resulting relative error is summarized by metric type as:

- median absolute wrist error;
- minimum and maximum absolute wrist error;
- number of significant wrist errors (`p <= 0.05`) out of all estimable metrics; and
- count of positive and negative wrist deviations.

Two overall rows are added: one across all estimable metrics and one restricted to metrics with significant wrist errors.

## Setup

```{r}
library(tidyverse)
library(gt)
library(readxl)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(emmeans)
library(here)
library(performance)
library(LightLogR)
```

```{r}
sig.level <- 0.05
wrist_color <- ggsci::pal_jco()(3)[3]
```

## Import prepared metrics

```{r}
load(here("data/prepared_metrics.RData"))
```

## Prepare wrist-versus-glasses metric data

```{r}
metrics_nested <-
  metrics |>
  filter(position %in% c("glasses", "wrist")) |>
  mutate(
    Date = replace_na(Date, as_date("2024-09-01")),
    position = factor(position, levels = c("glasses", "wrist")),
    Id = factor(Id),
    site = factor(site),
    Date = factor(Date)
  ) |>
  nest(data = -metric)
```

```{r}
metric_info <-
  read_xlsx(here("data/metric_types.xlsx")) |>
  mutate(
    family = map2(engine, note, \(engine, note) {
      if (!is.na(note)) {
        poisson()
      } else if (engine %in% c("lm", "lmer")) {
        gaussian()
      } else if (engine == "glmmTMB") {
        tweedie(link = "log")
      } else {
        gaussian()
      }
    })
  ) |>
  select(-note)

metrics_nested <-
  metrics_nested |>
  left_join(metric_info, by = join_by(metric == name))
```

Timing metrics that cross midnight are shifted so that values after midday are interpreted as times before midnight. This avoids artificial discontinuities in linear timing models.

```{r}
metrics_nested <-
  metrics_nested |>
  mutate(
    data = map2(
      data, metric,
      \(data, name) {
        if (name %in% c("darkest_10h_midpoint", "darkest_10h_onset")) {
          data |>
            mutate(value = ifelse(value >= 12 * 60 * 60, value - 24 * 60 * 60, value))
        } else {
          data
        }
      }
    )
  )
```

## Fit models

```{r}
metrics_nested <-
  metrics_nested |>
  mutate(H1 = paste0("~ site*position", random))

Formula_cols <- names(metrics_nested) |> keep(\(x) str_detect(x, "H1|H0_"))

make_formula <- function(response, rhs) {
  if (is.na(rhs) || is.na(response)) {
    return(NULL)
  }

  as.formula(paste(response, rhs))
}

metrics_nested <-
  metrics_nested |>
  mutate(
    response = str_replace(response, "metric", "value"),
    across(all_of(Formula_cols), \(x) map2(x, response, \(rhs, resp) make_formula(resp, rhs)))
  )
```

```{r}
main_formula <- value ~ site * position + site + position + (position | Id) + (1 | Date)
red_formula <- value ~ site * position + site + position + (position | Id)
red_formula2 <- value ~ site * position + site + position + (1 | Id)

rhs_same <- function(f1, f2) {
  identical(
    attr(terms(formula(f1, rhs = 1)), "term.labels"),
    attr(terms(formula(f2, rhs = 1)), "term.labels")
  )
}

fit_model <- function(data, formula, engine, family) {
  if (is.null(formula)) {
    return(NULL)
  }

  model <- switch(
    engine,
    "lm" = lm(formula, data = data, contrasts = list(site = "contr.sum")),
    "glm" = glm(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
    "lmer" = lmer(formula, data = data, contrasts = list(site = "contr.sum")),
    "glmer" = glmer(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
    "glmmTMB" = glmmTMB(formula, data = data, family = family, contrasts = list(site = "contr.sum"))
  )

  did_converge <- performance::check_convergence(model) && !performance::check_singularity(model)

  if (!did_converge) {
    if (rhs_same(main_formula, formula)) {
      formula <- update(formula, . ~ . - (1 | Date))
    } else if (rhs_same(red_formula, formula)) {
      formula <- update(formula, . ~ . - (position | Id) + (1 | Id))
    } else if (rhs_same(red_formula2, formula)) {
      formula <- update(formula, . ~ . - site:position)
    } else {
      return(model)
    }

    model <- fit_model(data, formula, engine, family)
  }

  model
}
```

```{r}
metrics_nested <-
  metrics_nested |>
  mutate(
    H1_model = pmap(list(data, H1, engine, family), fit_model, .progress = "H1 model fitting"),
    converged = map_lgl(H1_model, performance::check_convergence),
    singular = map_lgl(H1_model, performance::check_singularity)
  )
```

```{r}
all(metrics_nested$converged)
any(metrics_nested$singular)
```

## Extract wrist errors

```{r}
metric_results <-
  metrics_nested |>
  mutate(
    main_effect = map_dbl(
      H1_model,
      \(model) model |>
        emmeans(specs = ~1, at = list(position = "glasses")) |>
        as_tibble() |>
        pull(2)
    ),
    position_effect = map(
      H1_model,
      \(model) model |>
        emmeans(specs = ~position) |>
        contrast(method = "trt.vs.ctrl") |>
        as_tibble()
    ),
    wrist_error = map_dbl(position_effect, \(x) x$estimate[x$contrast == "wrist - glasses"]),
    p_wrist = map_dbl(position_effect, \(x) x$p.value[x$contrast == "wrist - glasses"]),
    wrist_error = wrist_error / abs(main_effect)
  ) |>
  select(metric, metric_family, metric_type, wrist_error, p_wrist)
```

```{r}
format_metric <- function(x) {
  case_when(
    x %in% c("MDER", "nvRD") ~ x,
    x == "centroidLE" ~ "Centroid LE",
    x == "midpointCE" ~ "Midpoint CE",
    x == "mean_MEDI" ~ "Mean mel EDI",
    TRUE ~ x |>
      str_replace_all("_", " ") |>
      str_to_sentence() |>
      str_replace_all("(\\d+)(?!\\d|h)", "\\1 lx mel EDI")
  )
}

metric_results <-
  metric_results |>
  mutate(across(c(metric, metric_family, metric_type), format_metric))
```

## Build `metrics_summary`

```{r}
summarize_wrist_errors <- function(data, metric_type = NULL, significant_only = FALSE) {
  data <-
    data |>
    mutate(
      p_wrist = p_wrist <= sig.level,
      wrist_error = if (significant_only) ifelse(p_wrist, wrist_error, NA_real_) else wrist_error
    )

  out <-
    data |>
    summarize(
      n = n(),
      wrist_median = median(abs(wrist_error), na.rm = TRUE),
      wrist_min = min(abs(wrist_error), na.rm = TRUE),
      wrist_max = max(abs(wrist_error), na.rm = TRUE),
      wrist_positive = sum(wrist_error > 0, na.rm = TRUE),
      wrist_negative = sum(wrist_error < 0, na.rm = TRUE),
      wrist_significant = sum(p_wrist, na.rm = TRUE),
      wrist_estimable = sum(!is.na(wrist_error))
    ) |>
    mutate(
      across(c(wrist_min, wrist_max), \(x) ifelse(is.infinite(x), NA_real_, x))
    )

  if (significant_only) {
    out <-
      out |>
      mutate(
        wrist_significant = NA_integer_,
        wrist_estimable = NA_integer_
      )
  }

  if (!is.null(metric_type)) {
    out <- mutate(out, metric_type = metric_type, .before = n)
  }

  out
}

summary_by_type <-
  metric_results |>
  group_by(metric_type) |>
  group_modify(\(.x, .y) summarize_wrist_errors(.x)) |>
  ungroup()

summary_overall <- summarize_wrist_errors(metric_results, metric_type = "Overall")
summary_significant <- summarize_wrist_errors(
  metric_results,
  metric_type = "Overall (significant)",
  significant_only = TRUE
)

metrics_summary_data <-
  bind_rows(summary_by_type, summary_overall, summary_significant) |>
  mutate(
    metric_type = factor(
      metric_type,
      levels = c(
        setdiff(metric_type, c("Overall", "Overall (significant)")),
        "Overall",
        "Overall (significant)"
      )
    )
  ) |>
  arrange(metric_type)
```

```{r}
metrics_summary <-
  metrics_summary_data |>
  gt(rowname_col = "metric_type") |>
  fmt_percent(c(wrist_median, wrist_min, wrist_max), decimals = 0) |>
  cols_merge(
    columns = c(
      wrist_median,
      wrist_min,
      wrist_max,
      wrist_significant,
      wrist_estimable,
      wrist_positive,
      wrist_negative
    ),
    pattern = "<b>{1}</b> ({2} - {3})<br><b><<signif {4}/{5}, >></b>+{6}/-{7}"
  ) |>
  cols_label(wrist_median = "Wrist") |>
  cols_hide(n) |>
  sub_missing() |>
  tab_footnote("median error (min - max), significant errors / all metrics, positive / negative deviations") |>
  tab_style(cell_text(weight = "bold"), list(cells_stub(), cells_column_labels())) |>
  tab_header(
    "Summary overview of metric errors due to wrist wearing position, compared to glasses",
    subtitle = "Errors are all based on their modelling scale"
  ) |>
  tab_style(
    cell_fill(color = "lightgrey", alpha = 0.25),
    locations = list(
      cells_body(rows = metric_type == "Overall (significant)"),
      cells_stub(rows = c("Overall", "Overall (significant)"))
    )
  ) |>
  tab_style(cell_text(color = wrist_color), cells_column_labels(wrist_median)) |>
  tab_style(cell_fill(color = wrist_color, alpha = 0.1), cells_body(wrist_median)) |>
  tab_style(
    cell_fill(color = wrist_color, alpha = 0.5),
    cells_body(wrist_median, metric_type %in% c("Overall", "Overall (significant)"))
  )

metrics_summary
```

```{r}
gtsave(metrics_summary, here("output/tables/Table_metrics_summary_wrist.png"), vwidth = 1000)
```

## Session info

```{r}
sessionInfo()
```
