---
title: "Comparison of metrics"
format: html
author: Johannes Zauner
---

## Setup

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(gt)
library(see)
library(readxl)
library(cowplot)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(emmeans)
library(car)
library(performance)
library(LightLogR)
library(broom)
library(broom.mixed)
```

## Import

```{r}
load("../data/prepared_metrics.RData")
```




## Visualizing metrics

```{r}
#| filename: helpers
breaks <- c(0, 10^(0:5))
colors <- ggsci::pal_jco()(3) |> set_names(c("glasses", "chest", "wrist"))
label <- function(x) format(x, 
      scientific = FALSE, big.mark = " ")
```

## Preparing metric table

```{r}
#| filename: nest data
metrics_nested <-
  metrics |> 
  mutate(Date = replace_na(Date, as_date("2024-09-01")),
         position = factor(position, levels = c("glasses", "chest", "wrist")),
         Id = factor(Id),
         site = factor(site),
         Date = factor(Date)) |> 
  nest(data = -c(metric))
```

```{r}
#| filename: Load metric information
metric_info <- read_xlsx("../data/metric_types.xlsx")
metric_info <- 
metric_info |> 
  mutate(family = map(engine, \(x) switch(x,
                              "lm" = gaussian(),
                              "lmer" = gaussian(),
                              "glmmTMB" = tweedie(link = "log"))),
         family = map2(family, note, 
                       \(x,y) ifelse(is.na(y), return(x), return(poisson()))
                       )
         ) |> 
  select(-note)
```

```{r}
#| filename: Combine metric data and metadata
metrics_nested <- 
metrics_nested |> 
  left_join(metric_info, by = join_by(metric == name))
```

### Scaling nighttime metrics

Metrics that capture timing and might cross midnight need to be rescaled in order to work with linear models. The next plot shows the distribution of timing metrics.

```{r}
#| fig-width: 15
#| fig-height: 15
#| filename: show distribution of timing metrics
metrics_nested |> 
  filter(metric_type == "timing") |> 
  unnest(data) |> 
    group_by(Id, metric) |> 
  mutate(
    position = fct_relevel(position, c("chest", "glasses", "wrist")),
    ) |>
  # filter(metric == "bright_cluster") |>
  ggplot(aes(x=position, y = value)) +
  # geom_line(aes(group = interaction(Id, Date)), alpha = 0.1) +
  geom_jitter(aes(col = position), height = 0,             
              width = 0.1, alpha = 0.5) +
  geom_boxplot(aes(col = position, fill = position), alpha = 0.5,
               width = 0.1,
              position = position_nudge(x = +0.2)) +
  geom_violinhalf(aes(x = position, y = value,
                   col = position, fill = position), 
              position = position_nudge(x = +0.3),
                inherit.aes = FALSE, alpha = 0.5) +
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  facet_wrap(~metric, ncol = 4, scales = "free_y") +
  guides(col = "none", fill = "none") +
  labs(y = "Scaled and centered metric values",
       x = "Wearing position",
       title = "Raincloud plots of scaled and centered metric results") +
  theme_cowplot() +
  scale_y_time()

```

Affected metrics are:

- darkest 10 hours midpoint
- darkest 10 hours onset

Both show a clear split of the data around midnight. To resolve this, we will shift the datas so that data after 12:00 midday are negative values (i.e., indicate time before midnight).

```{r}
#| fig-width: 15
#| fig-height: 15
#| filename: transform nighttime timing metrics
metrics_nested <- 
metrics_nested |> 
  mutate(data = map2(data, metric, 
                     \(data, name) {
                       if(name %in% c("darkest_10h_midpoint", "darkest_10h_onset")) {
                         data |> 
                           mutate(value = 
                                    ifelse(value >= (12*60*60), 
                                           value - 24*60*60, 
                                           value)
                                  )
                       } else data
                     }))
```

## Visualize metric distribution

```{r}
#| fig-width: 15
#| fig-height: 20
#| warning: false
#| filename: Show distribution of all metrics
metrics |> 
  group_by(Id, metric) |> 
  mutate(
    position = fct_relevel(position, c("chest", "glasses", "wrist")),
    value = base::scale(value)) |>
  # filter(metric == "bright_cluster") |>
  ggplot(aes(x=position, y = value)) +
  # geom_line(aes(group = interaction(Id, Date)), alpha = 0.1) +
  geom_jitter(aes(col = position), height = 0,             
              width = 0.1, alpha = 0.5) +
  geom_boxplot(aes(col = position, fill = position), alpha = 0.5,
               width = 0.1,
              position = position_nudge(x = +0.2)) +
  geom_violinhalf(aes(x = position, y = value,
                   col = position, fill = position), 
              position = position_nudge(x = +0.3),
                inherit.aes = FALSE, alpha = 0.5) +
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  facet_wrap(~metric, ncol = 5) +
  guides(col = "none", fill = "none") +
  labs(y = "Scaled and centered metric values",
       x = "Wearing position",
       title = "Raincloud plots of scaled and centered metric results") +
  theme_cowplot()

ggsave("../output/figures/Figure_metric_distribution.pdf", width = 15, height = 20)
```

## Prepare modelling

```{r}
#| filename: Add formulas
metrics_nested <- 
  metrics_nested |> 
  mutate(H1 = paste0("~ site*position", random))
```

```{r}
#| filename: Collect models
Formulas <-  names(metrics_nested) |> subset(grepl("H1|H0_", names(metrics_nested)))
```

```{r}
#| echo: false
metrics_nested |> 
  select(-data, -random) |>
  arrange(metric_type, metric_family, metric) |> 
  mutate(family = map(family, 1),
         rowname = seq_along(1:length(metric))) |> 
  group_by(metric_type) |> 
  gt() |> 
  cols_label_with(fn = \(x) x |> str_to_title() |> str_replace_all("_", " ")) |> 
  sub_missing(missing_text = "") |> 
  sub_values(values = "NULL", replacement = "") |> 
  fmt(c(metric, pre.transformed, response), fns = \(x) x |> str_to_title() |>  str_replace_all("_", " ")) |> 
  sub_values(values = "MDER", replacement = "MDER") |> 
  tab_header("Metric model specifications",
             subtitle = "Hypothesis: wearing position affects metric result within sites") |> 
  tab_footnote("Confirmatory analysis will use no p-value correction, as a non-significant difference result is the more desired result") |> 
  tab_footnote("These are the models used for the analysis per metric. H1 is tested with car::Anova to inspect the significant portions", locations = cells_column_labels(Formulas)) |> 
  cols_width(
    Formulas ~ px(200),
    c(metric, pre.transformed, response) ~ px(125),
    everything() ~ px(80))
```

```{r}
#| filename: Formula creation
make_formula <- function(response, rhs) {
  if(is.na(rhs) | is.na(response)) return()
  as.formula(paste(response, rhs))
}

metrics_nested <-
  metrics_nested |> 
  mutate(response = str_replace(response, "metric", "value"),
         across(all_of(Formulas),
                \(x) map2(x, response, \(y,z) make_formula(z, y))
                )
         )
```

## Fit models

```{r}
#| filename: Helper functions
main_formula <- value ~ site*position + site + position + (position|Id) + (1|Date) 
red_formula <- value ~ site*position + site + position + (position|Id)
red_formula2 <- value ~ site*position + site + position + (1|Id)

rhs_same <- function(f1, f2) {
  identical(
    attr(terms(formula(f1, rhs = 1)), "term.labels"),
    attr(terms(formula(f2, rhs = 1)), "term.labels")
  )
}

fit_model <- function(data, formula, engine, family) {
  if(is.null(formula)) return()
  model <- 
  switch(engine,
         "lm" = lm(formula, data = data, contrasts = list(site = "contr.sum")),
         "glm" = glm(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
         "lmer" = lmer(formula, data = data, contrasts = list(site = "contr.sum")),
         "glmer" = glmer(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
         "glmmTMB" = glmmTMB(formula, data = data, family = family, contrasts = list(site = "contr.sum"))
         )

did_converge <- 
  performance::check_convergence(model) & !check_singularity(model)
 
  if(!did_converge) {
    if(rhs_same(main_formula, formula)) {
      formula <- formula |> update(. ~ . - (1 | Date))
    } else if (rhs_same(red_formula, formula)) {
      formula <-  formula |> update(. ~ . - (position|Id) + (1|Id))
    } else if (rhs_same(red_formula2, formula)) {
      formula <-  formula |> update(. ~ . - site:position)
    } else return(model)
      
  model <- fit_model(data, formula, engine, family)
  }
  
  model

}
```

```{r}
#| filename: Model fit
#| message: false
#| warning: false
metrics_nested <-
  metrics_nested |> 
  mutate(
    H1_model = pmap(list(data, H1, engine, family), fit_model, .progress = "H1 model fitting"),
    converged = map_lgl(H1_model, check_convergence),
    singular = map_lgl(H1_model, check_singularity),
    H1_fin = map(H1_model, formula),
    H1_no_position = H1_fin |> map(\(x) update(x, . ~ . - site:position - position - (position|Id) + (1|Id))),
    H0 = H1_fin |> map(\(x) update(x, . ~ . - site*position)),
    H1_no_date = H1_fin |> map(\(x) update(x, . ~ . - (1|Date))),
    H0_no_date = H1_no_date |> map(\(x) update(x, . ~ . - site*position)),
    H1_no_slope = H1_fin |> map(\(x) update(x, . ~ . - (position|Id) + (1|Id))),
    H0_no_slope = H1_no_slope |> map(\(x) update(x, . ~ .  - site*position)),
    H0_model = pmap(list(data, H0, engine, family), fit_model, .progress = "H0 model fitting"),
    H1_no_position_model = pmap(list(data, H1_no_position, engine, family), fit_model, .progress = "H1 no position model fitting"),
    H1_no_date_model = pmap(list(data, H1_no_date, engine, family), fit_model, .progress = "H1 no date model fitting"),
    H0_no_date_model = pmap(list(data, H0_no_date, engine, family), fit_model, .progress = "H0 no date model fitting"),
    H1_no_slope_model = pmap(list(data, H1_no_slope, engine, family), fit_model, .progress = "H1 no position slope model fitting"),
    H0_no_slope_model = pmap(list(data, H0_no_slope, engine, family), fit_model, .progress = "H0 no position slope model fitting")
    )
```

After model fitting, all models converged and aren't singular. This was achieved through a stepwise reduction of model complexity. 

```{r}
#| filename: Checks for singularity and convergence
metrics_nested |> pull(converged) |> all() #did all converge
metrics_nested |> pull(singular) |> any() #any are singular?
```

## Extract model results

```{r}
#| filename: Extract model results
#| message: false
sig.level <- 0.05
metrics_nested <-
  metrics_nested |> 
  mutate(
         pvalues = map(H1_model, \(x) Anova(x) |> tidy()),
         diagnostics = map(H1_model, check_model),
         random_effects = map(H1_model,
                              \(x) x |>
                                  tidy() |>
                                  mutate(signif = p.value <= sig.level,
                                         estimate = exp(estimate) |> round(2)) |>
                                         filter(effect == "ran_pars") |>
                                          select(term, estimate)
                                          ),
         main_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ 1, at = list(position = "glasses")) |>
                             as.tibble()),
         position_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ position) |>
                             contrast(method = "trt.vs.ctrl") |>
                             as.tibble()),
         site_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ site | position) |>
                             contrast() |>
                             as.tibble()),
           r2_1 = map2(H1_model, H0_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_date = map2(H1_no_date_model, H0_no_date_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_slope = map2(H1_no_slope_model, H0_no_slope_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_position = map2(H1_no_position_model, H0_no_slope_model,
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_residual = r2_1 |> map_dbl(\(x) 1 - x$R2_conditional),
           r2_random = r2_1 |> map_dbl(\(x) x$R2_conditional - x$R2_marginal),
           r2_fixed = r2_1 |> map_dbl(\(x) x$R2_marginal),
           r2_position = map2_dbl(r2_1, r2_no_position,
                          \(x, y) x$R2_marginal - y$R2_marginal),
           r2_slope = map2_dbl(r2_1, r2_no_slope, 
                          \(x, y) (x$R2_conditional - x$R2_marginal) - (y$R2_conditional - y$R2_marginal)),
           r2_date = map2_dbl(r2_1, r2_no_date, 
                          \(x, y) (x$R2_conditional - x$R2_marginal) - (y$R2_conditional - y$R2_marginal)),
           r2_id = r2_random - r2_slope - r2_date,
           r2_site = r2_fixed - r2_position
         )
```

## Session info

```{r}
sessionInfo()
```

