---
title: "Comparison of metrics"
format: html
author: Johannes Zauner
editor_options: 
  chunk_output_type: console
---

## Preface

This document contains the analysis steps from the pre-calculated metrics (see @fig-fc_sample), through inference and the summary of results.

{{< include _metric_flowchart.qmd >}}

## Setup

```{r}
#| warning: false
#| message: false
library(tidyverse)
library(gt)
library(see)
library(readxl)
library(gghighlight)
library(cowplot)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(melidosData)
library(emmeans)
library(here)
library(car)
library(performance)
library(LightLogR)
library(broom)
library(broom.mixed)
library(katex)
library(gtsummary)
library(stringr)
library(knitr)
library(rmarkdown)
library(fontawesome)
performance::r2_nakagawa()
```

## Import

```{r}
load(here("data/prepared_metrics.RData"))
```

## Visualizing metrics

```{r}
#| filename: helpers
breaks <- c(0, 10^(0:5))
colors <- ggsci::pal_jco()(3) |> set_names(c("glasses", "chest", "wrist"))
label <- function(x) format(x, 
      scientific = FALSE, big.mark = " ")
```

## Preparing metric table

```{r}
#| filename: nest data
metrics_nested <-
  metrics |> 
  mutate(Date = replace_na(Date, as_date("2024-09-01")),
         position = factor(position, levels = c("glasses", "chest", "wrist")),
         Id = factor(Id),
         site = factor(site),
         Date = factor(Date)) |> 
  nest(data = -c(metric))
```

```{r}
#| filename: Load metric information
metric_info <- read_xlsx(here("data/metric_types.xlsx"))
metric_info <- 
metric_info |> 
  mutate(family = map(engine, \(x) switch(x,
                              "lm" = gaussian(),
                              "lmer" = gaussian(),
                              "glmer" = poisson(),
                              "glmmTMB" = tweedie(link = "log"))),
         family = map2(family, note, 
                       \(x,y) ifelse(is.na(y), return(x), return(poisson()))
                       )
         ) |> 
  select(-note)
```

```{r}
#| filename: Combine metric data and metadata
metrics_nested <- 
metrics_nested |> 
  left_join(metric_info, by = join_by(metric == name))
```

### Scaling nighttime metrics

Metrics that capture timing and might cross midnight need to be rescaled in order to work with linear models. The next plot shows the distribution of timing metrics.

```{r}
#| fig-width: 15
#| fig-height: 15
#| filename: show distribution of timing metrics
#| warning: false
metrics_nested |> 
  filter(metric_type == "timing") |> 
  unnest(data) |> 
    group_by(Id, metric) |> 
  mutate(
    position = fct_relevel(position, c("chest", "glasses", "wrist")),
    ) |>
  ggplot(aes(x=position, y = value)) +
  geom_jitter(aes(col = position), height = 0,             
              width = 0.1, alpha = 0.5) +
  geom_boxplot(aes(col = position, fill = position), alpha = 0.5,
               width = 0.1,
              position = position_nudge(x = +0.2)) +
  geom_violinhalf(aes(x = position, y = value,
                   col = position, fill = position), 
              position = position_nudge(x = +0.3),
                inherit.aes = FALSE, alpha = 0.5) +
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  facet_wrap(~metric, ncol = 4, scales = "free_y") +
  guides(col = "none", fill = "none") +
  labs(y = "Scaled and centered metric values",
       x = "Wearing position",
       title = "Raincloud plots of scaled and centered metric results") +
  theme_cowplot() +
  scale_y_time()

```

Affected metrics are:

- darkest 10 hours midpoint
- darkest 10 hours onset

Both show a clear split of the data around midnight. To resolve this, we will shift the datas so that data after 12:00 midday are negative values (i.e., indicate time before midnight).

```{r}
#| fig-width: 15
#| fig-height: 15
#| filename: Transform nighttime timing metrics
metrics_nested <- 
metrics_nested |> 
  mutate(data = map2(data, metric, 
                     \(data, name) {
                       if(name %in% c("darkest_10h_midpoint", "darkest_10h_onset")) {
                         data |> 
                           mutate(value = 
                                    ifelse(value >= (12*60*60), 
                                           value - 24*60*60, 
                                           value)
                                  )
                       } else data
                     }))
```

## Visualize metric distribution

```{r}
#| fig-width: 15
#| fig-height: 20
#| warning: false
#| filename: Show distribution of all metrics
metrics |> 
  group_by(Id, metric) |> 
  mutate(
    position = fct_relevel(position, c("chest", "glasses", "wrist")),
    value = base::scale(value)) |>
  ggplot(aes(x=position, y = value)) +
  geom_line(aes(group = interaction(Id, Date)), alpha = 0.05) +
  geom_jitter(aes(col = position), height = 0,             
              width = 0.1, alpha = 0.5) +
  geom_boxplot(aes(col = position, fill = position), alpha = 0.5,
               width = 0.1,
              position = position_nudge(x = +0.2)) +
  geom_violinhalf(aes(x = position, y = value,
                   col = position, fill = position), 
              position = position_nudge(x = +0.3),
                inherit.aes = FALSE, alpha = 0.5) +
  scale_color_manual(values = colors) + 
  scale_fill_manual(values = colors) + 
  facet_wrap(~metric, ncol = 5) +
  guides(col = "none", fill = "none") +
  labs(y = "Scaled and centered metric values",
       x = "Wearing position",
       title = "Raincloud plots of scaled and centered metric results") +
  theme_cowplot()

ggsave(here("output/figures/Figure_metric_distribution.pdf"), width = 15, height = 20, device = cairo_pdf)
```

## Prepare modelling

```{r}
#| filename: Add formulas
metrics_nested <- 
  metrics_nested |> 
  mutate(H1 = paste0("~ site*position", random))
```

```{r}
#| filename: Collect models
Formulas <-  names(metrics_nested) |> subset(grepl("H1|H0_", names(metrics_nested)))
```

```{r}
#| echo: false
metrics_nested |> 
  select(-data, -random) |>
  arrange(metric_type, metric_family, metric) |> 
  mutate(family = map(family, 1),
         rowname = seq_along(1:length(metric))) |> 
  group_by(metric_type) |> 
  gt() |> 
  cols_label_with(fn = \(x) x |> str_to_title() |> str_replace_all("_", " ")) |> 
  sub_missing(missing_text = "") |> 
  sub_values(values = "NULL", replacement = "") |> 
  fmt(c(metric, pre.transformed, response), fns = \(x) x |> str_to_title() |>  str_replace_all("_", " ")) |> 
  sub_values(values = "MDER", replacement = "MDER") |> 
  tab_header("Metric model specifications",
             subtitle = "Hypothesis: wearing position affects metric result within sites") |> 
  tab_footnote("Confirmatory analysis will use no p-value correction, as a non-significant difference result is the more desired result") |> 
  tab_footnote("These are the models used for the analysis per metric. H1 is tested with car::Anova to inspect the significant portions", locations = cells_column_labels(Formulas)) |> 
  cols_width(
    Formulas ~ px(200),
    c(metric, pre.transformed, response) ~ px(125),
    everything() ~ px(80))
```

```{r}
#| filename: Formula creation
make_formula <- function(response, rhs) {
  if(is.na(rhs) | is.na(response)) return()
  as.formula(paste(response, rhs))
}

metrics_nested <-
  metrics_nested |> 
  mutate(response = str_replace(response, "metric", "value"),
         across(all_of(Formulas),
                \(x) map2(x, response, \(y,z) make_formula(z, y))
                )
         )
```

## Fit models

```{r}
#| filename: Helper functions
main_formula <- value ~ site*position + site + position + (position|Id) + (1|Date) 
red_formula <- value ~ site*position + site + position + (position|Id)
red_formula2 <- value ~ site*position + site + position + (1|Id)

rhs_same <- function(f1, f2) {
  identical(
    attr(terms(formula(f1, rhs = 1)), "term.labels"),
    attr(terms(formula(f2, rhs = 1)), "term.labels")
  )
}

fit_model <- function(data, formula, engine, family) {
  if(is.null(formula)) return()
  model <- 
  switch(engine,
         "lm" = lm(formula, data = data, contrasts = list(site = "contr.sum")),
         "glm" = glm(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
         "lmer" = lmer(formula, data = data, contrasts = list(site = "contr.sum")),
         "glmer" = glmer(formula, data = data, family = family, contrasts = list(site = "contr.sum")),
         "glmmTMB" = glmmTMB(formula, data = data, family = family, contrasts = list(site = "contr.sum"))
         )

did_converge <- 
  performance::check_convergence(model) & !check_singularity(model)
 
  if(!did_converge) {
    if(rhs_same(main_formula, formula)) {
      formula <- formula |> update(. ~ . - (1 | Date))
    } else if (rhs_same(red_formula, formula)) {
      formula <-  formula |> update(. ~ . - (position|Id) + (1|Id))
    } else if (rhs_same(red_formula2, formula)) {
      formula <-  formula |> update(. ~ . - site:position)
    } else return(model)
      
  model <- fit_model(data, formula, engine, family)
  }
  
  model

}
```

```{r}
#| filename: Model fit
#| message: false
#| warning: false
metrics_nested <-
  metrics_nested |> 
  mutate(
    H1_model = pmap(list(data, H1, engine, family), fit_model, .progress = "H1 model fitting"),
    converged = map_lgl(H1_model, check_convergence),
    singular = map_lgl(H1_model, check_singularity),
    #For R^2:
    H1_fin = map(H1_model, formula),
    H1_no_position = H1_fin |> map(\(x) update(x, . ~ . - site:position - position - (position|Id) + (1|Id))),
    H0 = H1_fin |> map(\(x) update(x, . ~ . - site*position)),
    H1_no_date = H1_fin |> map(\(x) update(x, . ~ . - (1|Date))),
    H0_no_date = H1_no_date |> map(\(x) update(x, . ~ . - site*position)),
    H1_no_slope = H1_fin |> map(\(x) update(x, . ~ . - (position|Id) + (1|Id))),
    H0_no_slope = H1_no_slope |> map(\(x) update(x, . ~ .  - site*position)),
    H0_model = pmap(list(data, H0, engine, family), fit_model, .progress = "H0 model fitting"),
    H1_no_position_model = pmap(list(data, H1_no_position, engine, family), fit_model, .progress = "H1 no position model fitting"),
    H1_no_date_model = pmap(list(data, H1_no_date, engine, family), fit_model, .progress = "H1 no date model fitting"),
    H0_no_date_model = pmap(list(data, H0_no_date, engine, family), fit_model, .progress = "H0 no date model fitting"),
    H1_no_slope_model = pmap(list(data, H1_no_slope, engine, family), fit_model, .progress = "H1 no position slope model fitting"),
    H0_no_slope_model = pmap(list(data, H0_no_slope, engine, family), fit_model, .progress = "H0 no position slope model fitting")
    )
```

After model fitting, all models converged and aren't singular. This was achieved through a stepwise reduction of model complexity. 

```{r}
#| filename: Checks for singularity and convergence
metrics_nested |> pull(converged) |> all() #did all converge
metrics_nested |> pull(singular) |> any() #any are singular?
```

## Extract model results

```{r}
#| filename: Extract model results
#| message: false
sig.level <- 0.05
metrics_nested <-
  metrics_nested |> 
  mutate(
         pvalues = map(H1_model, \(x) Anova(x) |> tidy()),
         diagnostics = map(H1_model, check_model),
         random_effects = map(H1_model,
                              \(x) x |>
                                  tidy() |>
                                         filter(effect == "ran_pars") |>
                                          select(term, group, estimate)
                                          ),
         main_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ 1, at = list(position = "glasses")) |>
                             as.tibble()),
         position_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ position) |>
                             contrast(method = "trt.vs.ctrl") |>
                             as.tibble()),
         site_effect = map(H1_model,
                           \(x) x |>
                             emmeans(specs =  ~ site | position) |>
                             contrast() |>
                             as.tibble()),
           r2_1 = map2(H1_model, H0_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_date = map2(H1_no_date_model, H0_no_date_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_slope = map2(H1_no_slope_model, H0_no_slope_model, 
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_no_position = map2(H1_no_position_model, H0_no_slope_model,
                     \(x,y) r2_nakagawa(x, null_model = y)),
           r2_residual = r2_1 |> map_dbl(\(x) 1 - x$R2_conditional),
           r2_random = r2_1 |> map_dbl(\(x) x$R2_conditional - x$R2_marginal),
           r2_fixed = r2_1 |> map_dbl(\(x) x$R2_marginal),
           r2_position = map2_dbl(r2_1, r2_no_position,
                          \(x, y) x$R2_marginal - y$R2_marginal),
           r2_slope = map2_dbl(r2_1, r2_no_slope, 
                          \(x, y) (x$R2_conditional - x$R2_marginal) - 
                            (y$R2_conditional - y$R2_marginal)),
           r2_date = map2_dbl(r2_1, r2_no_date, 
                          \(x, y) (x$R2_conditional - x$R2_marginal) - 
                            (y$R2_conditional - y$R2_marginal)),
           r2_id = r2_random - r2_slope - r2_date,
           r2_site = r2_fixed - r2_position
         )
```

## Result table

```{r}
#| filename: Collect result table info
metric_table <-
  metrics_nested |> 
  mutate(
    scaling = 
      case_when(pre.transformed == "log_zero_inflated" ~ "log10",
                engine == "glmmTMB" ~ "log",
                response == "log_zero_inflated(value)" ~ "log10",
                response == "qlogis(value)" ~ "logit",
                map_chr(family, 1) == "poisson" ~ "log",
                .default = "linear"
                  ),
    pvalues = pvalues |> map(
      \(x) x |> 
        select(term, p.value) |> 
        pivot_wider(names_from = term, values_from = p.value, names_prefix = "p_")
    ),
    random_effects = random_effects |> map(
      \(x) x |> 
            mutate(group = replace_when(group, 
                                        term == "sd__positionchest" ~ "chest",
                                        term == "sd__positionwrist" ~ "wrist")) |> 
            filter_out(str_detect(term, "cor__")) |> 
            select(-term) |> 
            pivot_wider(names_from = group, values_from = estimate, names_prefix = "SD_")
    ),
    main_effect = main_effect |> map_dbl(2),
    position_effect = position_effect |> map(
      \(x) x |> 
            select(contrast, estimate, p = p.value) |> 
            pivot_wider(names_from = contrast, values_from = c(estimate, p))
              ),
    site_effect = site_effect |> map(
     \(x) x |> 
          mutate(contrast = str_remove(contrast, " effect")) |> 
          unite(name, contrast, position) |> 
          select(name, estimate, p = p.value) |> 
          pivot_wider(names_from = name, values_from = c(estimate, p)) 
    )
  ) |> 
  select(metric, metric_family, metric_type, scaling, pvalues,
         random_effects, main_effect, position_effect, site_effect,
         r2_position, r2_site, r2_id, r2_date, r2_slope, r2_residual) |> 
  unnest(c(pvalues, random_effects, position_effect, site_effect))
```


```{r}
#| filename: unformatted table for reference
metric_table |> group_by(metric_type) |> gt() |> opt_stylize() |> fmt_number()
```

```{r}
#| filename: Table construction

format_metric <- function(x) {
  case_when(
    x %in% c("MDER", "nvRD")  ~ x,
    x == "centroidLE"          ~ "Centroid LE",
    x == "midpointCE"          ~ "Midpoint CE",
    x == "mean_MEDI"           ~ "Mean mel EDI",
    TRUE ~ x |>
      str_replace_all("_", " ") |>
      str_to_sentence() |>
      str_replace_all("(\\d+)(?!\\d|h)", "\\1 lx mel EDI")
  )
} 

sites <- c("BAUA", "FUSPCEU", "IZTECH", "KNUST", "RISE", "THUAS", "TUM", "UCR")
site_colors <- melidosData::melidos_colors[sites]

# Returns a string of colored up/down SVG arrows for each significant site
make_arrows_html <- function(estimates, pvals, colors, sig_level) {
  sig <- !is.na(pvals) & pvals < sig_level
  if (!any(sig)) return("")
  dir  <- ifelse(estimates[sig] > 0, "arrow-up", "arrow-down")
  cols <- colors[sig]
  paste(
    mapply(\(d, c) as.character(fa(d, fill = c, height = "1em")), dir, cols),
    collapse = ""
  )
}

df <- metric_table |>
  mutate(across(c(metric, metric_family, metric_type), format_metric)) |>
  arrange(metric_type, metric, metric_family) |>
  rowwise() |>
  mutate(
    arrow_main = make_arrows_html(
      c_across(all_of(paste0("estimate_", sites, "_glasses"))),
      c_across(all_of(paste0("p_",        sites, "_glasses"))),
      site_colors, sig.level
    ),
    arrow_chest = make_arrows_html(
      c_across(all_of(paste0("estimate_", sites, "_chest"))),
      c_across(all_of(paste0("p_",        sites, "_chest"))),
      site_colors, sig.level
    ),
    arrow_wrist = make_arrows_html(
      c_across(all_of(paste0("estimate_", sites, "_wrist"))),
      c_across(all_of(paste0("p_",        sites, "_wrist"))),
      site_colors, sig.level
    ),
    across(c(`estimate_chest - glasses`, `estimate_wrist - glasses`,
             starts_with("SD_")),
           \(x) x/abs(main_effect)),
    across(starts_with("r2_"), \(x) replace_values(x, 0 ~ NA))
  ) |>
  ungroup()

table <-
df |>
  group_by(metric_type) |>
  gt(rowname_col = c("metric_family", "metric")) |>
  cols_width(
    # metric_type                ~ px(80),
    metric_family              ~ px(85),
    metric                     ~ px(340),
    main_effect                ~ px(120),
    arrow_main                 ~ px(60),
    `estimate_chest - glasses` ~ px(55),
    arrow_chest                ~ px(60),
    `estimate_wrist - glasses` ~ px(55),
    SD_Residual ~ px(70),
    arrow_wrist                ~ px(60),
    scaling ~ px(55),
    starts_with("p_") ~ px(65),
    starts_with("r2_") ~ px(65),
    everything()               ~ px(60)
  ) |>
  cols_move(main_effect,                after = scaling) |>
  cols_move(arrow_main,                 after = main_effect) |>
  cols_move(`estimate_chest - glasses`, after = arrow_main) |>
  cols_move(arrow_chest,                after = `estimate_chest - glasses`) |>
  cols_move(`estimate_wrist - glasses`, after = arrow_chest) |>
  cols_move(arrow_wrist,                after = `estimate_wrist - glasses`) |>
  cols_move(c(p_position, r2_position, SD_chest, SD_wrist, r2_slope, SD_Id, r2_id, SD_Date, r2_date, SD_Residual, r2_residual),
            after = arrow_wrist) |>
  fmt_scientific(main_effect) |>
  fmt_percent(
    columns  = c(`estimate_chest - glasses`, `estimate_wrist - glasses`), 
    decimals = 0, force_sign = TRUE
  ) |>
  fmt_percent(
    c(starts_with("r2_")), 
    decimals = 0,
  ) |>
  fmt_percent(
    c(starts_with("SD_")), 
    decimals = 0, pattern = "±{x}"
  ) |>
  fmt(starts_with("p_"), 
      fns = style_pvalue
        ) |> 
  cols_label(
    main_effect                = "Reference",
    arrow_main                 = "",
    p_position = md("$p_{Pos.}$"),
    p_site = md("$p_{Site}$"),
    `p_site:position` = md("$p_{Inter.}$"),
    r2_position = md("$R^2_{Pos.}$"),
    r2_slope = md("$R^2_{Slope}$"),
    r2_id= md("$R^2_{Id}$"),
    r2_date= md("$R^2_{Date}$"),
    r2_residual= md("$R^2_{Res.}$"),
    r2_site= md("$R^2_{Site}$"),
    scaling = "Scale",
    `estimate_chest - glasses` = "Chest",
    arrow_chest                = "",
    `estimate_wrist - glasses` = "Wrist",
    arrow_wrist                = ""
  ) |>
  cols_label_with(starts_with("SD_"), 
                  fn = \(x) x |> str_remove("SD_") |> str_to_title()) |> 
  # Tint significant cells: red if negative, blue if positive
  tab_style(
    style     = cell_fill(color = "#FFDDDD60"),
    locations = list(
      cells_body(
        columns = `estimate_chest - glasses`,
        rows    = `estimate_chest - glasses` < 0 & `p_chest - glasses` < sig.level
      ),
      cells_body(
        columns = `estimate_wrist - glasses`,
        rows    = `estimate_wrist - glasses` < 0 & `p_wrist - glasses` < sig.level
      )
    )
  ) |>
  tab_style(
    style     = cell_fill(color = "#D0EEFF60"),
    locations = list(
      cells_body(
        columns = `estimate_chest - glasses`,
        rows    = `estimate_chest - glasses` > 0 & `p_chest - glasses` < sig.level
      ),
      cells_body(
        columns = `estimate_wrist - glasses`,
        rows    = `estimate_wrist - glasses` > 0 & `p_wrist - glasses` < sig.level
      )
    )
  ) |>
  fmt(
    columns = c(arrow_main, arrow_chest, arrow_wrist),
    fns = function(x) lapply(x, gt::html)
  ) |>
  cols_hide(c(matches("(BAUA|FUSPCEU|IZTECH|KNUST|RISE|THUAS|TUM|UCR)"),
              `p_chest - glasses`, `p_wrist - glasses`)) |> 
  sub_missing() |> 
  tab_spanner(columns = c(`estimate_chest - glasses`, arrow_chest,
                          `estimate_wrist - glasses`,
                          arrow_wrist,  p_position, r2_position),
              "Wearing position effect (estimated error)") |> 
  tab_spanner(columns = c(SD_chest, SD_wrist, r2_slope, SD_Id, r2_id, SD_Date, r2_date, SD_Residual, r2_residual), 
              "Standard deviation (random effect)") |> 
  tab_spanner(columns = contains("_site"), "Site effect") |> 
    # Bold if position effect is significant
  tab_style(
    style     = cell_text(weight = "bold"),
    locations = list(
      cells_body(
      columns = `estimate_wrist - glasses`,
      rows    = `p_wrist - glasses` < sig.level),
      cells_body(
      columns = `estimate_chest - glasses`,
      rows    = `p_chest - glasses` < sig.level),
      cells_body(
      columns = p_position,
      rows    = p_position < sig.level),
      cells_body(
      columns = p_site,
      rows    = p_site < sig.level),
      cells_body(
      columns = `p_site:position`,
      rows    = `p_site:position` < sig.level),
      cells_column_labels(),
      cells_column_spanners(),
      cells_stub(columns =  "metric_family"), cells_group()
    )
  ) |> 
  tab_footnote(
    "Scale on which the reference and subsequent deviations from it (wearing position effect, site, random effects) are analyzed", 
    locations = cells_column_labels(scaling)
  ) |> 
  tab_footnote(
    md("Partial $R^2$ values of the effect compated to a model without the effect. All values within a row does not sum to 1, as they are from different models and variables are partially correlated."), 
    locations = cells_column_labels(contains("r2_"))
  )  |> tab_footnote(
    "No overall p-value correction has been performed, so as not to potentially underestimate individual deviations.", 
    locations = cells_column_labels(contains("p_"))
  )  |> tab_footnote(
    "Missing entries refer to models that had to be reduced in complexity to avoid convergence/singularity issues.", 
    locations = cells_column_spanners(2)
  ) |> 
  tab_header("Summary of wearing position errors") |> 
  tab_style(
    cell_text(color = colors[1]),
    cells_column_labels(main_effect)
  ) |> 
  tab_style(
    cell_text(color = colors[2]),
    cells_column_labels(`estimate_chest - glasses`)
  ) |> 
  tab_style(
    cell_text(color = colors[3]),
    cells_column_labels(`estimate_wrist - glasses`)
  )
```

```{r}
#| filename: Plot and save the table
table
gtsave(table, here("output/tables/Table_metrics.png"), vwidth = 2000)
```

```{r}
#| filename: Make a reduced dataframe for summaries
reduced_df <- 
df |> 
  select(metric, metric_type, metric_family, scaling, 
         p_chest = `p_chest - glasses`, p_wrist = `p_wrist - glasses`, error_chest = `estimate_chest - glasses`,
         error_wrist = `estimate_wrist - glasses`) 
```

```{r}
#| filename: Summaries by metric type
df_summary <- 
reduced_df |> 
  group_by(metric_type) |> 
  mutate(across(starts_with("p_"), \(x) x <= 0.05)) |> 
  summarize(
    n = n(),
    across(c(error_chest, error_wrist),
           list(
                median = ~ median(abs(.x), na.rm = TRUE),
                min = ~ min(abs(.x), na.rm = TRUE) |> replace_values(Inf ~ NA),
                max = ~ max(abs(.x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
                sign = ~ (.x > 0) |> sum(na.rm = TRUE),
                sign_n = ~ (.x < 0) |> sum(na.rm = TRUE),
                na = ~ is.na(.x) |> sum()
                )
           ),
    across(c(p_chest, p_wrist), sum),
    across(
      ends_with("_na"),
      \(x) n - x
    )
  )
```

```{r}
#| filename: Summary overall
df_summary2 <- 
reduced_df |> 
  mutate(across(starts_with("p_"), \(x) x <= 0.05),
         error_chest = ifelse(p_chest, error_chest, NA),
         error_wrist = ifelse(p_wrist, error_wrist, NA),
         )|> 
  summarize(
    metric_type = "Overall (significant)",
    n = n(),
    across(c(error_chest, error_wrist),
           list(
                median = ~ median(abs(.x), na.rm = TRUE),
                min = ~ min(abs(.x), na.rm = TRUE) |> replace_values(Inf ~ NA),
                max = ~ max(abs(.x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
                sign = ~ (.x > 0) |> sum(na.rm = TRUE),
                sign_n = ~ (.x < 0) |> sum(na.rm = TRUE),
                na = ~ NA
                )
           ),
    across(c(p_chest, p_wrist), sum),
    across(
      ends_with("_na"),
      \(x) n - x
    )
  )

```

```{r}
#| filename: Summary overall 2
df_summary3 <- 
reduced_df |> 
  mutate(across(starts_with("p_"), \(x) x <= 0.05),
         )|> 
  summarize(
    metric_type = "Overall",
    n = n(),
    across(c(error_chest, error_wrist),
           list(
                median = ~ median(abs(.x), na.rm = TRUE),
                min = ~ min(abs(.x), na.rm = TRUE) |> replace_values(Inf ~ NA),
                max = ~ max(abs(.x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
                sign = ~ (.x > 0) |> sum(na.rm = TRUE),
                sign_n = ~ (.x < 0) |> sum(na.rm = TRUE),
                na = ~ is.na(.x) |> sum()
                )
           ),
    across(c(p_chest, p_wrist), sum),
    across(
      ends_with("_na"),
      \(x) n - x
    )
  )

df_summary <- rbind(df_summary, df_summary3, df_summary2)
```

```{r}
#| filename: Summary-table by metric type
table_summary <- 
df_summary |> 
  gt(rowname_col = "metric_type") |> 
  fmt_percent(contains(c("_min", "_max", "_median")), decimals = 0) |> 
  cols_merge(contains("chest"), pattern = "<b>{1}</b> ({2} - {3})<br><b><<signif {7}/{6}, >></b>+{4}/-{5}") |> 
  cols_merge(contains("wrist"), pattern = "<b>{1}</b> ({2} - {3})<br><b><<signif {7}/{6}, >></b>+{4}/-{5}") |> 
  cols_label(error_chest_median = "Chest",
             error_wrist_median = "Wrist") |> 
    tab_footnote("median error (min - max), significant errors / all metrics, positive / negative deviations") |> 
  cols_hide(n) |> 
  tab_style(cell_text(weight = "bold"), list(cells_stub(), cells_column_labels())) |> 
  tab_header("Summary overview of metric errors due to wearing position, compared to glasses",
             subtitle = "Errors are all based on their modelling scale") |> 
  tab_style(cell_fill(color = "lightgrey", alpha = 0.25), 
            locations = list(
              cells_body(rows = metric_type == "Overall (significant)"),
              cells_stub(rows = c("Overall", "Overall (significant)"))
            )
            ) |> 
  tab_style(
    cell_text(color = colors[2]),
    cells_column_labels(error_chest_median)
  ) |> 
  tab_style(
    cell_text(color = colors[3]),
    cells_column_labels(error_wrist_median)
  ) |> 
  tab_style(
    cell_fill(color = colors[2], alpha = 0.1),
    cells_body(error_chest_median)
  ) |> 
  tab_style(
    cell_fill(color = colors[2], alpha = 0.5),
    cells_body(error_chest_median, metric_type %in% c("Overall", "Overall (significant)"))
  ) |> 
    tab_style(
    cell_fill(color = colors[3], alpha = 0.1),
    cells_body(error_wrist_median)
  ) |> 
  tab_style(
    cell_fill(color = colors[3], alpha = 0.5),
    cells_body(error_wrist_median, metric_type %in% c("Overall", "Overall (significant)"))
  )
  
```

```{r}
#| filename: Plot and save the summary table
table_summary
gtsave(table_summary, here("output/tables/Table_metrics_summary.png"), vwidth = 2000)
```

## Visualize results

```{r}
#| filename: Visualize absolute metric errors
reduced_df |> 
  select(metric, p_chest:error_wrist) |> 
  pivot_longer(-metric, names_sep = "_", names_to = c("type", "position")) |> 
  pivot_wider(names_from = type, values_from = value) |> 
  mutate(p = p< 0.05) |> 
  ggplot(aes(x = abs(error))) +
  geom_histogram(binwidth = 0.05, aes(fill = position)) +
  facet_wrap(~position) +
  scale_fill_manual(values = colors) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(0, 1, by = 0.2)) +
  theme_cowplot() +
  coord_cartesian(xlim = c(0,1)) +
  # guides(fill = "none") +
  labs(caption = "Binwidth: 5%", 
       title = "Histograms of (absolute) metric errors due to wearing position",
       subtitle = "Significant errors in the histogram are highlighted in the position color",
       y = "Count (n)", x = "absolute Error (%)") + 
  theme(panel.spacing = unit(2, "lines")) +
  gghighlight(p, use_group_by = FALSE,
              label_key = metric, calculate_per_facet = TRUE)

ggsave(here("output/figures/Fig_metric_abs_error.pdf"), device = cairo_pdf)
```

```{r}
#| filename: Visualize metric errors
reduced_df |> 
  select(metric, p_chest:error_wrist) |> 
  pivot_longer(-metric, names_sep = "_", names_to = c("type", "position")) |> 
  pivot_wider(names_from = type, values_from = value) |> 
  mutate(p = p< 0.05) |> 
  ggplot(aes(x = (error))) +
  geom_histogram(binwidth = 0.05, aes(fill = position)) +
  facet_wrap(~position) +
  scale_fill_manual(values = colors) +
  scale_x_continuous(labels = scales::label_percent(accuracy = 1),
                     breaks = seq(-1, 1, by = 0.2)) +
  theme_cowplot() +
  coord_cartesian(xlim = c(-0.8,0.5)) +
  # guides(fill = "none") +
  labs(caption = "Binwidth: 5%", 
       title = "Histograms of metric errors due to wearing position",
       subtitle = "Significant errors in the histogram are highlighted in the position color",
       y = "Count (n)", x = "Error (%)") + 
  theme(panel.spacing = unit(2, "lines")) +
  gghighlight(p, use_group_by = FALSE,
              label_key = metric, calculate_per_facet = TRUE)

ggsave(here("output/figures/Fig_metric_error.pdf"), device = cairo_pdf)
```


## Session info

```{r}
sessionInfo()
```

