---
title: "Preparation of light exposure data for contextual analysis of wear position errors"
format:
  html:
    toc: true
    embed-resources: true
authors: "Sietse de Vries"
number-sections: true
bibliography: ../references.bib
---

## Summary

This document describes the preparation of eye-level and body-worn dosimeter measurements for a statistical analysis which investigates whether contextual factors significantly affect the differences between light measurements at eye-level and on the body (chest and wrist), referred to as the wear position error.

The dosimeter measurements were collected at multiple sites as part of the MeLiDos project, following the protocol of Guidolin et al. [@guidolin2024]. The dataset of each site is available on [GitHub](https://github.com/MeLiDosProject):

-   BAUA, Dortmund, Germany [@broszio2025]

-   FUSPCEU, Madrid, Spain [@baezamoyano2025]

-   IZTECH, Izmir, Turkey [@didikoglu2025]

-   KNUST, Kumasi, Ghana [@akuffo2025]

-   MPI, Tübingen, Germany [@guidolin2025][^1]

-   RISE, Borås, Sweden [@nilssontengelin2026]

-   THUAS, the Hague, the Netherlands [@aerts2025a]

-   TUM, Munich, Germany [@hilden2025a]

-   UCR, San Pedro, San José, Costa Rica [@sancho-salas2026a]

[^1]: Data from this site were not included in the analysis as no chest-worn dosimeter measurements were collected and the wrist-worn measurements were made using a different type of dosimeter.

Each dataset contains dosimeter measurements from three body positions (eye level, chest, and wrist) alongside contextual data. These datasets were previously checked for gaps and irregularities, trimmed to only include measurements during the trial period, and adjusted to account for (if any) daylight saving time transitions.

All datasets were imported in R using the `MelidosData` library. Measurements at the three body positions were temporally aligned using LightLogR's `data2reference()` function [@zauner2025], and only observations (time points) with simultaneous data from all three dosimeters were retained. Measurements during sleep (derived from `Sleepdiaries.RData` ("sleepprep" and "out_ofbed")) or non-wear (derived from `Wearlog.RData` ("Start" and "end")) were removed. Additionally, measurements above 100,000 lx melanopic EDI were removed, as these are outside of the dosimeter's operating range.

Wear position errors were calculated as log-transformed differences (`logdif_x = log10(MEDI_x + 0.1) - log10(MEDI_h + 0.1)`), where x refers to either the chest (c) or wrist (w) dosimeter. The offset of 0.1 is to avoid taking the logarithm of zero. As this offset lies below the dosimeter's operating range, it has no meaningful impact on the results. Since the contextual data were gathered at an hourly resolution, the wear position errors were aggregated to hourly means.

Three participants had one of the dosimeters recording at a non-standard sampling interval: FUSPCEU_S014 (head dosimeter at a 60-second interval), IZTECH_S016 (wrist dosimeter at a 120-second interval), and RISE_S011 (head dosimeter at a 1-second interval). These were still included, as enough data remained for calculating hourly means. Days with flat (zero) signals across all three dosimeters were removed for participants FUSPCEU_S010 (03-11-2024), FUSPCEU_S017 (09-12-2024), IZTECH_S009 (18-03-2025 and 19-03-2025), and THUAS_S007 (11-04-2025).

```{r}
#| label: load-packages
#| message: false
#| warning: false


library(tidyverse)
library(rlang)
library(LightLogR)  # used for temporal alignment and photoperiod identification
library(melidosData)# used for importing the datasets
library(here)       # used to export the formatted dataset
library(scales)     # used to make plots
library(patchwork)  # used to combine plots
library(gt)         # used to make tables
library(moments)    # used fur kurtosis calculation
library(cowplot)    # used for plotting
library(ggh4x)      # used to color plots
library(ggsci)      # used for color palette (pal_jco)
library(knitr)      # used for chunk expansion

# Load timezones
source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/refs/heads/main/scripts/general_parameters.R")

# Load colors for head, chest, and wrist position
pos_colors <- setNames(pal_jco()(3), c("head", "chest", "wrist"))
```

## Function to count participants and participant-days

This function is executed at multiple points in the script to count the number of participants and participant-days remaining after each processing step. The resulting counts are presented in @tbl-data-counts.

```{r}
#| label: define-counter

# Initialize collector once
participant_stats <- list()

count_participants <- function(processing_step, data) {
  
  # Combine the datasets
  data <- bind_rows(data)
  
  result <- tibble(
    step = processing_step,
    participants = n_distinct(data$Id),
    participant_days = nrow(distinct(ungroup(data), Id, date = as_date(Datetime)))
  )
  
  participant_stats[[processing_step]] <<- result
  
  invisible(result) # Do not print results when calling the function
}
```

## Import light data

```{r}
#| label: import-light-data
#| message: false

# Load data from each site. These have been previously trimmed to only include measurements during the trial periods.
load_position <- function(dosimeter, site, position_label) {
  load_data(dosimeter, site = site) |>
    select(Id, Datetime, MEDI) |>
    mutate(position = position_label)
}

# Build the combined per site dataset
build_site <- function(site) {
  bind_rows(
    load_position("light_glasses", site, "h"),
    load_position("light_chest", site, "c"),
    load_position("light_wrist", site, "w")
  )
}

BAUA <- build_site("BAUA")
FUSPCEU <- build_site("FUSPCEU")
IZTECH <- build_site("IZTECH")
KNUST <- build_site("KNUST")
RISE <- build_site("RISE")
THUAS <- build_site("THUAS")
TUM <- build_site("TUM")
UCR <- build_site("UCR")

# MPI is imported manually, as there were no chest or wrist measurements in the repository
MPI <- load_position("light_glasses", "MPI", "h")


# Count participants and participant-days
count_participants("Trimmed by trial period",
                   list(BAUA, FUSPCEU, IZTECH, MPI, KNUST, RISE, THUAS, TUM, UCR))
```

Additionally, the number of participants and participant-days are counted for each dosimeter position.

```{r}
#| label: count-positions

datasets <- lst(BAUA, FUSPCEU, IZTECH, MPI, KNUST, RISE, THUAS, TUM, UCR)

summary_tbl <- datasets |>
  imap(~ .x |>
         ungroup() |>
         mutate(date = as_date(Datetime)) |>
         group_by(position) |>
         summarise(
           participants     = n_distinct(Id),
           participant_days = n_distinct(Id, date),
           .groups = "drop"
         ) |>
         mutate(dataset = .y, .before = 1)) |>
  list_rbind() |>
  # Format table contents
  pivot_wider(
    names_from  = position,
    values_from = c(participants, participant_days),
    values_fill = 0,
    names_glue  = "{position}_{.value}"
  ) |>
  select(dataset,
         starts_with("h_"),
         starts_with("c_"),
         starts_with("w_"))

# Calculate totals for each column
summary_tbl <- summary_tbl |>
  bind_rows(
    summary_tbl |>
      summarise(across(where(is.numeric), sum)) |>
      mutate(dataset = "Total")
  )

# Render table
summary_tbl |>
  gt(rowname_col = "dataset") |>
  tab_spanner(label = "Head (glasses)", columns = starts_with("h_")) |>
  tab_spanner(label = "Chest", columns = starts_with("c_")) |>
  tab_spanner(label = "Wrist", columns = starts_with("w_")) |>
  cols_label(
    ends_with("_participants")     ~ "Participants",
    ends_with("_participant_days") ~ "Participant-days"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_stub(),
      cells_column_labels(),
      cells_column_spanners()
    )
  ) |>
  tab_style(
    style = list(cell_text(weight = "bold"), cell_fill(color = "#f2f2f2")),
    locations = list(
      cells_stub(rows = "Total"),
      cells_body(rows = nrow(summary_tbl))
    )
  ) |>
  tab_header(
    title = "Trimmed by trial period"
  )
```

As no measurements at the chest or wrist were collected at the MPI site, this site is omitted from this point onward.

```{r}
#| label: list-datasets

datasets <- lst(BAUA, FUSPCEU, IZTECH, KNUST, RISE, THUAS, TUM, UCR)
rm(BAUA, FUSPCEU, IZTECH, KNUST, RISE, THUAS, TUM, UCR)
```

## Combine measurements

```{r}
#| label: combine-measurements

# Set timestamps of the chest and wrist measurements equal to the head measurements
combine_positions <- function(data) {
  h_data <- data |> filter(position == "h")
  c_data <- data |> filter(position == "c")
  w_data <- data |> filter(position == "w")
  
  data2reference(h_data, Reference.data = c_data, Reference.column = MEDI_c) |>
    left_join(
      data2reference(h_data, Reference.data = w_data, Reference.column = MEDI_w) |>
        select(Id, Datetime, MEDI_w),
      by = c("Id", "Datetime")) |>
    rename(MEDI_h = MEDI) |>
    select(-position)
}

datasets <- map(datasets, combine_positions)


# Remove observations where one or more MEDI value is missing
na_summary <- imap(datasets, function(data, name) {
  incomplete <- !complete.cases(data)
  n_before <- nrow(data)
  n_removed <- sum(incomplete)
  
  tibble(
    site = name,
    n_before = n_before,
    n_after = n_before - n_removed,
    n_removed = n_removed,
    pct_removed = n_removed / n_before * 100
  )
}) |> list_rbind()

# Apply the filtering to each dataset in the list
datasets <- map(datasets, ~ .x[complete.cases(.x), ])


# Print summary table
na_summary |>
  gt(rowname_col = "site") |>
  fmt_number(columns = c(n_before, n_after, n_removed),
             decimals = 0, sep_mark = ",") |>
  fmt_number(columns = pct_removed, decimals = 1, pattern = "{x}%") |>
  cols_label(
    n_before = "Before",
    n_after = "After",
    n_removed = "Removed",
    pct_removed = "% Removed"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
      locations = list(cells_stub(), cells_column_labels())
  ) |>
  tab_header(
    title = "Observations removed due to incomplete cases"
  )

# Count participants and participant-days
count_participants("Complete observations only", datasets)
```

At some sites, participants did not wear all three dosimeters, resulting in the removal of a relatively large number of observations. A missing data warning is produced for participants FUSPCEU_S014 and KNUST_S005, which is not problematic since the measurements will be averaged to hourly values.

## Filter for non-wear and out-of-range observations{#sec-filter-nonwear}

Observations, i.e. simultaneous eye-level, chest-worn, and wrist-worn dosimeter measurements at the same point in time, were excluded if non-wear was indicated (i.e. sleep or non-wear reported by the participant in the logbook) or if any of the three measurements exceeded the operating range of the dosimeter (\> 100,000 lx melanopic EDI).

```{r}
#| label: filter-nonwear-outofrange

remove_nonwear <- function(data, site_name) {
  sleep <- load_data("sleepdiaries", site = site_name)
  nonwear <- load_data("wearlog", site = site_name)
  
  exclude_intervals <- bind_rows(
    sleep |> select(Id, int_start = sleepprep, int_end = out_ofbed),
    nonwear |> select(Id, int_start = start, int_end = end)
  )
  
  data |>
    ungroup() |>
    anti_join(
      exclude_intervals,
      # Also removes observations with the exact "boundary" time stamp
      by = join_by(Id, Datetime >= int_start, Datetime <= int_end)
    ) |>
    group_by(Id)
}

remove_high_medi <- function(data) {
  data |>
    filter(if_all(starts_with("MEDI_"), ~ . < 1e5))
}

datasets <- imap(datasets, remove_nonwear)
datasets <- map(datasets, remove_high_medi)


# Count participants and participant-days
count_participants("Filtered for non-wear and out of range cases", datasets)
```

## Visualize light data

```{r}
#| label: visualize-light-data

vis_data <- function(data) {
  plot_data <- data |>
    pivot_longer(
      cols = starts_with("MEDI_"),
      names_to = "position",
      values_to = "MEDI"
    ) |>
    mutate(
      position = sub("^MEDI_", "", position),
      position = case_when(
        position == "h" ~ "head",
        position == "c" ~ "chest",
        position == "w" ~ "wrist"
      ),
      position = factor(position, levels = c("head", "chest", "wrist"))
    ) |>
    # Insert NA rows at gaps larger than 10 minutes
    # This is so geom line can be used, without having a continuous line through the night
    # (as geom line is much faster than geom point)
    arrange(Id, position, Datetime) |>
    group_by(Id, position) |>
    mutate(gap_break = as.numeric(difftime(Datetime, lag(Datetime), units = "mins")) > 10) |>
    mutate(MEDI = if_else(gap_break, NA_real_, MEDI)) |>
    select(-gap_break) |>
    group_by(Id)
  
  gg_days(
    dataset = plot_data,
    aes_col = position,
    geom = "line",
    alpha = 0.6,
    x.axis.limits = \(x) Datetime_limits(x, length = ddays(10))
  ) +
    scale_color_manual(values = pos_colors)
}
```

```{r}
#| label: visualize-light-data2
#| results: asis

chunks <- imap(datasets, \(data, name) {
  fig_height <- n_distinct(data$Id) * 1.8
  knit_expand(text = c(
    "### {{name}}",
    "```{r, fig.width=23, fig.height={{fig_height}}, fig.dpi=72, echo=FALSE, warning=FALSE, message=FALSE}",
    "vis_data(datasets[[\"{{name}}\"]])",
    "```"
  ))
})

cat(knit(text = unlist(chunks), quiet = TRUE))
```

Based on visual inspection, days with (almost) continuously zero signal from all three dosimeters were removed: participants FUSPCEU_S010 (03-11-2024), FUSPCEU_S017 (09-12-2024), IZTECH_S009 (18-03-2025 and 19-03-2025), and THUAS_S007 (11-04-2025).

```{r}
#| label: filter-flat

datasets$FUSPCEU <- datasets$FUSPCEU |>
  filter(!(Id == "FUSPCEU_S010" & as.Date(Datetime) == "2024-11-03"),
         !(Id == "FUSPCEU_S017" & as.Date(Datetime) == "2024-12-09"))

datasets$THUAS <- datasets$THUAS |>
  filter(!(Id == "THUAS_S007" & as.Date(Datetime) == "2025-04-11"))

datasets$IZTECH <- datasets$IZTECH |>
  filter(!(Id == "IZTECH_S009" & as.Date(Datetime) %in% as.Date(c("2025-03-18", "2025-03-19"))))

# Count participants and participant-days
count_participants("Filtered for flat signals", datasets)
```

## Import contextual data

Contextual data is added to each observation: time of day (hourly), day of experiment (integer counting from the first day) site name, photoperiod state (civil dawn and dusk, through LightLogR's `add_photoperiod()` function [@zauner2025]), and light exposure diaries (the primary and secondary lightsource, and columns with booleans for the type of activity). The time of day is represented on an hourly basis using the midpoint of each hour (e.g. 08:00 to 09:00 is represented as 8.5).

```{r}
#| label: import-contextual-data

# Add experiment day and time of day
# timeofday as hourly midpoint (e.g. 08:00-09:00 -> 8.5)
add_day_time <- function(data) {
  data |>
    mutate(
      timeofday = hour(Datetime) + 0.5,
      day = factor(as.integer(as.Date(Datetime) - min(as.Date(Datetime))) + 1)
    )
}

# Add site name
add_site <- function(data, name) {
  data |>
    mutate(site = factor(name))
}

# Add photoperiod state
add_photoperiodstate <- function(data, coords) {
  add_photoperiod(data, coords) |>
    mutate(photoperiod.state = factor(photoperiod.state)) |>
    select(-dawn, -dusk, -photoperiod)
}

# Add light exposure diaries
add_diary <- function(data, site_name) {
  diary <- load_data("lightexposurediary", site = site_name) |>
    select(
      Id, start, end,
      lightsource_primary, lightsource_secondary,
      act_sleep, act_home, act_road_vehicle, act_road_open,
      act_working_indoor, act_working_outdoor, act_free_outdoor, act_other
    ) |>
    mutate(end = end - seconds(1)) # e.g. 11:00:00 -> 10:59:59 (to prevent 2 entries for one timestamp)
  
  data |>
    ungroup() |>
    left_join(
      diary,
      by = join_by(Id, Datetime >= start, Datetime <= end)
    ) |>
    group_by(Id)
}

datasets <- datasets |>
  map(add_day_time) |>
  # Imap also gives the index of the name of dataset
  imap(add_site) |>
  # Coordinates loaded previously from general_parameters.R
  imap(\(data, name) add_photoperiodstate(data, coordinates[[name]])) |>
  imap(add_diary)
```

## Format activity data

```{r}
#| label: print-act-contexts

act_cat <- c(
  "act_sleep", "act_home", "act_road_vehicle", "act_road_open",     
  "act_working_indoor", "act_working_outdoor", "act_free_outdoor", "act_other")


act_overview <- bind_rows(datasets) |> ungroup()
total_n <- nrow(act_overview)


act_overview |>
  count(across(all_of(act_cat)), name = "n") |>
  mutate(
    perc = round(n / total_n * 100, 2),
    flag = case_when(
      if_all(all_of(act_cat), ~ . == FALSE) ~ "all FALSE",
      if_all(all_of(act_cat), is.na) ~ "all NA",
      TRUE ~ ""
    )
  ) |>
  arrange(desc(n)) |>
  gt() |>
  tab_header(
    title = "Distribution of activity category combinations"
  ) |>
  cols_label(
    n    = "Observations",
    perc = "% observations"
  ) |>
  fmt_number(columns = n, decimals = 0, sep_mark = ",") |>
  fmt_number(columns = perc, decimals = 1, pattern = "{x}%") |>
  tab_style(
    style = cell_fill(color = "orange"),
    locations = cells_body(rows = flag %in% c("all FALSE", "all NA"))
  ) |>
  tab_style(
    style = cell_fill(color = "lightgreen"),
    locations = map(act_cat, ~ cells_body(
      columns = all_of(.x),
      rows = !!sym(.x) == TRUE
    ))
  ) |>
  cols_hide(columns = flag) |>
  tab_options(table.font.size = px(10))
```

While measurements taken during sleep were excluded during previous filtering (section @sec-filter-nonwear), some may still be categorized as act_sleep due to the hourly time resolution of the contextual data.

Observations with missing contextual data (either all activities set to NA or FALSE) were removed.

```{r}
#| label: remove-empty-contexts

remove_no_activity <- function(data) {
  data |>
    filter(if_any(
      c(act_sleep, act_home, act_road_vehicle, act_road_open,
        act_working_indoor, act_working_outdoor, act_free_outdoor, act_other),
      ~ !is.na(.) & .
    ))
}

datasets <- map(datasets, remove_no_activity)

# Count participants and participant-days
count_participants("Complete contextual data only", datasets)
```

### Format environment (indoor/outdoor)

Based on the activity data, a column (`environment`) was added which indicates whether the participant was indoors (in a building) or outdoors:

-   `indoor` (*act_home* or *act_working_indoor* or *act_sleep* or *act_home* and *act_working_indoor* or *act_home* and *act_sleep*)

-   `outdoor` (*act_road_vehicle* or *act_free_outdoor* or *act_road_open* or *act_working_outdoor* or *act_road_vehicle* and *act_road_open* or *act_free_outdoor* and *act_road_open* or *act_free_outdoor* and *act_road_vehicle* or *act_road_open* and *act_working_outdoor*)

-   `other_env` (Any other category or combination)

```{r}
#| label: environment-format

add_environment <- function(data) {
  data |>
    ungroup() |>
    mutate(
      environment = case_when(
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == TRUE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == TRUE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == TRUE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == TRUE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == TRUE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == TRUE & act_home == TRUE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) ~ "indoor",
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == TRUE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == TRUE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == TRUE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == TRUE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == TRUE & act_other == FALSE) ~ "outdoor",
        TRUE ~ "other_env"
      ),
      environment = factor(environment, levels = c("indoor", "outdoor", "other_env"))
    )
}

datasets <- map(datasets, add_environment)
```

### Format simplified activities

To decrease the number of combinations, the data is formatted (`act_context` column) into general categories deemed representative of the various activity contexts:

-   `home` (*act_home*)

-   `working_indoor` (*act_working_indoor*)

-   `road_vehicle` (*act_road_vehicle*)

-   `outdoor_act` (*act_free_outdoor* or *act_road_open* or *act_working_outdoor* or *act_road_open* and *act_working_outdoor* or *act_road_open* and *act_free_outdoor*)

-   `other_act` (Any other category or combination)

Additionally, the activity boolean columns are dropped from the datasets, as they are no longer needed.

```{r}
#| label: activity-format

add_act_context <- function(data) {
  data |>
    ungroup() |>
    mutate(
      act_context = case_when(
        act_home == TRUE &
          act_sleep == FALSE & act_road_vehicle == FALSE & act_road_open == FALSE &
          act_working_indoor == FALSE & act_working_outdoor == FALSE &
          act_free_outdoor == FALSE & act_other == FALSE ~ "home",
        act_working_indoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_outdoor == FALSE &
          act_free_outdoor == FALSE & act_other == FALSE ~ "working_indoor",
        act_road_vehicle == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_open == FALSE &
          act_working_indoor == FALSE & act_working_outdoor == FALSE &
          act_free_outdoor == FALSE & act_other == FALSE ~ "road_vehicle",
        (act_free_outdoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == FALSE & act_working_indoor == FALSE &
          act_working_outdoor == TRUE & act_other == FALSE) |
        (act_free_outdoor == FALSE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == TRUE & act_other == FALSE) |
        (act_free_outdoor == TRUE &
          act_sleep == FALSE & act_home == FALSE & act_road_vehicle == FALSE &
          act_road_open == TRUE & act_working_indoor == FALSE &
          act_working_outdoor == FALSE & act_other == FALSE) ~ "outdoor_act",
        TRUE ~ "other_act"
      ),
      act_context = factor(act_context, levels = c(
        "home", "working_indoor", "road_vehicle", "outdoor_act", "other_act"
      ))
    ) |>
    select(-c(act_sleep, act_home, act_road_vehicle, act_road_open,
              act_working_indoor, act_working_outdoor, act_free_outdoor, act_other))
}

datasets <- map(datasets, add_act_context)
```

### Format indoor work lighting categories

A column (`work_light_context`) was added to describe the primary light source when participants were working indoors. For observations where this was not the case, the column is set to NA. Additionally, the primary and secondary light source columns are dropped from the datasets, as they are no longer needed.

```{r}
#| label: work-light-format

add_work_light_context <- function(data) {
  data |>
    ungroup() |>
    mutate(
      work_light_context = case_when(
        act_context == "working_indoor" &
          lightsource_primary == "Electric light source indoors" ~ "electric",
        act_context == "working_indoor" &
          lightsource_primary == "Daylight indoors" ~ "daylight",
        act_context == "working_indoor" &
          lightsource_primary == "Emissive display light" ~ "display"
      ),
      work_light_context = factor(
        work_light_context,
        levels = c("electric", "daylight", "display")
      )
    ) |>
    select(-c(lightsource_primary, lightsource_secondary))
}

datasets <- map(datasets, add_work_light_context)
```

## Calculate wear position errors

The wear position errors (mean hourly log-transformed differences between eye-level and body-worn dosimeter measurements, in melanopic EDI) were computed using a paired-differences approach. As the eye-level and body-worn measurements were at (approximately) the same time stamp, they were conducted under the same illumination field. Computing the log-difference within each pair before averaging preserves this paired structure and allows investigation of the measurement error. The alternative of taking the log of each position's hourly mean melanopic EDI discards the pairing and hence is thus only valid for an outcome focused analysis.

For wear position x, where x refers to either the chest- or wrist-worn dosimeter, the (hourly aggregated) wear position error (relative to a dosimeter worn on a pair of glasses) is defined as

$$
\bar{D}_x = \frac{1}{n}\sum_{i=1}^{n}\left[\log_{10}(E_{x,i} + 0.1) - \log_{10}(E_{glasses,i} + 0.1)\right]
$$where $E_{glasses,i}$ denotes the eye-level measurement and $E_{x,i}$ denotes the measurement at wear position $x$, both recorded at timestamp $i$, and $n$ is the number of measurements within the hour.

```{r}
#| label: calculate-logdif-average

to_hourly <- function(data) {
  data |>
    ungroup() |>
    mutate(
      absdif_c = MEDI_c - MEDI_h,
      absdif_w = MEDI_w - MEDI_h,
      logdif_c = log10(MEDI_c + 0.1) - log10(MEDI_h + 0.1),
      logdif_w = log10(MEDI_w + 0.1) - log10(MEDI_h + 0.1)
    ) |>
    group_by(Id, timeofday, day, site, environment, act_context, work_light_context) |>
    summarise(
      MEDI_h = mean(MEDI_h),
      MEDI_c = mean(MEDI_c),
      MEDI_w = mean(MEDI_w),
      absdif_c = mean(absdif_c),
      absdif_w = mean(absdif_w),
      logdif_c  = mean(logdif_c),
      logdif_w = mean(logdif_w),
      photoperiod.state = names(which.max(table(photoperiod.state))),
      .groups = "drop"
    ) |>
    mutate(photoperiod.state = as.factor(photoperiod.state)) |>
    arrange(Id, day, timeofday)
}

datasets <- map(datasets, to_hourly)
```

## Combine datasets

```{r}
#| label: combine-datasets

prepared_context_data <- bind_rows(datasets)
```

## Data overview

The distributions of the differences between (hourly mean) log-transformed eye-level and body-worn dosimeters measurements (`logdif_c` and `logdif_w`)) during daytime, nighttime, indoor, and outdoor conditions are given in @fig-2d-dis-day to @fig-2d-dis-outdoor.

```{r}
#| label: fig-2d-dis-func

make_hcw_plot <- function(data, filter_expr) {
  
  ALL_hcw <- data |>
    filter({{ filter_expr }}) |>
    select(Id, MEDI_h, MEDI_c, MEDI_w)

  # Non-zero / zero splits
  ALL_hc <- ALL_hcw |> filter(MEDI_h != 0, MEDI_c != 0) |> mutate(across(c(MEDI_h, MEDI_c), log10))
  ALL_hw <- ALL_hcw |> filter(MEDI_h != 0, MEDI_w != 0) |> mutate(across(c(MEDI_h, MEDI_w), log10))
  ALL_h0_c <- ALL_hcw |> filter(MEDI_h == 0, MEDI_c != 0) |> mutate(MEDI_c = log10(MEDI_c)) |> select(MEDI_c)
  ALL_h0_w <- ALL_hcw |> filter(MEDI_h == 0, MEDI_w != 0) |> mutate(MEDI_w = log10(MEDI_w)) |> select(MEDI_w)
  ALL_h_c0 <- ALL_hcw |> filter(MEDI_c == 0, MEDI_h != 0) |> mutate(MEDI_h = log10(MEDI_h)) |> select(MEDI_h)
  ALL_h_w0 <- ALL_hcw |> filter(MEDI_w == 0, MEDI_h != 0) |> mutate(MEDI_h = log10(MEDI_h)) |> select(MEDI_h)
  ALL_h0_c0 <- ALL_hcw |> filter(MEDI_h == 0, MEDI_c == 0)
  ALL_h0_w0 <- ALL_hcw |> filter(MEDI_h == 0, MEDI_w == 0)

  # Binning
  binwidth <- 0.2

  bin_2d <- function(df, v1, v2) {
    df |> ungroup() |>
      mutate(
        "{v1}_bin" := floor(.data[[v1]] / binwidth) * binwidth,
        "{v2}_bin" := floor(.data[[v2]] / binwidth) * binwidth
      )
  }

  bin_1d <- function(df, v) {
    df |> ungroup() |>
      mutate("{v}_bin" := floor(.data[[v]] / binwidth) * binwidth)
  }

  ALL_hc_bins <- bin_2d(ALL_hc, "MEDI_h", "MEDI_c")
  ALL_hw_bins <- bin_2d(ALL_hw, "MEDI_h", "MEDI_w")
  ALL_h0_c_bins <- bin_1d(ALL_h0_c, "MEDI_c")
  ALL_h0_w_bins <- bin_1d(ALL_h0_w, "MEDI_w")
  ALL_h_c0_bins <- bin_1d(ALL_h_c0, "MEDI_h")
  ALL_h_w0_bins <- bin_1d(ALL_h_w0, "MEDI_h")

  # Counts
  clamp_low <- function(df, ...) {
    vars <- ensyms(...)
    df |>
      mutate(across(c(!!!vars), \(x) ifelse(x < 0, -0.2, x))) |>
      group_by(across(c(!!!vars))) |>
      summarise(n = sum(n), .groups = "drop")
  }

  ALL_hc_bins_counts <- ALL_hc_bins  |> count(MEDI_h_bin, MEDI_c_bin, name = "n") |> clamp_low(MEDI_h_bin, MEDI_c_bin)
  ALL_hw_bins_counts <- ALL_hw_bins  |> count(MEDI_h_bin, MEDI_w_bin, name = "n") |> clamp_low(MEDI_h_bin, MEDI_w_bin)
  ALL_h0_c_bins_counts <- ALL_h0_c_bins |> count(MEDI_c_bin, name = "n") |> clamp_low(MEDI_c_bin)
  ALL_h0_w_bins_counts <- ALL_h0_w_bins |> count(MEDI_w_bin, name = "n") |> clamp_low(MEDI_w_bin)
  ALL_h_c0_bins_counts <- ALL_h_c0_bins |> count(MEDI_h_bin, name = "n") |> clamp_low(MEDI_h_bin)
  ALL_h_w0_bins_counts <- ALL_h_w0_bins |> count(MEDI_h_bin, name = "n") |> clamp_low(MEDI_h_bin)
  ALL_h0_c0_count <- ALL_h0_c0 |> ungroup() |> count(MEDI_h, name = "n")
  ALL_h0_w0_count <- ALL_h0_w0 |> ungroup() |> count(MEDI_h, name = "n")

  # Shared fill scale
  global_min <- min(ALL_hc_bins_counts$n, ALL_hw_bins_counts$n,
                    ALL_h0_c_bins_counts$n, ALL_h0_w_bins_counts$n,
                    ALL_h_c0_bins_counts$n, ALL_h_w0_bins_counts$n,
                    ALL_h0_c0_count$n, ALL_h0_w0_count$n)
  global_max <- max(ALL_hc_bins_counts$n, ALL_hw_bins_counts$n,
                    ALL_h0_c_bins_counts$n, ALL_h0_w_bins_counts$n,
                    ALL_h_c0_bins_counts$n, ALL_h_w0_bins_counts$n,
                    ALL_h0_c0_count$n, ALL_h0_w0_count$n)
  log_breaks <- 10^seq(floor(log10(5)), ceiling(log10(global_max)), by = 1)

  fill_scale <- scale_fill_viridis_c(
    trans = "log10", option = "plasma", name = "Count",
    limits = c(global_min, global_max),
    breaks = log_breaks, labels = label_number()
  )

  # Layout design
  design <- c(
    area(t = 1, l = 1, b = 1, r = 1),
    area(t = 1, l = 2, b = 1, r = 2),
    area(t = 2, l = 1, b = 2, r = 1),
    area(t = 2, l = 2, b = 2, r = 2)
  )

  # Plot builder helpers
  tick_segments <- function(positions, axis, direction, offset = 0.2) {
    # axis: "x" or "y", direction: "in" or "out"
    if (axis == "x") {
      y_base <- if (direction == "out") 0 else positions[length(positions)] + offset
      y_tip  <- if (direction == "out") -offset else positions[length(positions)]
      df <- data.frame(x = positions, xend = positions,
                       y = if (direction == "out") 0 else 5.2,
                       yend = if (direction == "out") -offset else 5)
    } else {
      df <- data.frame(y = positions, yend = positions,
                       x = if (direction == "out") 0 else 5.2,
                       xend = if (direction == "out") -offset else 5)
    }
    geom_segment(data = df, aes(x = x, xend = xend, y = y, yend = yend),
                 inherit.aes = FALSE, linewidth = 0.6)
  }

  make_main_plot <- function(counts, x_bin, y_bin) {
    ggplot() +
      geom_rect(data = counts,
                aes(xmin = .data[[x_bin]], xmax = .data[[x_bin]] + binwidth,
                    ymin = .data[[y_bin]], ymax = .data[[y_bin]] + binwidth,
                    fill = n)) +
      coord_fixed() +
      scale_x_continuous(limits = c(-0.2, 5.2), expand = c(0,0),
                         breaks = seq(0, 5), labels = NULL) +
      scale_y_continuous(limits = c(-0.2, 5.2), expand = c(0,0),
                         breaks = seq(0, 5), labels = NULL) +
      fill_scale +
      geom_abline(slope = 1, intercept = 0, color = "black",
                  linetype = "dashed", linewidth = 0.8) +
      theme_minimal() +
      theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)) +
      # tick marks
      geom_segment(data = data.frame(x=seq(0,5), xend=seq(0,5), y=0, yend=-0.2),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE, linewidth=0.6) +
      geom_segment(data = data.frame(y=seq(0,5), yend=seq(0,5), x=0, xend=-0.2),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE, linewidth=0.6) +
      geom_segment(data = data.frame(x=seq(0,5), xend=seq(0,5), y=5.2, yend=5),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE, linewidth=0.6) +
      geom_segment(data = data.frame(y=seq(0,5), yend=seq(0,5), x=5.2, xend=5),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE, linewidth=0.6) +
      labs(x = NULL, y = NULL, fill = "Count")
  }

  make_h0_plot <- function(counts, x_bin, x_lab) {
    ggplot() +
      geom_rect(data = counts,
                aes(xmin = .data[[x_bin]], xmax = .data[[x_bin]] + binwidth,
                    ymin = 0, ymax = binwidth, fill = n)) +
      coord_fixed() +
      scale_x_continuous(limits = c(-0.2, 5.2), expand = c(0,0),
                         breaks = seq(0, 5),
                         labels = function(x) label_number(accuracy=1)(10^x)) +
      scale_y_continuous(limits = c(0, binwidth), expand = c(0,0),
                         breaks = binwidth/2, labels = NULL) +
      fill_scale + guides(fill = "none") +
      labs(x = x_lab, y = NULL) +
      theme_minimal() +
      theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
            panel.grid = element_blank()) +
      geom_segment(data = data.frame(x=seq(0,5), xend=seq(0,5), y=0, yend=0.2),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE,
                   linewidth=0.6, color="black")
  }

  make_c0_plot <- function(counts, y_lab) {
    ggplot() +
      geom_rect(data = counts,
                aes(xmin = 0, xmax = binwidth,
                    ymin = MEDI_h_bin, ymax = MEDI_h_bin + binwidth, fill = n)) +
      coord_fixed() +
      scale_x_continuous(limits = c(0, binwidth), expand = c(0,0),
                         breaks = binwidth/2, labels = NULL) +
      scale_y_continuous(limits = c(-0.2, 5.2), expand = c(0,0),
                         breaks = seq(0, 5),
                         labels = function(x) label_number(accuracy=1)(10^x)) +
      fill_scale + guides(fill = "none") +
      labs(x = NULL, y = y_lab) +
      theme_minimal() +
      theme(panel.border = element_rect(colour = "black", fill = NA, linewidth = 1),
            panel.grid = element_blank()) +
      geom_segment(data = data.frame(x=0, xend=0.2, y=seq(0,5), yend=seq(0,5)),
                   aes(x=x,xend=xend,y=y,yend=yend), inherit.aes=FALSE,
                   linewidth=0.6, color="black")
  }

  make_h0c0_plot <- function(counts) {
    ggplot() +
      geom_rect(data = counts,
                aes(xmin=0, xmax=binwidth, ymin=0, ymax=binwidth, fill=n)) +
      coord_fixed() +
      scale_x_continuous(limits=c(0,binwidth), expand=c(0,0), breaks=binwidth/2, labels=0) +
      scale_y_continuous(limits=c(0,binwidth), expand=c(0,0), breaks=binwidth/2, labels=0) +
      fill_scale + guides(fill = "none") +
      theme_minimal() +
      theme(panel.border = element_rect(colour="black", fill=NA, linewidth=1),
            panel.grid = element_blank())
  }

  # Assemble the two composite plots
  plot_headvschest <- (
    make_c0_plot(ALL_h_c0_bins_counts, "Eye-level dosimeter [m-EDI lx]") +
    make_main_plot(ALL_hc_bins_counts, "MEDI_c_bin", "MEDI_h_bin") +
    make_h0c0_plot(ALL_h0_c0_count) +
    make_h0_plot(ALL_h0_c_bins_counts, "MEDI_c_bin", "Chest-worn dosimeter [m-EDI lx]")
  ) + plot_layout(design = design)

  plot_headvswrist <- (
    make_c0_plot(ALL_h_w0_bins_counts, NULL) +
    make_main_plot(ALL_hw_bins_counts, "MEDI_w_bin", "MEDI_h_bin") +
    make_h0c0_plot(ALL_h0_w0_count) +
    make_h0_plot(ALL_h0_w_bins_counts, "MEDI_w_bin", "Wrist-worn dosimeter [m-EDI lx]")
  ) + plot_layout(design = design)

  (plot_headvschest | plot_headvswrist) +
    plot_layout(guides = "collect") 
  }
```

```{r}
#| label: fig-2d-dis-day
#| fig-cap: "Distribution of head-worn and chest/wrist-worn dosimeter measurements during daytime."
#| fig-width: 13
#| fig-height: 8

make_hcw_plot(prepared_context_data, photoperiod.state == "day")
```

```{r}
#| label: fig-2d-dis-night
#| fig-cap: "Distribution of head-worn and chest/wrist-worn dosimeter measurements during nighttime."
#| fig-width: 13
#| fig-height: 8

make_hcw_plot(prepared_context_data, photoperiod.state == "night")
```

```{r}
#| label: fig-2d-dis-indoor
#| fig-cap: "Distribution of head-worn and chest/wrist-worn dosimeter measurements in indoor environments."
#| fig-width: 13
#| fig-height: 8

make_hcw_plot(prepared_context_data, environment == "indoor")
```

```{r}
#| label: fig-2d-dis-outdoor
#| fig-cap: "Distribution of head-worn and chest/wrist-worn dosimeter measurements in outdoor environments."
#| fig-width: 13
#| fig-height: 8

make_hcw_plot(prepared_context_data, environment == "outdoor")
```

```{r}
#| label: fig-dif-distr
#| fig-cap: "Distributions of hourly mean errors."
#| fig-width: 13
#| fig-height: 8

hist_data <- prepared_context_data |>
  ungroup() |>
  select(absdif_c, absdif_w, logdif_c, logdif_w) |>
  pivot_longer(
    cols = c(absdif_c, absdif_w, logdif_c, logdif_w),
    names_to = c("scale_type", "position"),
    names_pattern = "(absdif|logdif)_(c|w)",
    values_to = "value"
  ) |>
  mutate(
    position = factor(position, levels = c("c", "w"), labels = c("Chest", "Wrist"))
  )

p_abs <- hist_data |>
  filter(scale_type == "absdif") |>
  ggplot(aes(x = value, fill = position, color = position)) +
  geom_histogram(binwidth = 50, alpha = 0.85) +
  facet_wrap(~ position) +
  theme_cowplot() +
  scale_fill_manual(values = c(Chest = pos_colors[["chest"]],
                               Wrist = pos_colors[["wrist"]])) +
  scale_color_manual(values = c(Chest = pos_colors[["chest"]],
                                Wrist = pos_colors[["wrist"]])) +
  scale_y_continuous(labels = label_comma()) +
  coord_cartesian(xlim = c(-2500, 2500)) +
  guides(fill = "none", color = "none") +
  labs(x = "Absolute difference [m-EDI]", y = "Count (n)") +
  theme(panel.spacing = unit(1, "lines"))

p_log <- hist_data |>
  filter(scale_type == "logdif") |>
  ggplot(aes(x = value, fill = position)) +
  geom_histogram(binwidth = 0.1, alpha = 0.85) +
  facet_wrap(~ position) +
  theme_cowplot() +
  scale_fill_manual(values = c(Chest = pos_colors[["chest"]],
                               Wrist = pos_colors[["wrist"]])) +
  scale_y_continuous(labels = label_comma()) +
  guides(fill = "none") +
  labs(x = "Logarithmic difference [log10 m-EDI]", y = "Count (n)") +
  theme(panel.spacing = unit(1, "lines"),
        strip.background = element_blank(),
        strip.text = element_blank())

p_abs / p_log
```

Both logarithmic difference distributions are leptokurtic, meaning they have a sharp, narrow peak and heavy tails, as confirmed by kurtosis values substantially exceeding 3:

```{r}
#| label: kurtosis-test
cat("kurtosis logdif_c:", kurtosis(prepared_context_data$logdif_c), "\n")
cat("kurtosis logdif_w:", kurtosis(prepared_context_data$logdif_w), "\n")
```

```{r}
#| label: data-overview
#| fig-width: 18
#| fig-height: 5

# Observations per participant
prepared_context_data |>
  ungroup() |>
  count(Id, name = "n") |>
  ggplot(aes(x = Id, y = n)) +
  geom_col() +
  labs(x = "", y = "Number of observations") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

```{r}
#| label: data-overview2

summarise_by_category <- function(data, category_col, drop_na = FALSE) {
  category_col <- ensym(category_col)
  
  data <- data |> ungroup()
  if (drop_na) {
    data <- data |> filter(!is.na(!!category_col))
  }
  
  obs <- data |>
    count(!!category_col, name = "n_observations") |>
    mutate(perc_observations = n_observations / sum(n_observations) * 100)
  
  participants <- data |>
    distinct(!!category_col, Id) |>
    count(!!category_col, name = "n_participants")
  
  left_join(obs, participants, by = as_string(category_col))
}

format_summary_gt <- function(summary_df, category_col, title) {
  summary_df |>
    gt(rowname_col = category_col) |>
    fmt_number(columns = c(n_observations, n_participants),
               decimals = 0, sep_mark = ",") |>
    fmt_number(columns = perc_observations, decimals = 1, pattern = "{x}%") |>
    cols_label(
      n_observations = "Observations",
      perc_observations = "% observations",
      n_participants = "Participants"
    ) |>
    tab_footnote(
      footnote = "Number of unique participants in subset.",
      locations = cells_column_labels(columns = n_participants)
    ) |>
    tab_style(
      style = cell_text(weight = "bold"),
      locations = list(cells_stub(), cells_column_labels())
    ) |>
    tab_header(title = title)
}


# Observations per environment
prepared_context_data |>
  summarise_by_category(environment) |>
  format_summary_gt("environment", "Observations per environment")

# Observations per activity
prepared_context_data |>
  summarise_by_category(act_context) |>
  format_summary_gt("act_context", "Observations per activity")

# Observations per indoor work light context
prepared_context_data |>
  summarise_by_category(work_light_context, drop_na = TRUE) |>
  format_summary_gt("work_light_context", "Observations per indoor work light context")

# Observations per site
prepared_context_data |>
  summarise_by_category(site) |>
  format_summary_gt("site", "Observations per site")

# Observations per photoperiod.state
prepared_context_data |>
  summarise_by_category(photoperiod.state) |>
  format_summary_gt("photoperiod.state", "Observations per photoperiod state")

# Observations per timeofday
prepared_context_data |>
  ungroup() |>
  count(timeofday, name = "n") |>
  ggplot(aes(x = timeofday, y = n)) +
  geom_col() +
  scale_x_continuous(breaks = seq(0, 23, by = 3)) +
  scale_y_continuous(labels = label_comma()) +
  labs(x = "Time of day", y = "Number of observations",
       title = "Observations per time of day") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

```{r}
#| label: data-overview3

# Remove columns that will not be used in the statistical analysis
prepared_context_data <- prepared_context_data |>
  ungroup() |>
  select(-absdif_c, -absdif_w, -MEDI_h, -MEDI_c, -MEDI_w)

# Display variable overview
prepared_context_data |>
  summarise(across(everything(), list(
    type = ~ class(.)[1],
    example = ~ as.character(.[!is.na(.)][100]),
    reference = ~ if (is.factor(.)) levels(.)[1] else NA_character_
  ))) |>
  pivot_longer(everything(),
               names_to  = c("variable", ".value"),
               names_sep = "_(?=[^_]+$)") |>
  mutate(example = ifelse(
    !is.na(suppressWarnings(as.numeric(example))),
    suppressWarnings(as.character(round(as.numeric(example), 2))),
    example
  )) |>
  gt(rowname_col = "variable") |>
  cols_label(
    type = "Type",
    example = "Example value",
    reference = "Reference level"
  ) |>
  sub_missing(missing_text = "—") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_stub(), cells_column_labels())
  ) |>
  tab_header(
    title = "Variables in the analysis dataset"
  ) |>
  tab_footnote(
    footnote  = "Reference level shown only for factor variables.",
    locations = cells_column_labels(columns = reference)
  )
```

All datasets are saved as one `.RData` file.

```{r}
#| label: save-data

save(prepared_context_data, file = here("data/prepared_context_data.RData"))

```

## Overview of filtering counts

```{r}
#| label: tbl-data-counts
#| tbl-cap: "Data counts for each processing step"

participant_summary <- bind_rows(participant_stats)

participant_summary |>
  gt(rowname_col = "step") |>
  fmt_number(columns = c(participants, participant_days),
             decimals = 0, sep_mark = ",") |>
  cols_label(
    participants     = "Participants",
    participant_days = "Participant-days"
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_stub(), cells_column_labels(), cells_title(groups = "title"))
  ) |>
  tab_header(
    title    = "Participants and participant-days at each processing step"
  )
```

## Session info

```{r}
sessionInfo()
```

## Acknowledgments

Code optimization was assisted by Claude Opus 4.7 (Anthropic) and text formatting by Grammarly. All generated suggestions were reviewed and validated by the author, who takes full responsibility for the document.

## References {.unnumbered}
