---
title: "Impact of contextual factors on deviations between eye-level and body-worn dosimeter measurements - Data import"
format:
  html:
    toc: true
    embed-resources: true
authors: "Sietse de Vries"
bibliography: references.bib
---

## Summary

This file describes the import and filtering of eye-level and body-worn dosimeter measurements.

Personal light exposure measurements were collected at multiple sites as part of the MeLiDos project, following the protocol described by Guidolin et al. [@guidolin2024]. The dataset of each site is available on GitHub:

-   BAUA, Dortmund, Germany [@broszio2025]

-   FUSPCEU, Madrid, Spain [@baezamoyano2025]

-   IZTECH, Izmir, Turkey [@didikoglu2025]

-   KNUST, Kumasi, Ghana [@akuffo2025]

-   RISE, Borås, Sweden [@nilssontengelin2026]

-   THUAS, the Hague, the Netherlands [@aerts2025a]

-   TUM, Munich, Germany [@hilden2025a]

-   UCR, San Pedro, San José, Costa Rica [@sancho-salas2026a]

Each dataset contains dosimeter measurements from three body positions (eye-level, chest, and wrist) alongside contextual data. Measurements were temporally aligned using LightLogR's `data2reference()` function [@zauner2025], and only time points with simultaneous data from all three dosimeters were retained. Measurements during sleep (derived from `Sleepdiaries.RData` ("bedtime" and "out_ofbed")) and non-wear (derived from `Wearlog.RData` ("Start" and "end")) were removed from the datasets.

Light level differences (m-EDI) between each body-worn dosimeter (chest or wrist) and the head dosimeter were calculated as absolute differences (`absdif_x = MEDI_x − MEDI_h`) and log-transformed differences (`logdif_x = log10(MEDI_x + 0.1) − log10(MEDI_h + 0.1)`), where x refers to either the chest (c) or wrist (w) dosimeter.

Participants for whom one of the dosimeters recorded at a non-standard sampling interval, i.e. not at a 10-second interval, were removed: FUSPCEU_S014 (head dosimeter at a 1-minute interval), IZTECH_S016 (wrist dosimeter at 120-second interval), and RISE_S011 (head dosimeter at a 1-second interval). Additionally, days with flat (zero) signals across all three dosimeters were removed for participants FUSPCEU_S010 (03-11-2024), FUSPCEU_S017 (09-12-2024), IZTECH_S009 (18-03-2025 and 19-03-2025), and THUAS_S007 (11-04-2025). Daylight saving time corrections, where applicable (RISE, FUSPCEU, and THUAS), were applied prior to analysis.

```{r}
#| label: load-packages
#| message: false
#| warning: false

library(LightLogR)
library(rio)        # used only to load the study dates excel files
library(data.table) # used to match the light exposure diaries more efficiently
library(here)       # used to export the formatted datasets
library(tidyverse)

# Load timezones
source("https://raw.githubusercontent.com/MeLiDosProject/Data_Metadata_Conventions/refs/heads/main/scripts/general_parameters.R")
```

## Import light data

```{r}
#| label: import-light-data

import_light <- function(base_url) {
  positions <- list(
    h = "light_glasses",
    c = "light_chest",
    w = "light_wrist")
  
  purrr::map(names(positions), function(pos) {
    obj_name <- positions[[pos]]
    con <- url(paste0(base_url, obj_name, ".RData"))
    
    e <- new.env()
    load(con, envir = e)
    close(con)
    
    e[[obj_name]][, c("Id", "Datetime", "MEDI")] |>
      mutate(position = pos)
  }) |>
    purrr::list_rbind()
}

BAUA <- import_light("https://github.com/MeLiDosProject/BroszioEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
FUSPCEU <- import_light("https://github.com/MeLiDosProject/BaezaEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
IZTECH <- import_light("https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
KNUST <- import_light("https://github.com/MeLiDosProject/AkuffoEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
RISE <- import_light("https://github.com/MeLiDosProject/NilssonTengelinEtAl_Dataset_2026/raw/refs/heads/main/data/imported/light/")
THUAS <- import_light("https://github.com/MeLiDosProject/AertsEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
TUM <- import_light("https://github.com/MeLiDosProject/HildenEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
UCR <- import_light("https://github.com/MeLiDosProject/Sancho-SalasEtAl_Dataset_2025/raw/refs/heads/main/data/imported/light/")
```

## Filter light data

```{r}
#| label: filter-light-data

# Exclude participants for whom one of the dosimeters recorded at non-standard sampling intervals
FUSPCEU <- FUSPCEU |> filter(Id != "FUSPCEU_S014") # 60s interval
IZTECH <- IZTECH |> filter(Id != "IZTECH_S016") # 60s interval
RISE <- RISE |> filter(Id != "RISE_S011") # 1s interval


# Exclude any measurements outside of the trial period of each participant
filter_trial <- function(data, url, tz) {
  trial_period <- rio::import(url) |>
    select(subjectID_device, datetime_trial_start, datetime_trial_end) |>
    separate(subjectID_device, into = c("Id", "position"), sep = "_(?=[^_]+$)") |>
    filter(position %in% c("h", "c", "w")) |>
    mutate(
      datetime_trial_start = force_tz(datetime_trial_start, tz),
      datetime_trial_end = force_tz(datetime_trial_end, tz))
  
  data |>
    inner_join(trial_period, by = c("Id", "position")) |>
    filter(Datetime >= datetime_trial_start & Datetime <= datetime_trial_end) |>
    select(-datetime_trial_start, -datetime_trial_end)
}

BAUA <- filter_trial(
  data = BAUA,
  url  = "https://github.com/MeLiDosProject/BroszioEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_BAUA.xlsx",
  tz   = tzs["BAUA"])

FUSPCEU <- filter_trial(
  data = FUSPCEU,
  url  = "https://github.com/MeLiDosProject/BaezaEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_FUSPCEU.xlsx",
  tz   = tzs["FUSPCEU"])

IZTECH <- filter_trial(
  data = IZTECH,
  url  = "https://raw.githubusercontent.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/v1.0.1/data/Study_dates_MeLiDos_IZTECH.xlsx",
  tz   = tzs["IZTECH"])

KNUST <- filter_trial(
  data = KNUST,
  url  = "https://github.com/MeLiDosProject/AkuffoEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_KNUST.xlsx",
  tz   = tzs["KNUST"])

RISE <- filter_trial(
  data = RISE,
  url  = "https://github.com/MeLiDosProject/NilssonTengelinEtAl_Dataset_2026/raw/refs/heads/main/data/Study_dates_MeLiDos_RISE.xlsx",
  tz   = tzs["RISE"])

THUAS <- filter_trial(
  data = THUAS,
  url  = "https://github.com/MeLiDosProject/AertsEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_THUAS.xlsx",
  tz   = tzs["THUAS"])

TUM <- filter_trial(
  data = TUM,
  url  = "https://github.com/MeLiDosProject/HildenEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_TUM.xlsx",
  tz   = tzs["TUM"])

UCR <- filter_trial(
  data = UCR,
  url  = "https://github.com/MeLiDosProject/Sancho-SalasEtAl_Dataset_2025/raw/refs/heads/main/data/Study_dates_MeLiDos_UCR.xlsx",
  tz   = tzs["UCR"])


# Set timestamps of the chest and wrist measurements equal to the head measurements through LightLogR's data2reference function.
combine_positions <- function(data) {
  h <- data |> filter(position == "h")
  c <- data |> filter(position == "c")
  w <- data |> filter(position == "w")
  
  data2reference(h, Reference.data = c, Reference.column = MEDI_c) |>
    left_join(
      data2reference(h, Reference.data = w, Reference.column = MEDI_w) |>
        select(Id, Datetime, MEDI_w),
      by = c("Id", "Datetime")) |>
    rename(MEDI_h = MEDI) |>
    select(-position)
}

BAUA <- combine_positions(BAUA)
FUSPCEU <- combine_positions(FUSPCEU)
IZTECH <- combine_positions(IZTECH)
KNUST <- combine_positions(KNUST)
RISE <- combine_positions(RISE)
THUAS <- combine_positions(THUAS)
TUM <- combine_positions(TUM)
UCR <- combine_positions(UCR)

# Remove rows with missing values, which can occur when no chest or wrist measurement fell within 10 seconds of a corresponding head measurement. At some sites, participants did not wear all three dosimeters continuously, resulting in a relatively large number of rows being removed.
for (name in c("BAUA", "FUSPCEU", "IZTECH", "KNUST", "RISE", "THUAS", "TUM", "UCR")) {
  df <- get(name)
  incomplete <- !complete.cases(df)
  cat(name, ":", sum(incomplete),
      sprintf("NA rows removed (%.1f%%)\n", sum(incomplete) / nrow(df) * 100))
  assign(name, df[!incomplete, ])
}


# Exclude measurements during sleep or non-wear
remove_intervals <- function(data, url, obj_name, start_col, end_col) {
  e <- new.env()
  load(url(url), envir = e)
  
  intervals <- e[[obj_name]] |>
    select(Id, {{ start_col }}, {{ end_col }})
  
  data |>
    anti_join(
      data |>
        inner_join(intervals, by = "Id", relationship = "many-to-many") |>
        filter(Datetime >= {{ start_col }}, Datetime <= {{ end_col }}) |>
        distinct(Id, Datetime),
      by = c("Id", "Datetime"))
}

BAUA <- remove_intervals(BAUA,
                         url = "https://github.com/MeLiDosProject/BroszioEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                         obj_name = "sleepdiary",
                         start_col = bedtime,
                         end_col = out_ofbed)
BAUA <- remove_intervals(BAUA,
                         url = "https://github.com/MeLiDosProject/BroszioEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                         obj_name = "wearlog",
                         start_col = start,
                         end_col = end)

FUSPCEU <- remove_intervals(FUSPCEU,
                            url = "https://github.com/MeLiDosProject/BaezaEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                            obj_name = "sleepdiary",
                            start_col = bedtime,
                            end_col = out_ofbed)
FUSPCEU <- remove_intervals(FUSPCEU,
                            url = "https://github.com/MeLiDosProject/BaezaEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                            obj_name = "wearlog",
                            start_col = start,
                            end_col = end)

IZTECH <- remove_intervals(IZTECH,
                           url = "https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                           obj_name = "sleepdiary",
                           start_col = bedtime,
                           end_col = out_ofbed)
IZTECH <- remove_intervals(IZTECH,
                           url = "https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                           obj_name = "wearlog",
                           start_col = start,
                           end_col = end)

KNUST <- remove_intervals(KNUST,
                          url = "https://github.com/MeLiDosProject/AkuffoEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                          obj_name = "sleepdiary",
                          start_col = bedtime,
                          end_col = out_ofbed)
KNUST <- remove_intervals(KNUST,
                          url = "https://github.com/MeLiDosProject/AkuffoEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                          obj_name = "wearlog",
                          start_col = start,
                          end_col = end)

RISE <- remove_intervals(RISE,
                         url = "https://github.com/MeLiDosProject/NilssonTengelinEtAl_Dataset_2026/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                         obj_name = "sleepdiary",
                         start_col = bedtime,
                         end_col = out_ofbed)
RISE <- remove_intervals(RISE,
                         url = "https://github.com/MeLiDosProject/NilssonTengelinEtAl_Dataset_2026/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                         obj_name = "wearlog",
                         start_col = start,
                         end_col = end)

THUAS <- remove_intervals(THUAS,
                          url = "https://github.com/MeLiDosProject/AertsEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                          obj_name = "sleepdiary",
                          start_col = bedtime,
                          end_col = out_ofbed)
THUAS <- remove_intervals(THUAS,
                          url = "https://github.com/MeLiDosProject/AertsEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                          obj_name = "wearlog",
                          start_col = start,
                          end_col = end)

TUM <- remove_intervals(TUM,
                        url = "https://github.com/MeLiDosProject/HildenEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                        obj_name = "sleepdiary",
                        start_col = bedtime,
                        end_col = out_ofbed)
TUM <- remove_intervals(TUM,
                        url = "https://github.com/MeLiDosProject/HildenEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                        obj_name = "wearlog",
                        start_col = start,
                        end_col = end)

UCR <- remove_intervals(UCR,
                        url = "https://github.com/MeLiDosProject/Sancho-SalasEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/sleepdiaries.RData",
                        obj_name = "sleepdiary",
                        start_col = bedtime,
                        end_col = out_ofbed)
UCR <- remove_intervals(UCR,
                        url = "https://github.com/MeLiDosProject/Sancho-SalasEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/wearlog.RData",
                        obj_name = "wearlog",
                        start_col = start,
                        end_col = end)


# Remove days with continuous zero signal for all three dosimeters
FUSPCEU <- FUSPCEU |>
  filter(!(Id == "FUSPCEU_S010" & as.Date(Datetime) == "2024-11-03"),
         !(Id == "FUSPCEU_S017" & as.Date(Datetime) == "2024-12-09"))

THUAS <- THUAS |> filter(!(Id == "THUAS_S007" & as.Date(Datetime) == "2025-04-11"))

IZTECH <- IZTECH |>
  filter(!(Id == "IZTECH_S009" & as.Date(Datetime) == "2025-03-18"),
         !(Id == "IZTECH_S009" & as.Date(Datetime) == "2025-03-19"))
```

## Calculate absolute and log transformed differences

```{r}
#| label: calculate-differences

calculate_difs <- function(data) {
  data |>
    mutate(
      absdif_c = MEDI_c - MEDI_h,
      absdif_w = MEDI_w - MEDI_h,
      logdif_c = log10(MEDI_c + 0.1) - log10(MEDI_h + 0.1),
      logdif_w = log10(MEDI_w + 0.1) - log10(MEDI_h + 0.1))
}

BAUA <- calculate_difs(BAUA)
FUSPCEU <- calculate_difs(FUSPCEU)
IZTECH <- calculate_difs(IZTECH)
KNUST <- calculate_difs(KNUST)
RISE <- calculate_difs(RISE)
THUAS <- calculate_difs(THUAS)
TUM <- calculate_difs(TUM)
UCR <- calculate_difs(UCR)
```

## Visualize light data

```{r}
#| label: visualize-light-data

vis_data <- function(data) {
  plot_data <- data |>
    pivot_longer(
      cols = starts_with("MEDI_"),
      names_to = "position",
      values_to = "MEDI"
    ) |>
    mutate(
      position = sub("^MEDI_", "", position),
      position = case_when(
        position == "h" ~ "head",
        position == "c" ~ "chest",
        position == "w" ~ "wrist"
  )
)
  
  gg_days(
    dataset = plot_data,
    aes_col = position,
    geom = "point", alpha = 0.6, size = 0.01
  )
}
```

```{r}
#| label: visualize-light-data2
#| results: asis

dataset_names <- c("BAUA", "FUSPCEU", "IZTECH", "KNUST", "RISE", "THUAS", "TUM", "UCR")

chunks <- lapply(dataset_names, \(name) {
  fig_height <- n_distinct(get(name)$Id) * 50 / 25.4
  knitr::knit_expand(text = c(
    "#### {{name}}",
    "```{r, fig.width=23, fig.height={{fig_height}}, fig.dpi=40, echo=FALSE}",
    "vis_data({{name}})",
    "```"
  ))
})

cat(knitr::knit(text = unlist(chunks), quiet = TRUE))
```

## Import contextual data

```{r}
#| label: import-contextual-data

# Add experiment day and time of day
add_day_time <- function(data) {
  data |>
    mutate(
      timeofday = factor(hour(Datetime)),
      day = factor(as.integer(as.Date(Datetime) - min(as.Date(Datetime))) + 1)
    )
}

BAUA <- add_day_time(BAUA)
FUSPCEU <- add_day_time(FUSPCEU)
IZTECH <- add_day_time(IZTECH)
KNUST <- add_day_time(KNUST)
THUAS <- add_day_time(THUAS)
TUM <- add_day_time(TUM)
UCR <- add_day_time(UCR)


# Add site
add_site <- function(data) {
  data |>
    mutate(site = factor(sub("_.*", "", Id)))
}

BAUA <- add_site(BAUA)
FUSPCEU <- add_site(FUSPCEU)
IZTECH <- add_site(IZTECH)
KNUST <- add_site(KNUST)
THUAS <- add_site(THUAS)
TUM <- add_site(TUM)
UCR <- add_site(UCR)


# Add photoperiod state
add_photoperiodstate <- function(data, coords) {
  add_photoperiod(data, coords) |>
    mutate(photoperiod.state = factor(photoperiod.state)) |>
    dplyr::select(-dawn, -dusk, -photoperiod)
}

BAUA <- add_photoperiodstate(BAUA, coordinates$BAUA)
FUSPCEU <- add_photoperiodstate(FUSPCEU, coordinates$FUSPCEU)
IZTECH <- add_photoperiodstate(IZTECH, coordinates$IZTECH)
KNUST <- add_photoperiodstate(KNUST, coordinates$KNUST)
THUAS <- add_photoperiodstate(THUAS, coordinates$THUAS)
TUM <- add_photoperiodstate(TUM, coordinates$TUM)
UCR <- add_photoperiodstate(UCR, coordinates$UCR)


# Import light exposure diaries
urls <- list(
  BAUA    = "https://github.com/MeLiDosProject/BroszioEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  FUSPCEU = "https://github.com/MeLiDosProject/BaezaEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  IZTECH  = "https://github.com/MeLiDosProject/DidikogluEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  KNUST   = "https://github.com/MeLiDosProject/AkuffoEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  THUAS   = "https://github.com/MeLiDosProject/AertsEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  TUM     = "https://github.com/MeLiDosProject/HildenEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  UCR     = "https://github.com/MeLiDosProject/Sancho-SalasEtAl_Dataset_2025/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData",
  RISE     = "https://github.com/MeLiDosProject/NilssonTengelinEtAl_Dataset_2026/raw/refs/heads/main/data/imported/continuous/lightexposurediary.RData"
)

# If start and end are set to NA, rows are excluded
load_diary <- function(url) {
  load(url(url))
  lightexposurediary |> filter(!is.na(start) & !is.na(end))
}
diaries <- lapply(urls, load_diary)

cols_to_add <- c(
  "lightsource_primary", "lightsource_secondary",
  "act_sleep", "act_home", "act_road_vehicle", "act_road_open",
  "act_working_indoor", "act_working_outdoor", "act_free_outdoor", "act_other"
)

# data.table is used, as fuzzy_left_join is very slow
add_diary <- function(data, diary_data) {
  light   <- as.data.table(data)
  diary <- as.data.table(diary_data)[, c("Id", "start", "end", cols_to_add), with = FALSE]
  
  light[, `:=`(start = Datetime, end = Datetime)]
  setkey(diary, Id, start, end)
  setkey(light, Id, start, end)
  
  result <- foverlaps(
    light, diary, by.x = c("Id", "start", "end"), by.y = c("Id", "start", "end"),
    type = "within", nomatch = NA)
  result[, c("start", "end", "i.start", "i.end") := NULL]
  setcolorder(result, c(setdiff(names(result), cols_to_add), cols_to_add))
  
  as_tibble(result) |> group_by(Id)
}

results <- mapply(add_diary, mget(c("BAUA", "FUSPCEU", "IZTECH", "KNUST", "THUAS", "TUM", "UCR", "RISE")), diaries, SIMPLIFY = FALSE)
list2env(results, envir = .GlobalEnv)
```

## Data export

```{r}
#| label: export-data

save_data <- function(data, output_dir) {
  study_name <- deparse(substitute(data))
  
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }
  
  save(list = study_name, envir = parent.frame(), file = file.path(output_dir, paste0(study_name, ".RData")))
}

output_dir <- here("processed_data")

save_data(BAUA, output_dir)
save_data(FUSPCEU, output_dir)
save_data(IZTECH, output_dir)
save_data(KNUST, output_dir)
save_data(RISE, output_dir)
save_data(THUAS, output_dir)
save_data(TUM, output_dir)
save_data(UCR, output_dir)
```

## References {.unnumbered}
