---
title: "Stability of wearing-position bias with sample size"
format: html
author: Johannes Zauner
params:
  draws: 10000
execute:
  freeze: false
editor_options:
  chunk_output_type: inline
---

## Preface

The metric comparison established that wearing position (chest and wrist,
relative to the glasses reference) introduces outcome-level bias in personal
light-exposure metrics. This notebook asks a complementary, design-oriented
question:

> **How does the precision of the estimated bias change as participants and
> participant-days accumulate?**

The analysis includes every metric with daily observations. Interdaily
stability (IS) and intradaily variability (IV) are excluded because they are
defined across multiple days and therefore have no participant-day estimates.
The remaining metrics are summarized by the six metric classes used in the
primary analysis: duration, exposure history, level, spectrum, temporal
dynamics, and timing.

Two bootstrap axes are evaluated for both chest and wrist:

1. **N participants:** sample 1–200 participants, with either **3 or 7 days per
   sampled participant**.
2. **Days per participant:** sample 1–7 days per participant in studies with
   either **25 or 50 participants**.

At each sample size, the standard deviation (SD) of the bootstrapped mean bias
quantifies sampling uncertainty. Bias is calculated on each metric's model
scale and normalized to the absolute mean glasses reference on that scale. The
normalized SDs are then summarized pointwise by their median within each metric
class. A 5% SD is used as a practical precision tolerance.

## Setup and inputs

```{r}
#| label: setup
#| warning: false
#| message: false
#| filename: Setup
library(tidyverse)
library(readxl)
library(LightLogR)
library(cowplot)
library(gt)
library(here)

if (packageVersion("LightLogR") < package_version("0.10.3")) {
  stop("This analysis requires LightLogR 0.10.3 or newer.", call. = FALSE)
}
```

The prepared long-format metric data and their model metadata are loaded from
the same sources used by the metric-comparison notebook. The reusable
bootstrap helpers are kept in a separate script so the sampling and aggregation
logic can be tested independently.

```{r}
#| label: import-inputs
#| filename: Import metrics, metadata, and helpers
load(here("data/prepared_metrics.RData"))
metric_info <- read_xlsx(here("data/metric_types.xlsx"))
source(here("scripts/bootstrap_stability.R"))
```

A single seed makes all bootstrap scenarios reproducible. The source notebook
defaults to 10,000 draws for production. The draw count is a parameter so a
render can be tested quickly with, for example,
`quarto render notebooks/_JZ_error_stabilisation.qmd --execute -P draws:100`.

```{r}
#| label: analysis-settings
#| filename: Analysis settings
production_draws <- 10000L
draws <- as.integer(params$draws)
if (length(draws) != 1L || is.na(draws) || draws < 2L) {
  stop("`params$draws` must be one integer greater than one.", call. = FALSE)
}

days_per_participant <- c(3L, 7L)
day_axis_participants <- c(25L, 50L)
tolerance <- 5

set.seed(7391)

position_colors <- ggsci::pal_jco()(3) |>
  set_names(c("glasses", "chest", "wrist"))

metric_class_labels <- c(
  duration = "Duration",
  `exposure history` = "Exposure history",
  level = "Level",
  spectrum = "Spectrum",
  `temporal dynamics` = "Temporal dynamics",
  timing = "Timing"
)
```

`r if (draws < production_draws) paste0("This is a computational test render with **B = ", format(draws, big.mark = ","), "** draws; it verifies execution and layout, while the default production render uses B = ", format(production_draws, big.mark = ","), ".") else paste0("This production render uses **B = ", format(draws, big.mark = ","), "** bootstrap draws.")`

## Daily metric set

The inclusion rule is data-driven: retain all metrics present in both the
prepared data and metadata, except IS and IV. Explicit checks prevent a silent
drop if the prepared data and metadata diverge.

```{r}
#| label: select-daily-metrics
#| filename: Define and verify daily metric set
excluded_metrics <- c("interdaily_stability", "intradaily_variability")
data_metrics <- sort(unique(metrics$metric))

missing_exclusions <- setdiff(excluded_metrics, data_metrics)
if (length(missing_exclusions) > 0L) {
  stop(
    "Expected non-daily metrics are absent from the prepared data: ",
    paste(missing_exclusions, collapse = ", "),
    ".",
    call. = FALSE
  )
}

missing_metadata <- setdiff(data_metrics, metric_info$name)
if (length(missing_metadata) > 0L) {
  stop(
    "Prepared metrics lack metadata: ",
    paste(missing_metadata, collapse = ", "),
    ".",
    call. = FALSE
  )
}

target_metrics <- setdiff(data_metrics, excluded_metrics)
target_info <- metric_info |>
  filter(name %in% target_metrics)

if (nrow(target_info) != length(target_metrics)) {
  stop("The daily metric selection is not one-to-one with its metadata.", call. = FALSE)
}

metric_class_counts <- target_info |>
  count(metric_type, name = "n_metrics") |>
  mutate(
    metric_class = factor(
      metric_type,
      levels = names(metric_class_labels),
      labels = unname(metric_class_labels)
    )
  ) |>
  arrange(metric_class)
```

The analysis contains `r length(target_metrics)` daily metrics. Class sizes are
shown because the class curves summarize different numbers of metric-specific
traces using the pointwise median.

```{r}
#| label: tbl-metric-class-counts
#| filename: Daily metric inventory by class
metric_class_counts |>
  select(`Metric class` = metric_class, Metrics = n_metrics) |>
  gt() |>
  tab_header(
    title = "Daily metrics included in the stability analysis",
    subtitle = "IS and IV excluded because participant-day estimates are unavailable"
  ) |>
  fmt_integer(columns = Metrics) |>
  cols_align("center", columns = Metrics) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels()
  ) |>
  tab_options(table.width = px(620))
```

## Model-scale wearing-position bias

For each metric, values are placed on the same scale used in the primary
metric models. Positions are then aligned within participant-day and the paired
bias is calculated for chest and wrist against glasses:

$$
\text{bias}_{i d p m} =
\text{value}^{\text{model}}_{i d p m} -
\text{value}^{\text{model}}_{i d,\,\text{glasses},\,m}.
$$

The bootstrap SD for metric $m$ is divided by the absolute mean model-scale
glasses value for that metric and multiplied by 100. This normalization makes
the metric-specific SD curves commensurable. The publication figure retains
every metric curve and overlays its metric-class median.

```{r}
#| label: compute-all-biases
#| filename: Compute paired biases for all daily metrics
scales <- derive_metric_scales(metric_info)

bias_df <- metrics |>
  filter(metric %in% target_metrics) |>
  compute_bias(scales)

if (n_distinct(bias_df$metric) != length(target_metrics)) {
  stop("Not every selected daily metric produced a paired bias.", call. = FALSE)
}

bias_coverage <- bias_df |>
  count(metric, position, name = "participant_days")

if (nrow(bias_coverage) != 2L * length(target_metrics)) {
  stop("Not every daily metric has paired chest and wrist biases.", call. = FALSE)
}

bias_coverage |>
  summarise(
    metrics = n_distinct(metric),
    positions = n_distinct(position),
    `paired participant-days (minimum)` = min(participant_days),
    `paired participant-days (maximum)` = max(participant_days)
  )
```

## Bootstrap design

Both axes use hierarchical resampling with replacement. For the participant
axis, each bootstrap replicate samples $N$ participants and then samples exactly
3 or 7 valid paired days from each selected participant's empirical daily-bias
distribution. For the participant-day axis, each replicate first samples 25 or
50 participants and then samples 1–7 days from each. Sampling days with
replacement retains participants with fewer observed valid days while making
the target design size explicit.

The participant axis extends to 200 as an empirical-distribution projection,
although the prepared cohort contains `r n_distinct(bias_df$Id)` participants.
It therefore describes expected sampling precision under repeated sampling from
this cohort's empirical distribution, not the acquisition of new independent
population information.

```{r}
#| label: full-run
#| filename: Bootstrap all daily metrics and four design scenarios
#| cache: true
#| cache.extra: !expr list(draws, days_per_participant, day_axis_participants)
metric_stability <- bootstrap_stability(
  bias_df,
  n_range = 1:200,
  day_range = 1:7,
  B = draws,
  days_per_participant = days_per_participant,
  day_axis_participants = day_axis_participants
)
```

## Metric-class medians

Class medians are overlaid on the individual metric traces in the publication
figure and are the sole basis of the crossing table. For each position, design
scenario, and sample size, the pointwise median is taken across all normalized
metric-specific SDs in that class. This robust summary limits the influence of
unusually unstable metrics while the faint traces retain their information.

```{r}
#| label: median-metric-classes
#| filename: Take median metric-specific SD within class
class_stability <- median_metric_classes(metric_stability)

observed_class_counts <- class_stability |>
  distinct(metric_class, n_metrics) |>
  arrange(metric_class)

expected_class_counts <- metric_class_counts |>
  transmute(metric_class = metric_type, n_metrics) |>
  arrange(metric_class)

if (!identical(observed_class_counts, expected_class_counts)) {
  stop("The class-median metric denominators do not match the inventory.", call. = FALSE)
}
```

## Stabilisation figure

The eight panels cross wearing position (rows) with the four design scenarios
(columns). Faint colored lines show every daily metric, while the thicker lines
show their pointwise metric-class medians. The dashed line marks the 5% SD
tolerance. In each leftmost facet, right-aligned direct labels and arrows
identify the three individual metrics with the largest SD at 200 participants.
The first two columns vary participants from 1 to 200 with 3 or 7 days each.
The final two vary days per participant from 1 to 7 with 25 or 50 participants.
To preserve detail around the tolerance, the display is clipped at 20%; larger
SDs remain in the analysis.

```{r}
#| label: fig-stability
#| filename: Metric-class stability by design scenario
#| fig-width: 13.5
#| fig-height: 7.5
#| warning: false
fig_stability <- plot_stability(
  metric_stability,
  class_stability,
  position_colors = position_colors,
  metric_class_labels = metric_class_labels,
  tol = tolerance,
  ylim = c(0, 20),
  draws = draws
)

fig_stability
```

```{r}
#| label: save-figure
#| filename: Save publication figure
ggsave(
  here("output/figures/Fig_bias_stability.pdf"),
  fig_stability,
  width = 13.5,
  height = 7.5,
  device = cairo_pdf,
  bg = "white"
)
ggsave(
  here("output/figures/Fig_bias_stability.jpeg"),
  fig_stability,
  width = 13.5,
  height = 7.5,
  dpi = 600,
  bg = "white"
)
```

## Sample size at the 5% tolerance

For every class, position, and design scenario, the table reports the first
sample size at which the **class-median** SD is at or below 5%. A dash means the
tolerance is not reached within 1–200 participants or 1–7 days per participant.
The crossing applies to the class median and does not imply that every
individual metric in that class has crossed the tolerance.

```{r}
#| label: tbl-crossing
#| filename: Class-median sample size to reach 5 percent SD
crossing_wide <- crossing_table(class_stability, tol = tolerance) |>
  mutate(
    metric_class = factor(
      metric_class,
      levels = names(metric_class_labels),
      labels = unname(metric_class_labels)
    ),
    scenario_code = case_when(
      boot_type == "participants" & days_per_participant == 3L ~ "p3",
      boot_type == "participants" & days_per_participant == 7L ~ "p7",
      boot_type == "participant-days" & n_participants == 25L ~ "d25",
      boot_type == "participant-days" & n_participants == 50L ~ "d50",
      .default = NA_character_
    ),
    position = str_to_lower(position),
    output_column = paste(scenario_code, position, sep = "_")
  ) |>
  select(metric_class, n_metrics, output_column, n_within) |>
  pivot_wider(names_from = output_column, values_from = n_within) |>
  arrange(metric_class)

crossing_columns <- c(
  "p3_chest", "p3_wrist",
  "p7_chest", "p7_wrist",
  "d25_chest", "d25_wrist",
  "d50_chest", "d50_wrist"
)

if (!all(crossing_columns %in% names(crossing_wide))) {
  stop("The crossing table is missing at least one design-position column.", call. = FALSE)
}

crossing_wide <- crossing_wide |>
  select(metric_class, n_metrics, all_of(crossing_columns))

crossing_gt <- crossing_wide |>
  gt(rowname_col = "metric_class") |>
  tab_stubhead(label = "Metric class") |>
  cols_label(
    n_metrics = "Metrics",
    p3_chest = "Chest",
    p3_wrist = "Wrist",
    p7_chest = "Chest",
    p7_wrist = "Wrist",
    d25_chest = "Chest",
    d25_wrist = "Wrist",
    d50_chest = "Chest",
    d50_wrist = "Wrist"
  ) |>
  tab_spanner(
    label = "3 days each",
    columns = c(p3_chest, p3_wrist),
    level = 1,
    id = "participants_3",
    gather = FALSE
  ) |>
  tab_spanner(
    label = "7 days each",
    columns = c(p7_chest, p7_wrist),
    level = 1,
    id = "participants_7",
    gather = FALSE
  ) |>
  tab_spanner(
    label = "25 participants",
    columns = c(d25_chest, d25_wrist),
    level = 1,
    id = "days_25",
    gather = FALSE
  ) |>
  tab_spanner(
    label = "50 participants",
    columns = c(d50_chest, d50_wrist),
    level = 1,
    id = "days_50",
    gather = FALSE
  ) |>
  tab_spanner(
    label = "N participants",
    spanners = c("participants_3", "participants_7"),
    level = 2,
    id = "participant_axis"
  ) |>
  tab_spanner(
    label = "Days per participant",
    spanners = c("days_25", "days_50"),
    level = 2,
    id = "day_axis"
  ) |>
  fmt_integer(columns = c(n_metrics, all_of(crossing_columns))) |>
  sub_missing(columns = all_of(crossing_columns), missing_text = "—") |>
  cols_align("center", columns = c(n_metrics, all_of(crossing_columns))) |>
  cols_width(
    n_metrics ~ px(65),
    all_of(crossing_columns) ~ px(82)
  ) |>
  tab_header(
    title = "Sample size for class-median bias SD ≤ 5%",
    subtitle = paste0(
      length(target_metrics),
      " daily metrics; pointwise median within class; B = ",
      format(draws, big.mark = ","),
      " bootstrap draws"
    )
  ) |>
  tab_source_note(
    md(
      paste0(
        "**Note.** Cells give the first N participants (left) or days per ",
        "participant (right) meeting the tolerance. — indicates no crossing ",
        "within the evaluated range. A class-median crossing does not imply ",
        "that every metric in the class meets the tolerance."
      )
    )
  ) |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(cells_stub(), cells_column_labels(), cells_column_spanners())
  ) |>
  tab_style(
    style = cell_text(color = unname(position_colors[["chest"]]), weight = "bold"),
    locations = cells_column_labels(columns = ends_with("_chest"))
  ) |>
  tab_style(
    style = cell_text(color = unname(position_colors[["wrist"]]), weight = "bold"),
    locations = cells_column_labels(columns = ends_with("_wrist"))
  ) |>
  tab_style(
    style = cell_fill(color = unname(position_colors[["chest"]]), alpha = 0.08),
    locations = cells_body(columns = ends_with("_chest"))
  ) |>
  tab_style(
    style = cell_fill(color = unname(position_colors[["wrist"]]), alpha = 0.08),
    locations = cells_body(columns = ends_with("_wrist"))
  ) |>
  tab_options(
    table.width = px(1000),
    table.font.size = px(14),
    heading.align = "left",
    column_labels.background.color = "#F3F3F3",
    table_body.hlines.color = "#D9D9D9",
    source_notes.font.size = px(11)
  )

crossing_gt
```

```{r}
#| label: save-crossing-table
#| filename: Save publication crossing table
gtsave(
  crossing_gt,
  here("output/tables/Table_crossing.png"),
  vwidth = 1500,
  zoom = 2
)
```

## Interpretation

The figure isolates precision rather than bias magnitude. Increasing the number
of participants reduces uncertainty in the estimated population mean, whereas
adding days primarily reduces within-participant sampling uncertainty and is
ultimately limited by between-participant heterogeneity. The paired 3- versus
7-day participant scenarios show how much of the participant-axis precision is
attributable to longer monitoring. Likewise, the 25- versus 50-participant day
scenarios show how study size changes the value of additional days.

The thick class medians provide a concise design summary across all daily
metrics, while the faint individual traces expose metric-level heterogeneity.
The class medians should therefore be used to compare broad metric classes, not
to choose a sample size for a single specific metric. The direct labels in the
leftmost facets identify the three metrics with the largest SD at the maximum
participant sample size; they are not a separate inferential comparison. A
small SD indicates that the placement bias is estimated precisely; it does not
indicate that the placement bias itself is small.

## Session info

```{r}
#| label: session-info
sessionInfo()
```
