---
title: "Comparison of metrics - wake-only sensitivity"
subtitle: "Sensitivity analysis: documented non-wear and diary-defined sleep removed"
format: html
author: Johannes Zauner
editor_options:
  chunk_output_type: inline
---

## Preface

This document repeats the primary metric-comparison workflow using the wake-only dataset prepared in `_JZ_metric_preparation_wake_only.qmd`. Participant-day eligibility is inherited from the primary analysis, but documented non-wear and diary-defined sleep are removed before the 42 retained metrics are calculated.

The primary-analysis sample flowchart is not repeated because it describes a different state-handling scenario. Sample verification and the wake-only state audit are reported in the corresponding preparation document.

## Setup

```{r}
#| label: setup
#| warning: false
#| message: false
library(tidyverse)
library(gt)
library(see)
library(readxl)
library(gghighlight)
library(cowplot)
library(glmmTMB)
library(lme4)
library(lmerTest)
library(melidosData)
library(emmeans)
library(here)
library(car)
library(performance)
library(LightLogR)
library(broom)
library(broom.mixed)
library(katex)
library(gtsummary)
library(stringr)
library(knitr)
library(rmarkdown)
library(fontawesome)
library(patchwork)
```

## Import and verify metrics

```{r}
#| label: import-metrics
load(here("data/prepared_metrics_wake_only.RData"))
```

```{r}
#| label: verify-imported-metrics
#| filename: Verify wake-only metric set
excluded_metrics <- c(
  "interdaily_stability",
  "intradaily_variability",
  "nvRD",
  "darkest_10h_mean",
  "darkest_10h_midpoint",
  "darkest_10h_onset",
  "darkest_10h_offset",
  "dark_threshold",
  "dark_mean_level",
  "dark_cluster",
  "circadian_variation",
  "disparity_index"
)

if (n_distinct(metrics$metric) != 42L) {
  stop("The wake-only comparison requires exactly 42 metrics.", call. = FALSE)
}
if (any(excluded_metrics %in% metrics$metric)) {
  stop("The wake-only data contain at least one excluded metric.", call. = FALSE)
}
```

### Derive wake and sleep duration denominators

Timing differences are expressed relative to the behavioural interval in which the underlying metric is defined. The sleep interval follows the preparation workflow and extends from diary-recorded `sleepprep` (when the devices were positioned for the night) to `wake`. Diary intervals are restricted to participant-dates retained in the prepared metrics, with each interval assigned to its wake date. Mean sleep duration is the grand mean across all retained diary intervals, so each retained night receives equal weight. Mean wake duration is the complement of mean sleep duration to 24 hours. The three darkest-10-hour timing outcomes are classified as darkness-based; all other timing outcomes are classified as brightness-based. The darkness-based timing outcomes are not retained in this wake-only sensitivity analysis, but the sleep-duration denominator is still reported for consistency across scenarios.

```{r}
#| label: derive-state-duration-denominators
#| filename: Derive grand-mean wake and sleep durations
#| message: false
#| warning: false
analysis_days <-
  metrics |>
  filter(!is.na(Date)) |>
  transmute(
    Id = as.character(Id),
    Date = as.Date(Date)
  ) |>
  distinct()

sleep_intervals <-
  load_data("sleepdiaries") |>
  list_rbind(names_to = "diary_site") |>
  filter(
    !is.na(sleepprep),
    !is.na(wake)
  ) |>
  mutate(
    Id = as.character(Id),
    Date = as.Date(wake),
    sleep_duration_seconds = as.numeric(
      difftime(wake, sleepprep, units = "secs")
    )
  ) |>
  semi_join(analysis_days, by = join_by(Id, Date))

invalid_sleep_intervals <-
  sleep_intervals |>
  filter(
    !is.finite(sleep_duration_seconds) |
      sleep_duration_seconds <= 0 |
      sleep_duration_seconds >= 24 * 60 * 60
  )

if (nrow(invalid_sleep_intervals) > 0L) {
  stop(
    "At least one diary-defined sleep interval is outside (0, 24) hours.",
    call. = FALSE
  )
}

mean_sleep_duration_seconds <- mean(sleep_intervals$sleep_duration_seconds)
mean_wake_duration_seconds <- 24 * 60 * 60 - mean_sleep_duration_seconds

format_duration <- function(seconds) {
  total_minutes <- round(seconds / 60)
  sprintf(
    "%d h %02d min",
    total_minutes %/% 60,
    total_minutes %% 60
  )
}

mean_sleep_duration_label <- format_duration(mean_sleep_duration_seconds)
mean_wake_duration_label <- format_duration(mean_wake_duration_seconds)
mean_sleep_one_percent_label <-
  sprintf("%.1f min", mean_sleep_duration_seconds / 100 / 60)
mean_wake_one_percent_label <-
  sprintf("%.1f min", mean_wake_duration_seconds / 100 / 60)

darkness_timing_metrics <- c(
  "darkest_10h_midpoint",
  "darkest_10h_onset",
  "darkest_10h_offset"
)

timing_normalization_footnote <- paste0(
  "Timing effects and timing-model SDs use the grand mean ",
  "diary-defined wake duration (", mean_wake_duration_label,
  "; 1% = ", mean_wake_one_percent_label,
  ") for brightness-based timing metrics and the grand mean ",
  "diary-defined sleep duration (", mean_sleep_duration_label,
  "; 1% = ", mean_sleep_one_percent_label,
  ") for darkness-based timing metrics. All other metrics use the absolute ",
  "glasses reference on the modelling scale."
)
```

The resulting grand means are `r mean_wake_duration_label` of wake and `r mean_sleep_duration_label` of sleep. Consequently, 1% represents `r mean_wake_one_percent_label` for brightness-based timing metrics and `r mean_sleep_one_percent_label` for darkness-based timing metrics.

## Visualize metrics

```{r}
#| label: plot-helpers
#| filename: Plot helpers
breaks <- c(0, 10^(0:5))
colors <- ggsci::pal_jco()(3) |> set_names(c("glasses", "chest", "wrist"))
scenario_label <- "Sensitivity analysis: documented non-wear and diary-defined sleep removed"
label <- function(x) {
  format(x, scientific = FALSE, big.mark = " ")
}
```

All metrics are centered and scaled within participant-metric groups for this distribution overview. The figure describes the wake-only metric set and does not change the values used for modelling.

```{r}
#| label: fig-metric-distribution
#| fig-width: 15
#| fig-height: 20
#| warning: false
#| filename: Show distribution of wake-only metrics
metrics |>
  group_by(Id, metric) |>
  mutate(
    position = fct_relevel(position, c("chest", "glasses", "wrist")),
    value = base::scale(value)
  ) |>
  ggplot(aes(x = position, y = value)) +
  geom_line(aes(group = interaction(Id, Date)), alpha = 0.05) +
  geom_jitter(aes(col = position), height = 0, width = 0.1, alpha = 0.5) +
  geom_boxplot(
    aes(col = position, fill = position),
    alpha = 0.5,
    width = 0.1,
    position = position_nudge(x = 0.2)
  ) +
  geom_violinhalf(
    aes(x = position, y = value, col = position, fill = position),
    position = position_nudge(x = 0.3),
    inherit.aes = FALSE,
    alpha = 0.5
  ) +
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  facet_wrap(~metric, ncol = 5) +
  guides(col = "none", fill = "none") +
  labs(
    y = "Scaled and centered metric values",
    x = "Wearing position",
    title = "Raincloud plots of scaled and centered metric results",
    subtitle = scenario_label
  ) +
  theme_cowplot()

ggsave(
  here("output/figures/Figure_metric_distribution_wake_only.pdf"),
  width = 15, height = 20, device = cairo_pdf
)
ggsave(
  here("output/figures/Figure_metric_distribution_wake_only.jpeg"),
  width = 15, height = 20
)
```

## Prepare modelling

The data are nested by metric and joined to the existing model metadata. The darkest-period timing metrics that required recentering around midnight are not part of this sensitivity set, so no nighttime recentering is applied.

```{r}
#| label: nest-metrics
#| filename: Nest metric data
metrics_nested <-
  metrics |>
  mutate(
    Date = replace_na(Date, as_date("2024-09-01")),
    position = factor(position, levels = c("glasses", "chest", "wrist")),
    Id = factor(Id),
    site = factor(site),
    Date = factor(Date)
  ) |>
  nest(data = -c(metric))
```

```{r}
#| label: load-metric-info
#| filename: Load metric information
metric_info <- read_xlsx(here("data/metric_types.xlsx"))
metric_info <-
  metric_info |>
  mutate(
    family = map(
      engine,
      \(x) switch(
        x,
        "lm" = gaussian(),
        "lmer" = gaussian(),
        "glmer" = poisson(),
        "glmmTMB" = tweedie(link = "log")
      )
    ),
    family = map2(
      family,
      note,
      \(x, y) ifelse(is.na(y), return(x), return(poisson()))
    )
  ) |>
  select(-note)
```

```{r}
#| label: join-metric-metadata
#| filename: Combine metric data and metadata
metrics_nested <-
  metrics_nested |>
  left_join(metric_info, by = join_by(metric == name))

if (any(is.na(metrics_nested$engine)) || any(is.na(metrics_nested$metric_type))) {
  stop("At least one wake-only metric has no complete model metadata.", call. = FALSE)
}
```

The primary hypothesis is that wearing position affects a metric and may interact with study site. The stored random-effects specification is appended to the fixed-effects template.

```{r}
#| label: add-formulas
#| filename: Add formulas
metrics_nested <-
  metrics_nested |>
  mutate(H1 = paste0("~ site*position", random))
```

```{r}
#| label: collect-formula-names
#| filename: Collect model formula columns
Formulas <- names(metrics_nested) |>
  subset(grepl("H1|H0_", names(metrics_nested)))
```

```{r}
#| label: tbl-model-specs
#| echo: false
metrics_nested |>
  select(-data, -random) |>
  arrange(metric_type, metric_family, metric) |>
  mutate(
    family = map(family, 1),
    rowname = seq_along(metric)
  ) |>
  group_by(metric_type) |>
  gt() |>
  cols_label_with(fn = \(x) x |> str_to_title() |> str_replace_all("_", " ")) |>
  sub_missing(missing_text = "") |>
  sub_values(values = "NULL", replacement = "") |>
  fmt(
    c(metric, pre.transformed, response),
    fns = \(x) x |> str_to_title() |> str_replace_all("_", " ")
  ) |>
  sub_values(values = "MDER", replacement = "MDER") |>
  tab_header(
    "Metric model specifications",
    subtitle = paste(
      "Hypothesis: wearing position affects metric result within sites.",
      scenario_label
    )
  ) |>
  tab_footnote(
    "No p-value correction is applied, matching the primary analysis."
  ) |>
  tab_footnote(
    "H1 is tested with car::Anova to inspect the model terms.",
    locations = cells_column_labels(Formulas)
  ) |>
  cols_width(
    Formulas ~ px(200),
    c(metric, pre.transformed, response) ~ px(125),
    everything() ~ px(80)
  )
```

```{r}
#| label: create-formulas
#| filename: Formula creation
make_formula <- function(response, rhs) {
  if (is.na(rhs) || is.na(response)) {
    return(NULL)
  }
  as.formula(paste(response, rhs))
}

metrics_nested <-
  metrics_nested |>
  mutate(
    response = str_replace(response, "metric", "value"),
    across(
      all_of(Formulas),
      \(x) map2(x, response, \(rhs, response) make_formula(response, rhs))
    )
  )
```

## Fit models

The same engines and stepwise complexity reduction used by the primary analysis are applied. If required, the date random effect is removed first, then the participant-specific position slope is reduced to an intercept, and finally the site-position interaction is removed.

```{r}
#| label: fit-model-helpers
#| filename: Model fitting helpers
main_formula <- value ~ site * position + site + position + (position | Id) + (1 | Date)
red_formula <- value ~ site * position + site + position + (position | Id)
red_formula2 <- value ~ site * position + site + position + (1 | Id)

rhs_same <- function(f1, f2) {
  identical(
    attr(terms(formula(f1, rhs = 1)), "term.labels"),
    attr(terms(formula(f2, rhs = 1)), "term.labels")
  )
}

fit_model <- function(data, formula, engine, family) {
  if (is.null(formula)) {
    return(NULL)
  }

  model <- switch(
    engine,
    "lm" = lm(formula, data = data, contrasts = list(site = "contr.sum")),
    "glm" = glm(
      formula, data = data, family = family,
      contrasts = list(site = "contr.sum")
    ),
    "lmer" = lmer(formula, data = data, contrasts = list(site = "contr.sum")),
    "glmer" = glmer(
      formula, data = data, family = family,
      contrasts = list(site = "contr.sum")
    ),
    "glmmTMB" = glmmTMB(
      formula, data = data, family = family,
      contrasts = list(site = "contr.sum")
    )
  )

  did_converge <-
    performance::check_convergence(model) && !check_singularity(model)

  if (!did_converge) {
    if (rhs_same(main_formula, formula)) {
      formula <- update(formula, . ~ . - (1 | Date))
    } else if (rhs_same(red_formula, formula)) {
      formula <- update(formula, . ~ . - (position | Id) + (1 | Id))
    } else if (rhs_same(red_formula2, formula)) {
      formula <- update(formula, . ~ . - site:position)
    } else {
      return(model)
    }
    model <- fit_model(data, formula, engine, family)
  }

  model
}
```

```{r}
#| label: fit-all-models
#| filename: Fit all models
#| message: false
#| warning: false
metrics_nested <-
  metrics_nested |>
  mutate(
    H1_model = pmap(
      list(data, H1, engine, family), fit_model,
      .progress = "H1 model fitting"
    ),
    converged = map_lgl(H1_model, check_convergence),
    singular = map_lgl(H1_model, check_singularity),
    H1_fin = map(H1_model, formula),
    H1_no_position = map(
      H1_fin,
      \(x) update(x, . ~ . - site:position - position - (position | Id) + (1 | Id))
    ),
    H0 = map(H1_fin, \(x) update(x, . ~ . - site * position)),
    H1_no_date = map(H1_fin, \(x) update(x, . ~ . - (1 | Date))),
    H0_no_date = map(H1_no_date, \(x) update(x, . ~ . - site * position)),
    H1_no_slope = map(
      H1_fin,
      \(x) update(x, . ~ . - (position | Id) + (1 | Id))
    ),
    H0_no_slope = map(H1_no_slope, \(x) update(x, . ~ . - site * position)),
    H0_model = pmap(
      list(data, H0, engine, family), fit_model,
      .progress = "H0 model fitting"
    ),
    H1_no_position_model = pmap(
      list(data, H1_no_position, engine, family), fit_model,
      .progress = "H1 no position model fitting"
    ),
    H1_no_date_model = pmap(
      list(data, H1_no_date, engine, family), fit_model,
      .progress = "H1 no date model fitting"
    ),
    H0_no_date_model = pmap(
      list(data, H0_no_date, engine, family), fit_model,
      .progress = "H0 no date model fitting"
    ),
    H1_no_slope_model = pmap(
      list(data, H1_no_slope, engine, family), fit_model,
      .progress = "H1 no position slope model fitting"
    ),
    H0_no_slope_model = pmap(
      list(data, H0_no_slope, engine, family), fit_model,
      .progress = "H0 no position slope model fitting"
    )
  )
```

```{r}
#| label: check-convergence
#| filename: Check singularity and convergence
all(metrics_nested$converged)
any(metrics_nested$singular)

if (!all(metrics_nested$converged) || any(metrics_nested$singular)) {
  stop("At least one wake-only H1 model failed convergence or remained singular.", call. = FALSE)
}
```

## Extract model results

P-values, diagnostics, random-effect SDs, reference and position contrasts, site contrasts, and partial $$R^2$$ components are extracted exactly as in the primary analysis.

```{r}
#| label: extract-model-results
#| filename: Extract model results
#| message: false
sig.level <- 0.05
metrics_nested <-
  metrics_nested |>
  mutate(
    pvalues = map(H1_model, \(x) Anova(x) |> tidy()),
    diagnostics = map(H1_model, check_model),
    random_effects = map(
      H1_model,
      \(x) x |>
        tidy() |>
        filter(effect == "ran_pars") |>
        select(term, group, estimate)
    ),
    main_effect = map(
      H1_model,
      \(x) x |>
        emmeans(specs = ~1, at = list(position = "glasses")) |>
        as.tibble()
    ),
    position_effect = map(
      H1_model,
      \(x) x |>
        emmeans(specs = ~position) |>
        contrast(method = "trt.vs.ctrl") |>
        as.tibble()
    ),
    site_effect = map(
      H1_model,
      \(x) x |>
        emmeans(specs = ~site | position) |>
        contrast() |>
        as.tibble()
    ),
    r2_1 = map2(H1_model, H0_model, \(x, y) r2_nakagawa(x, null_model = y)),
    r2_no_date = map2(
      H1_no_date_model, H0_no_date_model,
      \(x, y) r2_nakagawa(x, null_model = y)
    ),
    r2_no_slope = map2(
      H1_no_slope_model, H0_no_slope_model,
      \(x, y) r2_nakagawa(x, null_model = y)
    ),
    r2_no_position = map2(
      H1_no_position_model, H0_no_slope_model,
      \(x, y) r2_nakagawa(x, null_model = y)
    ),
    r2_residual = map_dbl(r2_1, \(x) 1 - x$R2_conditional),
    r2_random = map_dbl(r2_1, \(x) x$R2_conditional - x$R2_marginal),
    r2_fixed = map_dbl(r2_1, \(x) x$R2_marginal),
    r2_position = map2_dbl(
      r2_1, r2_no_position,
      \(x, y) x$R2_marginal - y$R2_marginal
    ),
    r2_slope = map2_dbl(
      r2_1, r2_no_slope,
      \(x, y) (x$R2_conditional - x$R2_marginal) -
        (y$R2_conditional - y$R2_marginal)
    ),
    r2_date = map2_dbl(
      r2_1, r2_no_date,
      \(x, y) (x$R2_conditional - x$R2_marginal) -
        (y$R2_conditional - y$R2_marginal)
    ),
    r2_id = r2_random - r2_slope - r2_date,
    r2_site = r2_fixed - r2_position
  )
```

## Result table

For non-timing metrics, placement effects and random-effect SDs are divided by the absolute glasses reference on the modelling scale. Clock time has an arbitrary zero point, so an intercept-relative percentage such as a percentage of 16:00 is not meaningful. Brightness-based timing metrics are therefore divided by the grand mean diary-defined wake duration, whereas darkness-based timing metrics would be divided by the grand mean diary-defined sleep duration. The latter outcomes are excluded from this wake-only metric set.

```{r}
#| label: collect-result-table
#| filename: Collect result-table information
metric_table <-
  metrics_nested |>
  mutate(
    scaling = case_when(
      pre.transformed == "log_zero_inflated" ~ "log10",
      engine == "glmmTMB" ~ "log",
      response == "log_zero_inflated(value)" ~ "log10",
      response == "qlogis(value)" ~ "logit",
      map_chr(family, 1) == "poisson" ~ "log",
      .default = "linear"
    ),
    pvalues = map(
      pvalues,
      \(x) x |>
        select(term, p.value) |>
        pivot_wider(
          names_from = term,
          values_from = p.value,
          names_prefix = "p_"
        )
    ),
    random_effects = map(
      random_effects,
      \(x) x |>
        mutate(
          group = replace_when(
            group,
            term == "sd__positionchest" ~ "chest",
            term == "sd__positionwrist" ~ "wrist"
          )
        ) |>
        filter_out(str_detect(term, "cor__")) |>
        select(-term) |>
        pivot_wider(
          names_from = group,
          values_from = estimate,
          names_prefix = "SD_"
        )
    ),
    main_effect = map_dbl(main_effect, 2),
    position_effect = map(
      position_effect,
      \(x) x |>
        select(contrast, estimate, p = p.value) |>
        pivot_wider(
          names_from = contrast,
          values_from = c(estimate, p)
        )
    ),
    site_effect = map(
      site_effect,
      \(x) x |>
        mutate(contrast = str_remove(contrast, " effect")) |>
        unite(name, contrast, position) |>
        select(name, estimate, p = p.value) |>
        pivot_wider(names_from = name, values_from = c(estimate, p))
    )
  ) |>
  select(
    metric, metric_family, metric_type, scaling, pvalues,
    random_effects, main_effect, position_effect, site_effect,
    r2_position, r2_site, r2_id, r2_date, r2_slope, r2_residual
  ) |>
  unnest(c(pvalues, random_effects, position_effect, site_effect))
```

```{r}
#| label: tbl-main-results
#| filename: Construct detailed wake-only result table
format_metric <- function(x) {
  case_when(
    x %in% c("MDER", "nvRD") ~ x,
    x == "centroidLE" ~ "Centroid LE",
    x == "midpointCE" ~ "Midpoint CE",
    x == "mean_MEDI" ~ "Mean melEDI",
    TRUE ~ x |>
      str_replace_all("_", " ") |>
      str_to_sentence() |>
      str_replace_all("(\\d+)(?!\\d|h)", "\\1 lx melEDI")
  )
}

sites <- c("BAUA", "FUSPCEU", "IZTECH", "KNUST", "RISE", "THUAS", "TUM", "UCR")
site_colors <- melidosData::melidos_colors[sites]

make_arrows_html <- function(estimates, pvals, colors, sig_level) {
  sig <- !is.na(pvals) & pvals < sig_level
  if (!any(sig)) {
    return("")
  }
  direction <- ifelse(estimates[sig] > 0, "arrow-up", "arrow-down")
  arrow_colors <- colors[sig]
  paste(
    mapply(
      \(direction, color) as.character(
        fa(direction, fill = color, height = "1em")
      ),
      direction,
      arrow_colors
    ),
    collapse = ""
  )
}

if (any(darkness_timing_metrics %in% metric_table$metric)) {
  stop(
    "The wake-only analysis unexpectedly contains a darkness-based timing metric.",
    call. = FALSE
  )
}

df <-
  metric_table |>
  mutate(
    is_timing_metric = metric_type == "timing",
    is_darkness_timing = metric %in% darkness_timing_metrics,
    normalization_denominator = case_when(
      is_darkness_timing ~ mean_sleep_duration_seconds,
      is_timing_metric ~ mean_wake_duration_seconds,
      .default = abs(main_effect)
    ),
    normalization_basis = case_when(
      is_darkness_timing ~ paste0(
        "Mean sleep duration (", mean_sleep_duration_label, ")"
      ),
      is_timing_metric ~ paste0(
        "Mean wake duration (", mean_wake_duration_label, ")"
      ),
      .default = "Absolute glasses reference"
    ),
    raw_estimate_chest = `estimate_chest - glasses`,
    raw_estimate_wrist = `estimate_wrist - glasses`,
    across(c(metric, metric_family, metric_type), format_metric)
  ) |>
  arrange(metric_type, metric, metric_family) |>
  rowwise() |>
  mutate(
    arrow_main = if (!is.na(p_site) && p_site < sig.level) {
      make_arrows_html(
        c_across(all_of(paste0("estimate_", sites, "_glasses"))),
        c_across(all_of(paste0("p_", sites, "_glasses"))),
        site_colors,
        sig.level
      )
    } else {
      ""
    },
    arrow_chest = if (
      !is.na(`p_site:position`) && `p_site:position` < sig.level
    ) {
      make_arrows_html(
        c_across(all_of(paste0("estimate_", sites, "_chest"))),
        c_across(all_of(paste0("p_", sites, "_chest"))),
        site_colors,
        sig.level
      )
    } else {
      ""
    },
    arrow_wrist = if (
      !is.na(`p_site:position`) && `p_site:position` < sig.level
    ) {
      make_arrows_html(
        c_across(all_of(paste0("estimate_", sites, "_wrist"))),
        c_across(all_of(paste0("p_", sites, "_wrist"))),
        site_colors,
        sig.level
      )
    } else {
      ""
    },
    across(
      c(
        `estimate_chest - glasses`,
        `estimate_wrist - glasses`,
        starts_with("SD_")
      ),
      \(x) x / normalization_denominator
    ),
    across(starts_with("r2_"), \(x) replace_values(x, 0 ~ NA))
  ) |>
  ungroup()

expected_normalization_denominator <- case_when(
  df$is_darkness_timing ~ mean_sleep_duration_seconds,
  df$is_timing_metric ~ mean_wake_duration_seconds,
  .default = abs(df$main_effect)
)

if (!isTRUE(all.equal(
  df$normalization_denominator,
  expected_normalization_denominator,
  check.attributes = FALSE
))) {
  stop("Wake-only normalization denominators are incorrect.", call. = FALSE)
}

if (!isTRUE(all.equal(
  df$`estimate_chest - glasses`,
  df$raw_estimate_chest / df$normalization_denominator,
  check.attributes = FALSE
)) || !isTRUE(all.equal(
  df$`estimate_wrist - glasses`,
  df$raw_estimate_wrist / df$normalization_denominator,
  check.attributes = FALSE
))) {
  stop("Wake-only bias normalization failed its numerical check.", call. = FALSE)
}

table <-
  df |>
  group_by(metric_type) |>
  gt(rowname_col = c("metric_family", "metric")) |>
  cols_width(
    metric_family ~ px(85),
    metric ~ px(340),
    main_effect ~ px(120),
    arrow_main ~ px(60),
    `estimate_chest - glasses` ~ px(58),
    arrow_chest ~ px(60),
    `estimate_wrist - glasses` ~ px(55),
    SD_Residual ~ px(70),
    arrow_wrist ~ px(60),
    scaling ~ px(55),
    starts_with("p_") ~ px(65),
    starts_with("r2_") ~ px(65),
    everything() ~ px(60)
  ) |>
  cols_move(main_effect, after = scaling) |>
  cols_move(arrow_main, after = main_effect) |>
  cols_move(`estimate_chest - glasses`, after = arrow_main) |>
  cols_move(arrow_chest, after = `estimate_chest - glasses`) |>
  cols_move(`estimate_wrist - glasses`, after = arrow_chest) |>
  cols_move(arrow_wrist, after = `estimate_wrist - glasses`) |>
  cols_move(
    c(
      p_position, r2_position, SD_chest, SD_wrist, r2_slope, SD_Id,
      r2_id, SD_Date, r2_date, SD_Residual, r2_residual
    ),
    after = arrow_wrist
  ) |>
  fmt_scientific(main_effect) |>
  fmt_percent(
    columns = c(`estimate_chest - glasses`, `estimate_wrist - glasses`),
    decimals = 0,
    force_sign = TRUE
  ) |>
  fmt_percent(starts_with("r2_"), decimals = 0) |>
  fmt_percent(starts_with("SD_"), decimals = 0, pattern = "±{x}") |>
  fmt(starts_with("p_"), fns = style_pvalue) |>
  cols_label(
    main_effect = "Reference",
    arrow_main = "",
    p_position = html("p<sub>Pos.</sub>"),
    p_site = html("p<sub>Site</sub>"),
    `p_site:position` = html("p<sub>Inter.</sub>"),
    r2_position = html("R<sup>2</sup><sub>Pos.</sub>"),
    r2_slope = html("R<sup>2</sup><sub>Slope</sub>"),
    r2_id = html("R<sup>2</sup><sub>Id</sub>"),
    r2_date = html("R<sup>2</sup><sub>Date</sub>"),
    r2_residual = html("R<sup>2</sup><sub>Res.</sub>"),
    r2_site = html("R<sup>2</sup><sub>Site</sub>"),
    scaling = "Scale",
    `estimate_chest - glasses` = "Chest",
    arrow_chest = "",
    `estimate_wrist - glasses` = "Wrist",
    arrow_wrist = ""
  ) |>
  cols_label_with(
    starts_with("SD_"),
    fn = \(x) x |> str_remove("SD_") |> str_to_title()
  ) |>
  tab_style(
    style = cell_fill(color = "#FFDDDD60"),
    locations = list(
      cells_body(
        columns = `estimate_chest - glasses`,
        rows = `estimate_chest - glasses` < 0 &
          `p_chest - glasses` < sig.level
      ),
      cells_body(
        columns = `estimate_wrist - glasses`,
        rows = `estimate_wrist - glasses` < 0 &
          `p_wrist - glasses` < sig.level
      )
    )
  ) |>
  tab_style(
    style = cell_fill(color = "#D0EEFF60"),
    locations = list(
      cells_body(
        columns = `estimate_chest - glasses`,
        rows = `estimate_chest - glasses` > 0 &
          `p_chest - glasses` < sig.level
      ),
      cells_body(
        columns = `estimate_wrist - glasses`,
        rows = `estimate_wrist - glasses` > 0 &
          `p_wrist - glasses` < sig.level
      )
    )
  ) |>
  fmt(
    columns = c(arrow_main, arrow_chest, arrow_wrist),
    fns = \(x) lapply(x, gt::html)
  ) |>
  cols_hide(
    c(
      is_timing_metric, is_darkness_timing,
      normalization_denominator, normalization_basis,
      raw_estimate_chest, raw_estimate_wrist,
      matches("(BAUA|FUSPCEU|IZTECH|KNUST|RISE|THUAS|TUM|UCR)"),
      `p_chest - glasses`, `p_wrist - glasses`
    )
  ) |>
  sub_missing() |>
  tab_spanner(
    columns = c(
      `estimate_chest - glasses`, arrow_chest,
      `estimate_wrist - glasses`, arrow_wrist,
      p_position, r2_position
    ),
    "Chest/wrist versus glasses (estimated bias)"
  ) |>
  tab_spanner(
    columns = c(
      SD_chest, SD_wrist, r2_slope, SD_Id, r2_id, SD_Date,
      r2_date, SD_Residual, r2_residual
    ),
    "Standard deviation (random effect)"
  ) |>
  tab_spanner(columns = contains("_site"), "Site effect") |>
  tab_style(
    style = cell_text(weight = "bold"),
    locations = list(
      cells_body(
        columns = `estimate_wrist - glasses`,
        rows = `p_wrist - glasses` < sig.level
      ),
      cells_body(
        columns = `estimate_chest - glasses`,
        rows = `p_chest - glasses` < sig.level
      ),
      cells_body(columns = p_position, rows = p_position < sig.level),
      cells_body(columns = p_site, rows = p_site < sig.level),
      cells_body(
        columns = `p_site:position`,
        rows = `p_site:position` < sig.level
      ),
      cells_column_labels(),
      cells_column_spanners(),
      cells_stub(columns = "metric_family"),
      cells_group()
    )
  ) |>
  tab_footnote(
    "Scale on which the glasses reference and subsequent deviations (chest/wrist contrasts, site, random effects) are analyzed",
    locations = cells_column_labels(scaling)
  ) |>
  tab_footnote(
    timing_normalization_footnote,
    locations = list(
      cells_column_labels(`estimate_chest - glasses`),
      cells_column_labels(`estimate_wrist - glasses`),
      cells_column_labels(starts_with("SD_"))
    )
  ) |>
  tab_footnote(
    html("Partial R<sup>2</sup> values compare the effect model with a model without that effect. Values within a row do not sum to one because they come from different models and partially correlated components."),
    locations = cells_column_labels(contains("r2_"))
  ) |>
  tab_footnote(
    "No overall p-value correction has been performed, so as not to potentially underestimate individual metric deviations.",
    locations = cells_column_labels(contains("p_"))
  ) |>
  tab_footnote(
    "Missing entries refer to models that had to be reduced in complexity to avoid convergence/singularity issues.",
    locations = cells_column_spanners(2)
  ) |>
  tab_footnote(
    html(paste0(
      "Colored arrows indicate significant site-specific deviations above (↑) or below (↓) the column value. Arrow color by site: ",
      paste(
        mapply(
          \(site, color) sprintf(
            '<span style="color:%s; font-weight:bold">%s</span>',
            color,
            site
          ),
          paste0(
            melidos_cities[names(site_colors)],
            " (",
            c("DE", "ES", "TR", "GH", "SE", "NL", "DE", "CR"),
            ")"
          ),
          site_colors
        ),
        collapse = ", "
      )
    )),
    locations = list(
      cells_column_labels(main_effect),
      cells_column_labels(`estimate_chest - glasses`),
      cells_column_labels(`estimate_wrist - glasses`)
    )
  ) |>
  tab_header(
    html("Summary of outcome-level bias:<br>body-worn versus eye-level measurement"),
    subtitle = scenario_label
  ) |>
  tab_style(cell_text(color = colors[1]), cells_column_labels(main_effect)) |>
  tab_style(
    cell_text(color = colors[2]),
    cells_column_labels(`estimate_chest - glasses`)
  ) |>
  tab_style(
    cell_text(color = colors[3]),
    cells_column_labels(`estimate_wrist - glasses`)
  )
```

```{r}
#| label: display-main-table
#| filename: Save detailed wake-only result table
table
gtsave(
  table,
  here("output/tables/Table_metrics_wake_only.png"),
  vwidth = 2020
)
```

## Summary table

The reduced result data retain the normalized chest and wrist effects and the variance components needed for summaries. Timing values use the fixed-day denominator; other metrics use the absolute glasses reference.

```{r}
#| label: reduce-for-summary
#| filename: Reduce results for summaries
reduced_df <-
  df |>
  select(
    metric, metric_type, metric_family, scaling,
    p_chest = `p_chest - glasses`,
    p_wrist = `p_wrist - glasses`,
    error_chest = `estimate_chest - glasses`,
    error_wrist = `estimate_wrist - glasses`,
    p_position,
    r2_position,
    r2_id,
    SD_Id
  )

make_general_html <- function(
  r2_pos_med, r2_pos_min, r2_pos_max,
  r2_id_med, r2_id_min, r2_id_max,
  sd_id_med, sd_id_min, sd_id_max
) {
  pct <- function(x) paste0(round(x * 100), "%")
  pos_str <- if (is.na(r2_pos_med)) {
    "---"
  } else {
    paste0(
      pct(r2_pos_med), " (", pct(r2_pos_min), "–", pct(r2_pos_max), ")"
    )
  }
  id_str <- paste0(
    pct(r2_id_med), " (", pct(r2_id_min), "–", pct(r2_id_max), ")"
  )
  sd_str <- paste0(
    "±", round(sd_id_med * 100), "% (",
    round(sd_id_min * 100), "–", round(sd_id_max * 100), "%)"
  )
  paste0(
    "<b>R²<sub>Pos.</sub>:</b> ", pos_str, "<br>",
    "<b>R²<sub>Id</sub>:</b> ", id_str, "<br>",
    "<b>SD<sub>Id</sub>:</b> ", sd_str
  )
}
```

```{r}
#| label: summary-by-metric-type
#| filename: Summaries by metric type
df_summary <-
  reduced_df |>
  group_by(metric_type) |>
  mutate(across(starts_with("p_"), \(x) x <= 0.05)) |>
  summarize(
    n = n(),
    across(
      c(error_chest, error_wrist),
      list(
        median = \(x) median(abs(x), na.rm = TRUE),
        min = \(x) min(abs(x), na.rm = TRUE) |> replace_values(Inf ~ NA),
        max = \(x) max(abs(x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
        sign = \(x) sum(x > 0, na.rm = TRUE),
        sign_n = \(x) sum(x < 0, na.rm = TRUE),
        na = \(x) sum(is.na(x))
      )
    ),
    across(c(p_chest, p_wrist), sum),
    across(ends_with("_na"), \(x) n - x),
    r2_pos_median = median(r2_position[p_position], na.rm = TRUE),
    r2_pos_min = min(r2_position[p_position], na.rm = TRUE) |>
      replace_values(Inf ~ NA),
    r2_pos_max = max(r2_position[p_position], na.rm = TRUE) |>
      replace_values(-Inf ~ NA),
    r2_id_median = median(r2_id, na.rm = TRUE),
    r2_id_min = min(r2_id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    r2_id_max = max(r2_id, na.rm = TRUE) |> replace_values(-Inf ~ NA),
    sd_id_median = median(SD_Id, na.rm = TRUE),
    sd_id_min = min(SD_Id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    sd_id_max = max(SD_Id, na.rm = TRUE) |> replace_values(-Inf ~ NA)
  )
```

```{r}
#| label: summary-significant-only
#| filename: Summary of significant metrics
df_summary_significant <-
  reduced_df |>
  mutate(
    across(starts_with("p_"), \(x) x <= 0.05),
    error_chest = ifelse(p_chest, error_chest, NA),
    error_wrist = ifelse(p_wrist, error_wrist, NA)
  ) |>
  summarize(
    metric_type = "Overall (significant)",
    n = n(),
    across(
      c(error_chest, error_wrist),
      list(
        median = \(x) median(abs(x), na.rm = TRUE),
        min = \(x) min(abs(x), na.rm = TRUE) |> replace_values(Inf ~ NA),
        max = \(x) max(abs(x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
        sign = \(x) sum(x > 0, na.rm = TRUE),
        sign_n = \(x) sum(x < 0, na.rm = TRUE),
        na = \(x) NA_integer_
      )
    ),
    across(c(p_chest, p_wrist), sum),
    across(ends_with("_na"), \(x) n - x),
    r2_pos_median = median(r2_position[p_position], na.rm = TRUE),
    r2_pos_min = min(r2_position[p_position], na.rm = TRUE) |>
      replace_values(Inf ~ NA),
    r2_pos_max = max(r2_position[p_position], na.rm = TRUE) |>
      replace_values(-Inf ~ NA),
    r2_id_median = median(r2_id, na.rm = TRUE),
    r2_id_min = min(r2_id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    r2_id_max = max(r2_id, na.rm = TRUE) |> replace_values(-Inf ~ NA),
    sd_id_median = median(SD_Id, na.rm = TRUE),
    sd_id_min = min(SD_Id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    sd_id_max = max(SD_Id, na.rm = TRUE) |> replace_values(-Inf ~ NA)
  )
```

```{r}
#| label: summary-overall
#| filename: Overall summary
df_summary_overall <-
  reduced_df |>
  mutate(across(starts_with("p_"), \(x) x <= 0.05)) |>
  summarize(
    metric_type = "Overall",
    n = n(),
    across(
      c(error_chest, error_wrist),
      list(
        median = \(x) median(abs(x), na.rm = TRUE),
        min = \(x) min(abs(x), na.rm = TRUE) |> replace_values(Inf ~ NA),
        max = \(x) max(abs(x), na.rm = TRUE) |> replace_values(-Inf ~ NA),
        sign = \(x) sum(x > 0, na.rm = TRUE),
        sign_n = \(x) sum(x < 0, na.rm = TRUE),
        na = \(x) sum(is.na(x))
      )
    ),
    across(c(p_chest, p_wrist), sum),
    across(ends_with("_na"), \(x) n - x),
    r2_pos_median = median(r2_position[p_position], na.rm = TRUE),
    r2_pos_min = min(r2_position[p_position], na.rm = TRUE) |>
      replace_values(Inf ~ NA),
    r2_pos_max = max(r2_position[p_position], na.rm = TRUE) |>
      replace_values(-Inf ~ NA),
    r2_id_median = median(r2_id, na.rm = TRUE),
    r2_id_min = min(r2_id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    r2_id_max = max(r2_id, na.rm = TRUE) |> replace_values(-Inf ~ NA),
    sd_id_median = median(SD_Id, na.rm = TRUE),
    sd_id_min = min(SD_Id, na.rm = TRUE) |> replace_values(Inf ~ NA),
    sd_id_max = max(SD_Id, na.rm = TRUE) |> replace_values(-Inf ~ NA)
  )

df_summary <-
  bind_rows(df_summary, df_summary_overall, df_summary_significant) |>
  mutate(
    general = pmap_chr(
      list(
        r2_pos_median, r2_pos_min, r2_pos_max,
        r2_id_median, r2_id_min, r2_id_max,
        sd_id_median, sd_id_min, sd_id_max
      ),
      make_general_html
    )
  )
```

```{r}
#| label: tbl-summary-construction
#| filename: Construct wake-only summary table
table_summary <-
  df_summary |>
  gt(rowname_col = "metric_type") |>
  fmt_percent(
    columns = contains(c("error_chest", "error_wrist")) &
      contains(c("_min", "_max", "_median")),
    decimals = 0
  ) |>
  cols_merge(
    contains("chest"),
    pattern = "<b>bias: {1}</b> ({2} - {3})<br><b><<signif: {7}/{6}, >></b>direction: +{4}/-{5}"
  ) |>
  cols_merge(
    contains("wrist"),
    pattern = "<b>bias: {1}</b> ({2} - {3})<br><b><<signif: {7}/{6}, >></b>direction: +{4}/-{5}"
  ) |>
  fmt(columns = general, fns = \(x) lapply(x, gt::html)) |>
  cols_label(
    error_chest_median = "Chest",
    error_wrist_median = "Wrist",
    general = "General"
  ) |>
  tab_footnote(
    "median bias (min - max), significant bias / all metrics, positive / negative deviations"
  ) |>
  tab_footnote(
    paste0(
      timing_normalization_footnote,
      " Overall rows combine these normalization bases."
    ),
    locations = list(
      cells_column_labels(error_chest_median),
      cells_column_labels(error_wrist_median),
      cells_column_labels(general)
    )
  ) |>
  tab_footnote(
    html("R²<sub>Pos.</sub>: partial R² for the specific position effect (glasses, chest, or wrist; median min–max), significant metrics only. R²<sub>Id</sub> and SD<sub>Id</sub>: individual ID random effect, all metrics."),
    locations = cells_column_labels(general)
  ) |>
  cols_hide(
    c(
      n, r2_pos_median, r2_pos_min, r2_pos_max,
      r2_id_median, r2_id_min, r2_id_max,
      sd_id_median, sd_id_min, sd_id_max
    )
  ) |>
  tab_style(
    cell_text(weight = "bold"),
    list(cells_stub(), cells_column_labels())
  ) |>
  tab_header(
    title = html(
      "Summary overview of outcome-level bias<br>for body-worn versus eye-level measurement"
    ),
    subtitle = html(paste0(
      scenario_label,
      "<br>Timing metrics use mean wake/sleep durations;<br>",
      "other metrics to the absolute glasses reference."
    ))
  ) |>
  tab_options(table.width = px(900)) |>
  tab_style(
    cell_text(color = colors[2]),
    cells_column_labels(error_chest_median)
  ) |>
  tab_style(
    cell_text(color = colors[3]),
    cells_column_labels(error_wrist_median)
  ) |>
  tab_style(
    cell_fill(color = colors[2], alpha = 0.1),
    cells_body(error_chest_median)
  ) |>
  tab_style(
    cell_fill(color = colors[2], alpha = 0.5),
    cells_body(
      error_chest_median,
      metric_type %in% c("Overall", "Overall (significant)")
    )
  ) |>
  tab_style(
    cell_fill(color = colors[3], alpha = 0.1),
    cells_body(error_wrist_median)
  ) |>
  tab_style(
    cell_fill(color = colors[3], alpha = 0.5),
    cells_body(
      error_wrist_median,
      metric_type %in% c("Overall", "Overall (significant)")
    )
  ) |>
  tab_style(
    cell_borders(sides = "top", color = "black", weight = px(2.5)),
    locations = list(
      cells_body(rows = metric_type == "Overall"),
      cells_stub(rows = "Overall")
    )
  )
```

```{r}
#| label: display-summary-table
#| filename: Save wake-only summary table
table_summary
gtsave(
  table_summary,
  here("output/tables/Table_metrics_summary_wake_only.png"),
  vwidth = 2000
)
```

## Visualize normalized results

The absolute and signed histograms use the normalized values shown in the tables. Significant metric-specific contrasts are highlighted and labelled.

```{r}
#| label: fig-absolute-errors
#| filename: Visualize absolute wake-only metric errors
#| fig-height: 4
#| fig-width: 8
color_strips <- function(plot, fills) {
  grob <- ggplotGrob(plot)
  strip_idx <- grep("strip-t", grob$layout$name)
  for (i in seq_along(strip_idx)) {
    grob$grobs[[strip_idx[i]]]$grobs[[1]]$children[[1]]$gp$fill <- fills[i]
  }
  grob
}

normalization_caption <- paste(
  paste0(
    "Brightness-based timing metrics: mean wake duration ",
    mean_wake_duration_label, " (1% = ", mean_wake_one_percent_label, ")."
  ),
  paste0(
    "Darkness-based timing metrics: mean sleep duration ",
    mean_sleep_duration_label, " (1% = ", mean_sleep_one_percent_label, ")."
  ),
  "All other metrics: absolute glasses reference.",
  sep = "\n"
)
combined_bias_caption <- paste(
  "Bin width = 5%; significant errors are highlighted in the position color.",
  normalization_caption,
  sep = "\n"
)

fig_abs_error <-
  reduced_df |>
  select(metric, p_chest:error_wrist) |>
  pivot_longer(-metric, names_sep = "_", names_to = c("type", "position")) |>
  pivot_wider(names_from = type, values_from = value) |>
  mutate(p = p < 0.05) |>
  ggplot(aes(x = abs(error))) +
  geom_histogram(binwidth = 0.05, aes(fill = position)) +
  facet_wrap(~position, labeller = labeller(position = str_to_title)) +
  scale_fill_manual(values = colors) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(0, 1, by = 0.2)
  ) +
  theme_cowplot() +
  coord_cartesian(xlim = c(0, 1)) +
  labs(
    y = "Count (n)",
    x = "Absolute normalized outcome-level bias (%)"
  ) +
  theme(panel.spacing = unit(2, "lines")) +
  gghighlight(
    p,
    use_group_by = FALSE,
    label_key = metric,
    calculate_per_facet = TRUE
  )

fig_abs_error +
  labs(subtitle = scenario_label, caption = normalization_caption) +
  theme(plot.caption = element_text(hjust = 0, size = 9))
```

```{r}
#| label: fig-signed-errors
#| filename: Visualize signed wake-only metric errors
#| fig-height: 4
#| fig-width: 8
fig_sign_error <-
  reduced_df |>
  select(metric, p_chest:error_wrist) |>
  pivot_longer(-metric, names_sep = "_", names_to = c("type", "position")) |>
  pivot_wider(names_from = type, values_from = value) |>
  mutate(p = p < 0.05) |>
  ggplot(aes(x = error)) +
  geom_histogram(binwidth = 0.05, aes(fill = position)) +
  facet_wrap(~position, labeller = labeller(position = str_to_title)) +
  scale_fill_manual(values = colors) +
  scale_x_continuous(
    labels = scales::label_percent(accuracy = 1),
    breaks = seq(-1, 1, by = 0.2)
  ) +
  theme_cowplot() +
  coord_cartesian(xlim = c(-0.8, 0.5)) +
  labs(
    y = "Count (n)",
    x = "Normalized outcome-level bias (%)"
  ) +
  theme(panel.spacing = unit(2, "lines")) +
  gghighlight(
    p,
    use_group_by = FALSE,
    label_key = metric,
    calculate_per_facet = TRUE
  )

fig_sign_error +
  labs(subtitle = scenario_label, caption = normalization_caption) +
  theme(plot.caption = element_text(hjust = 0, size = 9))
```

```{r}
#| label: fig-errors
#| filename: Combine and save wake-only metric errors
#| fig-height: 8
#| fig-width: 8
strip_fills <- unname(colors[c("chest", "wrist")])

combined_error_figure <-
  wrap_elements(color_strips(fig_sign_error, strip_fills)) /
  wrap_elements(color_strips(fig_abs_error, strip_fills)) +
  plot_annotation(
    tag_levels = "A",
    title = scenario_label,
    caption = combined_bias_caption,
    theme = theme(
      plot.title = element_text(hjust = 0, size = 14),
      plot.caption = element_text(hjust = 0, size = 9),
      plot.margin = margin(8, 12, 8, 12)
    )
  )

combined_error_figure

ggsave(
  here("output/figures/Fig_metric_error_wake_only.pdf"),
  plot = combined_error_figure,
  device = cairo_pdf,
  width = 8,
  height = 8
)
ggsave(
  here("output/figures/Fig_metric_error_wake_only.jpeg"),
  plot = combined_error_figure,
  width = 8,
  height = 8
)
```

## Session info

```{r}
#| label: session-info
sessionInfo()
```
